### 2022-05-27 article.tex
Start of LaTeX

\begin{abstract}

Here we define paradata as the data that describes the generation of data.
In genetic epidemiology, the data generated is mostly the results 
of an analysis (e.g. predicting a person having a disease),
as done by computer code.
In such context, paradata is usually the scientific paper that
describes what the computer code does.
However, this has the unrealtistic 
assumption that there is a perfect match between the paper and the code.
In this chapter it is argued that the source code should should be supplied,
as this is the true paradata: if the paper and code disagree, it is the
code that has generated the results.
The chapter concludes by some rules how to allow code to serve as paradata,
and hence allowing computational research to be truly reproducible.

\end{abstract}

{\bf Keywords:} paradata, reproducible research, code, software,
FAIR data, computational research, Open Science, DevOps, best practices,
genetic epidemiology

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abbreviations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[h]
  \begin{tabular}{ll}
    Abbreviation & Full                                    \\
    \hline
    AI           & Artificial intelligence                 \\
    GCAE         & Genotype convolutional autoencoder      \\
    GWAS         & Genome-wide association study           \\
    ML           & Machine learning                        \\
    PCA          & Principal Component Analysis
  \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{Karesuando_church.jpg}
  \caption{Karesuando's church}
  \label{fig:karesuando_church}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We found ...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This article and its metadata can be found at 
\url{https://github.com/richelbilderbeek/chapter_paradata}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Vancouver style
\bibliographystyle{unsrtnat}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\section{Supplementary materials}

% Figures start from one and are prepended with an S
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

% Tables start from one and are prepended with an S
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Datasets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Stuff

### 2022-05-27 tab_snp_associations.tex
Start of LaTeX

% In the end, use a latex table generated in R by xtable
\begin{table}[ht]
\centering
\begin{tabular}{rlllll}
  \hline
  SNP       & Primary/Conditional hit & Cis\_Trans & Effect strength & SE   & P-value \\
  \hline
  rs1653281 & Primary                 & Trans     & 0.1             & 0.01 & 2.90E-09 \\
  rs2701020 & Primary                 & Trans     & 0.2             & 0.02 & 2.11E-10 \\
  rs3875965 & Primary                 & Trans     & -1.3            & 0.13 & 2.62E-11 \\
  rs4984576 & Primary                 & Trans     & 0.4             & 0.14 & 4.26E-12 \\
  rs5058124 & Primary                 & Cis       & -0.5            & 0.05 & 3.71E-13 \\
  \hline
\end{tabular}
\caption{
  SNP association.
  Effect strength = some measure of effect strength
  \richel{
    I adapted this from the SKAT biomarkers paper. I think the
    data will look different (e.g. the first column should have
    single or multiple SNPs, as their effects may only occur when combined)
  }
} 
\label{tab:snp_associations}
\end{table}

### 2022-05-27 scripts/create.sh
Add script to create LaTeX

@@ -1,0 +1,16 @@
+#!/bin/bash
+#
+# Create the PDF 'article.pdf'
+#
+# It does so by converting the TeX to PDF, 
+# as well as stitching some other PDFs to it
+#
+# Usage:
+#
+#   ./scripts/create.sh
+#
+pdflatex header.tex
+bibtex header >/dev/null
+pdflatex header.tex >/dev/null
+pdflatex header.tex
+mv header.pdf article.pdf

### 2022-05-29 scripts/edit.sh
Bibliography is now in article.bib

@@ -1,0 +1,2 @@
+#!/bin/bash
+mousepad header.tex article.bib article.tex

### 2022-05-29 article.tex
Convert to LaTeX

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Abbreviations}
+\section*{Definitions}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{table}[h]
-  \begin{tabular}{ll}
-    Abbreviation & Full                                    \\
+  \begin{tabular}{lll}
+    Term      & Definition                                            & Example                                                         \\
     \hline
-    AI           & Artificial intelligence                 \\
-    GCAE         & Genotype convolutional autoencoder      \\
-    GWAS         & Genome-wide association study           \\
-    ML           & Machine learning                        \\
-    PCA          & Principal Component Analysis
+    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
+    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
+    Paradata  & Data that describes the generation of data            & The code to conclude that a SNP has a significant association   \\
+    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
+    Metadata  & Data that provides information about other data       & The article that describes an experiment                        \\
+    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
   \end{tabular}
@@
 
+% This section has paragraph headers. This helps me, the author,
+% to focus the contents of my paragraphs.
+% In the final version of this manuscript, these will be removed.
+
+\paragraph{Universities produce knowledge}
+
+The currency that universities generate is knowledge, which in turn is distributed
+in mostly lectures and academic papers.
+Academic papers contain knowledge, either as the
+report of measurements or the outcome of an experiment.
+These reports are described in English, the current lingua franca of science.
+
+\paragraph{A paper contains data and should be FAIR data}
+
+A paper contains and/or describes data, where we define
+data as 'individual facts, statistics, or items of information, often 
+numeric' [OECD 2008].
+To ensure unhampered knowledge growth, this data should [TODO: Why? Reference!] 
+follow the FAIR 
+principles.
+The FAIR principles [Wilkinson et al., 2016] are that data should be 
+[TODO: there are probably better examples]
+Findable (e.g. found by web),
+Accessible (e.g. DataDryad and or Zenoda), 
+Interoperable (e.g. comma-separated files) 
+and Reusable (e.g. work on all operating systems)
+
+\paragraph{A paper is data and should be FAIR data}
+
+A paper in itself is data, as it is a collection of text and
+figures that we itself can do measurements upon.
+Also here, a paper should be FAIR, 
+[TODO: there are probably better examples]
+i.e.
+Findable (e.g. Google Scholar)
+Accessible (e.g. view it online), 
+Interoperable (e.g. written in English)
+and Reusable (e.g. it is valid to cite and build upon earlier academic papers).
+This paper itself will later specify the standpoint that a paper
+is metadata (i.e. data about data), instead of paradata (i.e.
+data that describes the collection of data)
+
+\paragraph{For some fields, the experiment is actually run by code}
+
+This paper focusses on research fields that uses computation
+to do experiments and uses genetic epidemiology as an example.
+Genetic epidemiology is a field within biology that, among 
+others, measures the spread (hence 'epidemiology') 
+of heritable (i.e. genetic) traits, as well as the relation
+between having a certain genetic makeup and a certain trait,
+where the trait can be any human property, such as weight, height,
+the amount of metabolites and having a disease yes/no.
+
 \begin{figure}[!htbp]
@@
   \includegraphics[width=\linewidth]{Karesuando_church.jpg}
-  \caption{Karesuando's church}
+  \caption{
+    Figure F2: Picture of Karesuando's church,
+    the village where the Northern Swedish Population
+    Health Study started.
+    From \cite{hopfner2005}
+  }
   \label{fig:karesuando_church}
 \end{figure}
+
+As an example, we'll use a pseudorandomly selected paper
+from [Ahsan et al., 2017]. The data used by that paper is
+from a population study called the Northern Swedish Population
+Health Study (NSPHS) that started in 2010 [Igl et al., 2010]. 
+The approximately 1000 participants were initially mostly surveyed
+about lifestyle [Igl et al., 2010] and follow-up studies
+provided the type of data relevant for this paper, 
+which are (1) the genetic data [Johansson et al., 2013],
+(2) the concentration of certain proteins in the 
+blood [Enroth et al., 2014][Enroth et al., 2015].
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{1189px-Eukaryote_DNA-en.png}
+  \caption{
+    Fig F3: A cell has a nucleus that contains chromosomes. 
+    Each of these chromosomes (46 in humans) consist out of DNA. 
+    DNA itself conists out of 4 nucleotides, 
+    as depicted by the horizontal sticks 
+    with the colors red, yellow, green and blue.
+    From [Sponk et al., 2012]
+  }
+  \label{fig:eukakyote_dna}
+\end{figure}
+
+The first type of data, the (final form of the) genetic data,
+consists out of single nucleotide polymorphisms (SNPs).
+SNPs consist out of a name, a position and a nucleotide. 
+DNA (organized into chromosomes and present in every (nucleated) cell, 
+see figure F3) consists out of many nucleotides, of which there are four types,
+called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
+as A, C, G and T respectively.
+One SNP example is, that 'rs12133641' is a SNP located at position 154,428,283,
+where some people have a certain nucleotide. In this case, 
+67 percent of the people within this study have an A,
+and 33 percent have a G (also from [Ahsan et al., 2017], Table S3).
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{DNA_to_protein_or_ncRNA.png}
+  \caption{
+    Fig F5: Parts of DNA (so-called 'genes') code for proteins. 
+    The DNA, that always stays put in the cell's nucleus, 
+    is transcripted to messenger RNA (mRNA).
+    mRNA leaves the nucleus and its code gets translated to 
+    a protein sequence.
+    Near the start of a gene are regions that determine the amount
+    of proteins produced (not shown in figure).
+    Adapted from [Shafee, 2015]
+  }
+  \label{fig:dna_to_protein}
+\end{figure}
+
+The second type of data are concentrations of certain proteins in the 
+blood. DNA contains the code for building proteins (see Figure F5), 
+as well as the rate
+at which a protein is created. Some proteins end up in the blood and
+their presence can be used to assess the health of an individual.
+IL6RA is a protein [TODO: more info]
+
+The field of genetic epidemiology looks -among others- for
+correlations between genetic data and biological traits.
+For example, figure F10 (from [Ahsan et al., 2017]) shows that
+SNP rs12133641 is highly correlated (p value is 3.0 * 10e-73, 
+961 individuals) with protein IL6RA. What this results does
+not yet teach us, is how this correlation works, yet
+figure F20 shows us the direction of the association:
+the X axis shows the possible genetic make-ups (aka the genotype) of the individuals,
+where 0 denotes the individuals with genotype
+AA (the individual inherited one adenosine 
+from his/here mother and one adenosine from his/her father), 
+1.0 denotes AG (one A is inherited from one parent, 
+where the G is inherited from the other parent) and 2.0 denotes GG.
+From Figure F20 we can conclude that, on average,
+the more guanines are inherited at that SNPs location,
+the higher concentration of IL6RA can be found in a human's blood.
+
+The amount of variance that can be explained by an association (i.e.
+the R squared value) is rarely 100 percent, which means that a trait (in
+this case, IL6RA concentrations) cannot be perfectly explained
+from the genotype alone. As we can see in figure F20, 
+43 percent of the variance can be attributed to an individuals'
+genotype and additional factors, 
+such as the effect
+of the environment (e.g. geographic location, time of day), 
+lifestyle (e.g. smoking yes/no) or having a disease (e.g. diabetes) 
+are needed to explain the variation between individuals better.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{ahsan2017relative_table_2_sub.png}
+  \caption{
+    Figure F10 shows an example result of a genetic epideological research.
+    It shows that the SNP named rs12133641 (located at position 154,428,283
+    of chromosome 1) is highly correlated (p value is 3.0 * 10e-73, 
+    961 individuals) to the concentration of the protein IL6RA, as measured
+    in blood. The table is a simplified result from [Ahsan et al., 2017].
+  }
+  \label{fig:ahsan2017relative_table_2_sub}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{ahsan2017relative_s6.png}
+  \caption{
+    Figure F20: the relation between the genotype for SNP rs12133641 
+    and the protein concentration of IL6RA is relatively strong.
+    The X axis shows the the genotype of the individuals,
+    where 0 denotes AA, 1.0 denotes AG and 2.0 denotes GG.
+    The Y axis shows the concentration of the protein IL6RA 
+    as found in the participants' blood. n = 1 denotes the number
+    of SNPs that were determined to be involved.
+  }
+  \label{fig:ahsan2017relative_s6}
+\end{figure}
+
+\paragraph{The experiments within genetic epidemology works are done by code}
+
+The experiment described above is run by code. There is no fieldwork,
+nor lab-work involved. It is the researcher that writes the code,
+after which the code does the work. A genetic epidemiologist that
+looks for associations between genotype and phenotype,
+one does not need a lab.
+
+To **obtain** the genetic data, yes, there may be fieldwork and/or lab-work
+involved. For example, a researcher needs to go into the field to take
+samples (e.g. blood) of humans or other species. To analyse these samples,
+a researcher needs a lab. Depending on the technique being used on
+how to analyse the samples, there may be a big bioinformatics step
+to aggregate the measurements into useful genetic data. 
+Although this paper does not focus on the code run to 
+aggragate the raw data into useful genetic data, the same arguments
+can be made for that code as in the example: that code is 
+the ground truth of how the raw data is collected into useful genetic data.
+
+\paragraph{The code is the paradata for the results of an experiment}
+
+Programming code is paradata, as it is data that describes how data is collected,
+which is the definition of paradata [TODO: REF].
+Programming code is data that is usually in the form of text, 
+spread over one or more files, that describes the experiment.
+The experiment collects the data.
+The data collected is the data we call the results of an experiment.
+
+The article is metadata, as it is data that describes other
+data, which is the definition of metadata [TODO: REF]:
+an article describes the
+experiment and hence the programming
+code in English. However, it is not the best
+candidate to describe how the data is collected,
+as it has a loose connection with collecting the data.
+Again, if the code and article of an experiment disagree,
+it is the code that actually let the data be collected.  
+Instead, an article is metadata about a research.
+
+\paragraph{Code should be published}
+
+For science to be reproducible (a fundamental property of the scientific 
+method), it is the code that should be published, instead of the
+English description of what the code does (i.e. the article), if one
+would have to pick one, as it is the code does the actual work.
+Hence, from a knowledge management perspective, emphasis should be
+put more on the preservation of code, as it is the most important actor
+in an experiment.
+
+To quote a review on reproducibility [Peng and Hicks, 2021], 
+'Reproducibility is typically thwarted by a 
+lack of availability of the original data and computer code'.
+In the case of genetic epidemiology, it is a given that the original
+data cannot be published as this data is sensitive.
+
+\paragraph{Example of code being the ground truth}
+
+Code should be published, as it holds the
+ground truth of an experiment; it does the actual work.
+The more complex the computation pipeline is, the easier it is
+to have a mismatch between the article (that describes what the
+code does) and the code (that actually does the work).
+It is easy (tempting?) to overlook how easy it is to have a mismatch
+between paper and code.
+
+To illustrate how easy it is to get a mismatch between a paper
+and the code, consider this fictional example
+of a text in a paper:
+
+> We compared the values of `x` and `y` using a one-tailed Student's T-test,
+> as we expect the average value of `x` to be less than the average value of `y`.
+
+We ignore the choices of words and style of this sentence: what is
+important is the content: a one-tailed T test is performed.
+Taking a look at the (in this case, the programming language R) code, 
+we find the following line:
+
+\begin{verbatim}
+t.test(x, y)
+\end{verbatim}
+
+The line of code is simplified (as the result of the T test is never stored),
+but the code is correct: 
+`t.test` is indeed the name of an R function to do a T-test
+and it is reasonable to assume that this code, when found in the code
+of an article, is assumed to be correct.
+
+Here, however, we see a mismatch between the English description and the code:
+by default, the R function `t.test` does a **two**-tailed T-test.
+A consequence of this fictional example is that the published p-values are
+higher than needed, resulting in less significant findings, which
+results in less conclusions drawn from a paper.
+
+It is the paper that accompanies the code,
+as it is the code that generates the results.
+When humans are fallible and code gets bigger, the likelihood of
+a mismatch between the English paper and the code increases.
+But regardless of the size of the code, 
+it is the code that is the ground truth.
+
+\paragraph{Code itself has important properties}
+
+The code of a computational experiment is
+more important to preserve than the article describing
+the experiment, even when both describe the experiment identically,
+correctly and reproducibly. 
+Here I will argue why this is the
+case and the implications and best practices for knowledge preservation.
+To summarize, (1) it gives an honest indication of the quality 
+of a computational experiment, ....
+
+Optionally, code can actively teach what it does.
+Here, I will go into more detail of each point.
+
+\paragraph{Example of code being convincing}
+
+As code is paradata, as it describes how data is being collected,
+it can convince (or not) that it does so correctly.
+Similar to a cell biologist that can be working sterile to avoid contamination
+by airborne bacteria, 
+or sloppy and likelier to have contaminated samples,
+code can be examplary or sloppy.
+
+\paragraph{Code is hard to manage}
+
+Unlike English, code is fragile in standing the test of time.
+Containers can alleviate this.
+Code has a limited lifetime.
+
+\paragraph{Hosted code has important metadata}
+
+ * Number of commits
+ * Commit history
+ * Code coverage
+ * Objective indicator of work being done
+
+\paragraph{Runnable code has other important features}
+
+ * Reproduce an experiment
+ * Code can teach the user, see babette
+ * Build upon an experiment
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The code that does the work in a computational experiment should
+always be published together with a scholarly article.
+For research to be reproducible, one ideally has access to
+both the data used and the code.
+In some fields, such as genetic epidemiology, the data is
+sensitive, hence cannot be released,
+yet there are methods being devised to run code on sensitive
+data with assured privacy [Zhang et al., 2016][Azencott, 2018].
+
+Code has additional useful information, similar to confidence intervals,
+that allow a reader to gauge how much he/she trusts the results.
+The most important way to determine the quality of code
+is the amount of unit tests.
+When following a set of best practices, such as DevOps, TDD, Agile,
+writing unit tests is an essential 
+part in writing code.
+The amount of unit tests is an honest signal 
+for code correctness (i.e. it does what it is supposed to do, as opposed
+to 'it does something').
+In academia, to uncover the truth, code correctness is essential.
+To make a comparison with cell biologists, where working
+sterile is essential to have the correct results,
+unit tests allow a computational biologist to have the same.
+
+Code is harder to preserve than an English text.
+
+Although code is the primary actor in computational experiments,
+there is no incentive to submit code alongside a publication.
+Most academic journal do not require authors to submit their code,
+nor it the submitted code peer reviewed.
+
+Although the code of computation experiments can be archived well, 
+there is no incentive to do so.
+
+Although a runnable version of a computation experiment can be archived well, 
+there is no incentive to do so.
 
@@
 
-We found ...
+\paragraph{Definition}
+
+Code may or may not be paradata, depending on how the definition
+is interpreted.
+Here we repeat the definition of 'paradata' and discuss 
+its (numbered) constituents.
+This paper defined paradata as 'data (1) about the collecting (2) of the data (3)',
+where (1) code must be seen as data, (2) downloading raw data
+and doing calculations must be seen as collecting, and (3) the
+results of an experiment must be seen as data.
+This paper argues, that (1) code is data in the form of text spread
+over one or more files, that has useful measurable properties, 
+(2) downloading raw datas and doing calculations, such as a T-test,
+does describe how the bits and pieces of an end result is collected,
+and (3) an experimental results is data, as it can be measured and
+used as the raw data of a next experiment.
+
+\paragraph{The drawbacks of publishing code}
+
+Publishing code may be disadvantageous for an author.
+For science, yes, as this allows reproducible research.
+For an author, publishing code alongside an experiment opens up
+the possibily to receive questions regarding that code.
+For a publisher
+For a knowledge archivist,
+
+\paragraph{The drawbacks of publishing version-controlled code}
+
+Is it worth it to publish version-controlled code?
+For an author, 
+there is additional training involved, and also here,
+publishing code alongside an experiment opens up
+the possibily to receive questions regarding that code.
+
+\paragraph{The drawbacks of publishing a running version of code}
+
+Is it worth it to archive running versions of code?
+For an author, 
+there is additional training involved.
+There is no FAIR infrastructure for Singularity containers.
+
+\paragraph{Recommendations}
+
+Being able to re-do an experiment is a core principle of the scientific method.
+Publishing only a paper about a computational experiment is not enough,
+as the results are too likely to mismatch that English description.
+Journal should make it mandatory for authors
+to publish their code alongside a computational experiment.
+Authors should follow the FAIR principes for their code as well,
+as can be done using the infrastructure as supplied by, 
+for example, GitHub and GitLab, opening up new angles in
+metascience regarding computational experiments.
+Knowledge managers should create the infrastructure for the preservation
+of runnable experiments, to allow scientist to upload a Singularity
+container, so these are as well following the FAIR principles.
+
+\paragraph{Final word}
+
+The world of science would be a more open, humble, trustworthy, truthful
+and helpful would the code that accompanies a scientific paper
+be treated like a first class citizen. As doing so in an exemplary way
+in yet to be rewarded, hence it has to be the idealististic scientists
+to wage this battle. I feel the truth and science are worth fighting for
+and I hope this paper helps others to join.
 

### 2022-05-29 scripts/edit.sh

@@ -1,2 +1,2 @@
 #!/bin/bash
-mousepad header.tex article.bib article.tex
+mousepad header.tex header.bib article.tex

### 2022-05-29 scripts/view.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+evince article.pdf

### 2022-05-29 article.tex
Must be 'article.bib'

@@
 data as 'individual facts, statistics, or items of information, often 
-numeric' [OECD 2008].
+numeric' \cite{glossary2003oecd}.
 To ensure unhampered knowledge growth, this data should [TODO: Why? Reference!] 
@@
 principles.
-The FAIR principles [Wilkinson et al., 2016] are that data should be 
+The FAIR principles \cite{wilkinson2016fair} are that data should be 
 [TODO: there are probably better examples]

### 2022-05-29 scripts/edit.sh
Must be 'article.bib'

@@ -1,2 +1,2 @@
 #!/bin/bash
-mousepad header.tex header.bib article.tex
+mousepad header.tex article.bib article.tex

### 2022-05-29 article.tex
All refs are in

@@
 As an example, we'll use a pseudorandomly selected paper
-from [Ahsan et al., 2017]. The data used by that paper is
+from \cite{ahsan2017relative}. The data used by that paper is
 from a population study called the Northern Swedish Population
-Health Study (NSPHS) that started in 2010 [Igl et al., 2010]. 
+Health Study (NSPHS) that started in 2010 \cite{igl2010northern}. 
 The approximately 1000 participants were initially mostly surveyed
-about lifestyle [Igl et al., 2010] and follow-up studies
+about lifestyle \cite{igl2010northern} and follow-up studies
 provided the type of data relevant for this paper, 
-which are (1) the genetic data [Johansson et al., 2013],
+which are (1) the genetic data \cite{johansson2013identification},
 (2) the concentration of certain proteins in the 
-blood [Enroth et al., 2014][Enroth et al., 2015].
+blood \cite{enroth2014strong,enroth2015effect}.
 
@@
 67 percent of the people within this study have an A,
-and 33 percent have a G (also from [Ahsan et al., 2017], Table S3).
+and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).
 
@@
 correlations between genetic data and biological traits.
-For example, figure F10 (from [Ahsan et al., 2017]) shows that
+For example, figure F10 (from \cite{ahsan2017relative}) shows that
 SNP rs12133641 is highly correlated (p value is 3.0 * 10e-73, 
@@
     961 individuals) to the concentration of the protein IL6RA, as measured
-    in blood. The table is a simplified result from [Ahsan et al., 2017].
+    in blood. The table is a simplified result from \cite{ahsan2017relative}.
   }
@@
 
-To quote a review on reproducibility [Peng and Hicks, 2021], 
+To quote a review on reproducibility \cite{peng2021reproducible}, 
 'Reproducibility is typically thwarted by a 
@@
 but the code is correct:
-`t.test` is indeed the name of an R function to do a T-test
+\verb{t.test} is indeed the name of an R function to do a T-test
 and it is reasonable to assume that this code, when found in the code
@@
 Here, however, we see a mismatch between the English description and the code:
-by default, the R function `t.test` does a **two**-tailed T-test.
+by default, the R function \verb{t.test} does a \textbf{two}-tailed T-test.
 A consequence of this fictional example is that the published p-values are
@@
 yet there are methods being devised to run code on sensitive
-data with assured privacy [Zhang et al., 2016][Azencott, 2018].
+data with assured privacy \cite{zhang2016review,azencott2018machine}.
 

### 2022-05-29 article.tex
Fix refs

@@
   \caption{
-    Figure F2: Picture of Karesuando's church,
+    Picture of Karesuando's church,
     the village where the Northern Swedish Population
@@
   \caption{
-    Fig F3: A cell has a nucleus that contains chromosomes. 
+    A cell has a nucleus that contains chromosomes. 
     Each of these chromosomes (46 in humans) consist out of DNA. 
@@
     with the colors red, yellow, green and blue.
-    From [Sponk et al., 2012]
+    From \cite{sponk2012}
   }
@@
 DNA (organized into chromosomes and present in every (nucleated) cell, 
-see figure F3) consists out of many nucleotides, of which there are four types,
+see figure \ref{fig:eukakyote_dna}) consists out of many nucleotides, of which there are four types,
 called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
 as A, C, G and T respectively.
-One SNP example is, that 'rs12133641' is a SNP located at position 154,428,283,
+One SNP example is \verb|rs12133641|, which is a SNP located at position 
+154,428,283,
 where some people have a certain nucleotide. In this case, 
@@
   \caption{
-    Fig F5: Parts of DNA (so-called 'genes') code for proteins. 
+    Parts of DNA (so-called 'genes') code for proteins. 
     The DNA, that always stays put in the cell's nucleus, 
@@
     of proteins produced (not shown in figure).
-    Adapted from [Shafee, 2015]
+    Adapted from \cite{shafee2015}
   }
@@
 The second type of data are concentrations of certain proteins in the 
-blood. DNA contains the code for building proteins (see Figure F5), 
+blood. DNA contains the code for building proteins (see Figure 
+\ref{fig:dna_to_protein}), 
 as well as the rate
@@
 correlations between genetic data and biological traits.
-For example, figure F10 (from \cite{ahsan2017relative}) shows that
-SNP rs12133641 is highly correlated (p value is 3.0 * 10e-73, 
+For example, figure \ref{fig:ahsan2017relative_table_2_sub} (from \cite{ahsan2017relative}) shows that
+SNP \verb|rs12133641| is highly correlated (p value is 3.0 * 10e-73, 
 961 individuals) with protein IL6RA. What this results does
 not yet teach us, is how this correlation works, yet
-figure F20 shows us the direction of the association:
+figure \ref{fig:ahsan2017relative_s6} shows us the direction of the association:
 the X axis shows the possible genetic make-ups (aka the genotype) of the individuals,
@@
 where the G is inherited from the other parent) and 2.0 denotes GG.
-From Figure F20 we can conclude that, on average,
+From Figure \ref{fig:ahsan2017relative_s6} we can conclude that, on average,
 the more guanines are inherited at that SNPs location,
@@
 this case, IL6RA concentrations) cannot be perfectly explained
-from the genotype alone. As we can see in figure F20, 
+from the genotype alone. As we can see in figure \ref{fig:ahsan2017relative_s6}, 
 43 percent of the variance can be attributed to an individuals'
@@
   \caption{
-    Figure F10 shows an example result of a genetic epideological research.
+    An example result of a genetic epideological research.
     It shows that the SNP named rs12133641 (located at position 154,428,283
@@
   \caption{
-    Figure F20: the relation between the genotype for SNP rs12133641 
+    The relation between the genotype for SNP rs12133641 
     and the protein concentration of IL6RA is relatively strong.
@@
 
-> We compared the values of `x` and `y` using a one-tailed Student's T-test,
-> as we expect the average value of `x` to be less than the average value of `y`.
+\begin{verbatim}
+We compared the values of x and y using a one-tailed Student's T-test,
+as we expect the average value of x to be less than the average value of y.
+\end{verbatim}
 
@@
 but the code is correct:
-\verb{t.test} is indeed the name of an R function to do a T-test
+\verb|t.test| is indeed the name of an R function to do a T-test
 and it is reasonable to assume that this code, when found in the code
@@
 Here, however, we see a mismatch between the English description and the code:
-by default, the R function \verb{t.test} does a \textbf{two}-tailed T-test.
+by default, the R function \verb|t.test| does a \textbf{two}-tailed T-test.
 A consequence of this fictional example is that the published p-values are

### 2022-05-29 article.tex
Progress

@@
 the amount of metabolites and having a disease yes/no.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Example study}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Code is more important than the (English) paper}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \paragraph{The experiments within genetic epidemology works are done by code}
@@
 
-To **obtain** the genetic data, yes, there may be fieldwork and/or lab-work
+To \emph{obtain} the genetic data, yes, there may be fieldwork and/or lab-work
 involved. For example, a researcher needs to go into the field to take
@@
 
-\paragraph{Code itself has important properties}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{There are multiple way to publish code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The code of a computational experiment is
+There are multiple ways of publishing code,
+that are increasingly hard to do, yet result
+in more reproducible experiments:
+(1) publishing the code as text only, e.g. by printing
+it in a Supplementary Materials, (2) publishing
+the code on a hosted website, e.g on GitHub, with
+added features (3) publishing 
+the installed code within a virtual environment,
+for example, using a Singularity container.
+Here I will go into details of each of these ways.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Publishing raw code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Code is commonly published alongside of a 
+paper, pasted as text in the Supplementary Materials 
+(for example, in \cite{labrecque2019interpretation}).
+This is the rawest form of publishing code,
+as it is just text.
+
+\paragraph{Raw code is important}
+
+The raw code of a computational experiment is
 more important to preserve than the article describing
 the experiment, even when both describe the experiment identically,
-correctly and reproducibly. 
-Here I will argue why this is the
-case and the implications and best practices for knowledge preservation.
-To summarize, (1) it gives an honest indication of the quality 
-of a computational experiment, ....
+correctly and reproducibly, 
+as it has/should been that code that generated the results of an experiment.
+Again: if the paper and code differ, it is the code that did
+the actual work with the actual/true results.
 
-Optionally, code can actively teach what it does.
-Here, I will go into more detail of each point.
+\paragraph{Raw code is an indicator of quality}
 
-\paragraph{Example of code being convincing}
+Raw code gives an indication of the quality 
+of a computational experiment and several metrics are
+devised to get an idea of the quality of code.
 
-As code is paradata, as it describes how data is being collected,
-it can convince (or not) that it does so correctly.
-Similar to a cell biologist that can be working sterile to avoid contamination
-by airborne bacteria, 
-or sloppy and likelier to have contaminated samples,
-code can be examplary or sloppy.
+The simplest metric to assess the quality of code is 
+the amount of (single) lines of code (SLOC). 
+Code that has a lower SLOC count, i.e. it is more concise,
+is usually regarded as having a lower chance of containing bugs.
+
+A more interesting metric is the cyclomatic complexity of the code.
+The cyclomatic complexity is defined as the number of ways that
+the code can be executed. 
+For example, code that has only one \verb|if| statement
+has a cyclomatic complexity of 2, as the condition within the \verb|if|
+statement can be true or false.
+The cyclomatic complexity correlates with code complexity,
+where more complex code is likelier to contain bugs.
+
+A feature that deserves more attention
+is that code can actively teach what it does.
+A mechanism to do so, which is present in most programming
+languages, is error handling: errors are the mechanism
+by which a programmer can teach the user what to do/expect.
+One example is when calculating the variance of a distribution.
+To calculate the variance of a distribution, at least two values
+are required. When the variance of only one value is requested,
+an error message can helpfully indicate that one needs at least two values.
+
+\paragraph{The problems with raw code}
+
+There are multiple problems with publishing raw code.
+First, there is no guarantee that that code \emph{can} actually
+run, for example, due to a typo added after the code was pasted.
+Secondly, there is no way to determine if code that worked yesterday
+works today. These problems can be prevented by, instead of
+pasting raw code, using a website to host that code (see below).
+
+\paragraph{Raw code is already better than no code}
+
+Publishing raw code is already better than not publishing code at all,
+as that raw code can be assumed to have run successfully at least once.
+To reproduce an experiment from raw code can be painful, 
+yet way better than rewriting the same code from scratch.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Publishing hosted code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Code is sometimes published alongside a paper, 
+in the form of a hosted website.
+This way of publishing code gives additional options,
+both for the authors of that code, as well as its users.
+
+\paragraph{Example of hosted code}
+
+There are multiple websites for hosting code, with GitHub,
+and GitLab being the most well-known.
+GitHub is the most commonly used website to host code in general,
+with millions of users.
+GitHub users can create dedicated websites (called 'repositories')
+to upload code. GitHub hosts that uploaded by and provides
+a user interface for every visiter of such a website.
+An example from the author is the website \cite{bbbq_article_issue_157},
+that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}.
+
+\paragraph{Hosted code has important metadata}
+
+Hosted code has many additional features, such as
+a version-controlled history, continuous integration scripts
+and automatically generated metrics.
+
+\paragraph{Hosted code has a commit history}
+
+Hosted code commonly has a commit history. For example, GitHub
+uses a version control system called \verb|git| (hence its name).
+In practice, this means that when a change is made to the code,
+a new version is created. In the case that the change was harmful,
+one can go back to an earlier version and continue again from there.
+
+\paragraph{A commit history helps assess honest reporting}
+
+A commit history helps to assess the honesty 
+in the reporting of the code. 
+Ideally, the full version history of the code 
+has been generated, by hosting the code directly at its inception.
+However, in some repositories, only the final version of the code has been
+posted, which means all historical information is lost.
+From a code history, it can be observed, for example, 
+how much statistical tests have really/actually been done 
+and see if all are reported:
+not reporting a statistical tests is considered a Dubious Research Practice.
+
+\paragraph{Continuous integration helps assess code quality}
+
+\paragraph{Metrics help assess the honest division of labour}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Publishing a virtual environment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Code has a limited lifetime.
-
-\paragraph{Hosted code has important metadata}
-
- * Number of commits
- * Commit history
- * Code coverage
- * Objective indicator of work being done
 

### 2022-05-29 article.tex
Fix refs
Done most

@@
 a user interface for every visiter of such a website.
-An example from the author is the website \cite{bbbq_article_issue_157},
+An example from the author is the website \cite{bbbqarticleissue157},
 that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}.
@@
 and see if all are reported:
-not reporting a statistical tests is considered a Dubious Research Practice.
+not reporting a statistical tests is considered a Dubious Research Practice 
+\richel{REF}, as this allows dishonest reporting of p-value hacking.
 
@@
 
+Some code hosts, among others, GitHub, 
+allows the researcher to start software scripts when uploading a new version.
+To determine the reproducibility of the code, one such script should
+run the code (on trivial data) to verify if the code works at all.
+If the functionality of the code is lost, the researcher
+can immediatly see which small change was the cause of this.
+A more sophisticated check for the code's correctness is
+to run units tests and determine the percentage of code that has been
+tested, which is called the code coverage. The higher the code coverage,
+the \richel{ref}.
+
 \paragraph{Metrics help assess the honest division of labour}
+
+Most code hosts, among others, GitHub and GitLab,
+automatically keep track of certain metrics.
+One such metric is the amount of code each contributor has made.
+In that way, authors involved in the writing of the software
+are honestly acknowledged.
+
+\paragraph{A reader can contact the author}
+
+A feature that gets too few attention is that code hosts
+allow users to post bug reports and/or ask questions.
+This allows other researchers to assess the health
+of the code.
+
+\paragraph{The problems with hosted code}
+
+Albeith hosted code is vastly superior over raw code
+in case of reproduciblity and correctness,
+there are some problems with hosted code as well.
+The most important one is that hosted code needs maintanance:
+the libraries that are used by an experiment
+will change over time, inevitably causing the experiment to fail
+over time. This process is called bit rot and is a fundamental
+problem \richel{ref}. A solution that has more 
+longevity is to publishing the code installed in a virtual environment,
+as described below.
+
+A lesser problem, which is only a problem to some, is, that
+the author of the code can be contacted and that bugs can be reported.
+It can be argued that no energy should be wasted on published code.
+This is however, a rather selfish point: a bug should be reported and
+given proper attention, for the sake of the correctness of the 
+results published.
 
@@
 
-\paragraph{Code is hard to manage}
+The most reproducible way of submitting the code of an experiment,
+is by providing the code and all it needs 
+installed in a virtual container.
+Virtualization allows one to provide a standard environment,
+that can be used on any other computer.
+The best known examples of container software are Docker and Singularity,
+where the latter is the only viable option to be able to run
+an experiment on a high performance computing (HPC) environment,
+such as a computer cluster.
+
+\paragraph{Containers allow for runnable code, for longer periods of time}
 
 Unlike English, code is fragile in standing the test of time.
-Containers can alleviate this.
-Code has a limited lifetime.
+Containers can alleviate this, as these can be created at a point
+in time and then run later. Because the container is stand-alone,
+all libraries the code depends on remain in their current/old state.
+As with hosted code, an experiment may not run with newer versions
+of libraries anymore, but a containerized experiment can run regardless, if
+one refrains from updating.
 
-\paragraph{Runnable code has other important features}
+Ideally, to re-do an experiment, one would need a one-liner to a container,
+for example, \verb|singularity run my_experiment.sif run|.
 
- * Reproduce an experiment
- * Code can teach the user, see babette
- * Build upon an experiment
+\paragraph{Runnable code can precisely reproduce an experiment}
+
+Containers allow a computation experiment to be highly reproducible:
+given the same data, an experiment put into a container will give
+the same results on different platforms, at least in theory.
+In practice, differences may be observed when peripheral factors
+are different, such as the random numbers as generated by an operating
+system, or data that is downloaded from additional online sources.
+
+\paragraph{Runnable code allows others to re-use an experiment}
+
+Containers allow a computation experiment to be highly re-usable,
+as any scientist can work with it and tweak code.
+
+\paragraph{The problems with runnable code}
+
+There are some problems with providing a container with a runnable container.
+First, a container is several gigabytes and need to be stored somewhere
+online. Ideally, this would be at Syslabs, as that one is free, yet
+there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
+be entered, such as the URL of the article, nor the URL of hosted code.
+Also, journals do not request hosted code, nor runnable code.
+
+\paragraph{Runnable code is the pinnacle of reproducible research}
+
+When research truly needs to be reproducible, putting the code 
+of an experiment into a container is a good solution, as it 
+can stand the test of time well. Rewarding this, however, is not
+done yet by journals.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Stuff

### 2022-05-30 article.tex
Add refs

@@
     \hline
+    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
     Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
@@
 is usually regarded as having a lower chance of containing bugs.
+Additionally, following a style guide (e.g. \cite{style_guide}), 
+improves software quality (\cite{fang2001}).
 
@@
 a user interface for every visiter of such a website.
+The use of code hosting websites, such as GitHub,
+accommodates collaboration (\cite{perez2016ten}).
+\begin{sloppypar}\url{https://github.com/richelbilderbeek/babette},\end{sloppypar}
+and improves transparency (\cite{gorgolewski2016practical}).
 An example from the author is the website \cite{bbbqarticleissue157},
@@
 
-A commit history helps to assess the honesty 
-in the reporting of the code. 
+Hosted code improves transparency (\cite{gorgolewski2016practical}).
+The history of changes, called the commit history, 
+helps to assess the honesty in the reporting of the code. 
 Ideally, the full version history of the code 
@@
 can immediatly see which small change was the cause of this.
+A formalized version of this workflow is called continuous integration (CI).
+Using a CI service (such as GitHub Actions), is known to significantly 
+increase the number of bugs exposed (\cite{vasilescu2015}) and increases
+the speed at which new features are added (\cite{vasilescu2015}).
+
 A more sophisticated check for the code's correctness is
 to run units tests and determine the percentage of code that has been
-tested, which is called the code coverage. The higher the code coverage,
-the \richel{ref}.
+tested, which is called the code coverage. Code coverage
+correlates with code quality (\cite{horgan1994,del1995correlation}).
 
@@
 This article and its metadata can be found at 
-\url{https://github.com/richelbilderbeek/chapter_paradata}.
+\begin{sloppypar}\url{https://github.com/richelbilderbeek/chapter_paradata}\end{sloppypar}.
 

### 2022-05-30 scripts/do_it.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+rm article.pdf        &&
+  ./scripts/edit.sh   &&
+  ./scripts/create.sh &&
+  ./scripts/view.sh 
+

### 2022-05-30 article.tex
Fix ref
EOD

@@
 
+\paragraph{Science should be reproducible}
+
+A scientific claim must be reproducible before it is accepted
+by the scientific community.
+Reproducibility (i.e. to reproduce the same results) 
+and replicability (i.e. to reconclude a conclusion)
+are fundamental characteristics of scientific studies (\cite{patil2019visual}).
+For computational science, it is relatively easy to 
+reproduce an experiment, as all it takes is a computer, electricity,
+an optional internet connection, the code and the data.
+In computational science, however, it appears that 
+culture to reproduce results has been lost 
+and to counter this, it has been suggested to make
+reproducible research a minimal requirement for 
+publication (\cite{peng2011reproducible}).
+
 \paragraph{Code should be published}
 
-For science to be reproducible (a fundamental property of the scientific 
-method), it is the code that should be published, instead of the
-English description of what the code does (i.e. the article), if one
-would have to pick one, as it is the code does the actual work.
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{peng2011reproducible_fig_1.png}
+  \caption{
+    Levels of reproducibility, from \cite{peng2011reproducible}
+  }
+  \label{fig:peng2011reproducible}
+\end{figure}
+
+The code of a computational experiment must be published
+for an experiment to reproducible.
+It is however (too) common that the original data
+and/or code are unavailable (\cite{peng2021reproducible}).
+One sad example is described in \cite{haibe2020importance}, where
+an algorithm that detects breast cancer better than a human expert,
+was irreproducible.
+In the case of genetic epidemiology, it is a given that the original
+data cannot be published as this data is sensitive.
+The code of a computational experiment, however, can be
+distributed without such problems.
+It is, however, not code that is published, but an English
+description of what that code does instead.
+From a knowledge management's perspective,
+to conserve an experiment,
+it is the code that must be archived and not the
+English description of what it does:
+it is the code does the actual work.
 Hence, from a knowledge management perspective, emphasis should be
@@
 in an experiment.
-
-To quote a review on reproducibility \cite{peng2021reproducible}, 
-'Reproducibility is typically thwarted by a 
-lack of availability of the original data and computer code'.
-In the case of genetic epidemiology, it is a given that the original
-data cannot be published as this data is sensitive.
 
@@
 
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{baggerly2009deriving_fig_1a_zoom.png}
+  \includegraphics[width=\linewidth]{baggerly2009deriving_fig_1a.png}
+  \caption{
+    A published off-by-one error.
+    Left: a zoom in. 
+    Horizontal axis is the sample name.
+    Vertical axis is UNKNOWN.
+    Color in the cell denotes gene expression.
+    The colors above the cells, with the dendrogram, shows
+    red for non-responders (red) and responsive (blue).
+    Right: visisualization of sample that shows samples are duplicated,
+    as annotated by the pink rectangles. Note that, due to this, some
+    samples are both non-responders and responders.
+    Adapted from \cite{baggerly2009deriving}
+  }
+  \label{fig:baggerly2009deriving}
+\end{figure}
+
 It is the paper that accompanies the code,
@@
 a mismatch between the English paper and the code increases.
+We know that the most common errors are 
+simple (and may legitimately fear that those simple errors are common) 
+(\cite{baggerly2009deriving}),
+and this example is not too far off from the
+examples given in that paper.
+
 But regardless of the size of the code, 
@@
 
-There are multiple websites for hosting code, with GitHub,
-and GitLab being the most well-known.
-GitHub is the most commonly used website to host code in general,
+There are multiple websites for hosting code, with BitBucket, GitHub,
+GitLab and SourceForge being the most well-known.
+Of these, GitHub is the most popular website to host code in general,
 with millions of users.
@@
 
+\paragraph{Code coverage helps assess code quality}
+
 A more sophisticated check for the code's correctness is
@@
 correlates with code quality (\cite{horgan1994,del1995correlation}). 
+The non-profit organisation rOpenSci, which peer-reviews code,
+has made it a prerequisite to have a code coverage of 100\%.
 

### 2022-05-31 article.tex
Use correct picture

@@
   \centering
-  \includegraphics[width=\linewidth]{DNA_to_protein_or_ncRNA.png}
+  \includegraphics[width=\linewidth]{DNA_to_protein.png}
   \caption{
@@
   \centering
-  \includegraphics[width=\linewidth]{baggerly2009deriving_fig_1a_zoom.png}
   \includegraphics[width=\linewidth]{baggerly2009deriving_fig_1a.png}
   \caption{
-    A published off-by-one error.
-    Left: a zoom in. 
-    Horizontal axis is the sample name.
-    Vertical axis is UNKNOWN.
-    Color in the cell denotes gene expression.
+    Gene expression result due to an off-by-one error.
+    The cells in the main table show how much mRNA is produced per cell 
+    line (the columns) for different proteins (the rows).
     The colors above the cells, with the dendrogram, shows
-    red for non-responders (red) and responsive (blue).
-    Right: visisualization of sample that shows samples are duplicated,
-    as annotated by the pink rectangles. Note that, due to this, some
-    samples are both non-responders and responders.
-    Adapted from \cite{baggerly2009deriving}
+    red for cells that do not respond (red) and do respond (blue) to an 
+    antibiotic.
+    The purple rectangles show samples that are duplicated.
+    Note that, due to this, some
+    samples that behave identically are both non-responders and responders.
+    From \cite{baggerly2009deriving}
   }

### 2022-05-31 article.tex
Badges are useful

@@
 code that has generated the results.
-The chapter concludes by some rules how to allow code to serve as paradata,
+The chapter concludes by some rules how to better code to serve as paradata,
 and hence allowing computational research to be truly reproducible.
@@
 {\bf Keywords:} paradata, reproducible research, code, software,
-FAIR data, computational research, Open Science, DevOps, best practices,
+FAIR data, computational research, Open Science, best practices,
 genetic epidemiology
@@
 \begin{table}[h]
-  \begin{tabular}{lll}
+  \begin{tabular}{lp{5cm}p{5cm}}
     Term      & Definition                                            & Example                                                         \\
@@
 the speed at which new features are added (\cite{vasilescu2015}).
+Using CI opens up the possibility to informally annotate features of
+a repository. For example, Figure \ref{fig:build_badge} shows a 
+build badge that indicates that a LaTeX document (i.e. this paper)
+could be built.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{build_badge.png}
+  \caption{
+    A build badge for the GitHub repository of this article.
+    The description indicates which process it is that the badge
+    signals success (as in this case) or failure of. 
+  }
+  \label{fig:build_badge}
+\end{figure}
 
@@
 has made it a prerequisite to have a code coverage of 100\%.
+Also here, CI opens up the possibility to informally annotate the
+code coverage. For example, Figure \ref{fig:badge_codecov} shows a 
+build badge that indicates that certain R package's code 
+has a 100\% code coverage.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[]{badge_codecov.png}
+  \caption{
+    A build badge for the code coverage of an R 
+    package (\url{https://github.com/ropensci/beautier}).
+    The umbrella signals that the code coverage is hosted by
+    CodeCov (\url{https://codecov.io/}). 
+    The text indicates which process it is that the badge
+    signals, which in this case, is that the code is fully covered
+    by tests.
+  }
+  \label{fig:badge_codecov}
+\end{figure}
 
@@
 
+\paragraph{Suggestions for knowledge management}
+
+Here are multiple suggestions to make hosted code FAIR.
+To make the hosted code findable, commonly a paper refers to its URL.
+At the homepage of hosted code, there is no standardized metadata, to,
+for example, link to the paper.
+Using a popular code host ensures the code is easily accessible.
+There is some interoperabily supplied by a version control manager,
+such as in obtaining the history of the code.
+The reusability of code is reasonably easy to assess, by using build badges.
+However, there is no guarantee that a build badge that claimes success,
+actually is an honest signal (i.e. that code may not actually build),
+nor are there standards for what 'success' entails.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 installed in a virtual container.
+This is close to the golden standard suggested by 
+\cite{peng2011reproducible} (see also Figure \ref{fig:peng2011reproducible}),
+however, that was before virtualization existed.
 Virtualization allows one to provide a standard environment,
@@
 Also, journals do not request hosted code, nor runnable code.
+
+\paragraph{Suggestions for knowledge management}
+
+Here are multiple suggestions to make containers with runnable code FAIR.
+To make such containers findable, the most common 
+environments (Syslabs and DockerHub) to upload containers must be indexed.
+Uploading a container to those environments is, however, limited by
+a restricted amount of free storage, with the possible consequence that
+those containers are uploaded somewhere where they escape proper indexing.
+Having containers uploaded at the existing websites ensures they are
+accessible, interoperable and reusable.
+One could argue however, that the scientific community need its own,
+similar website to upload containers, as the metadata for these
+sites are limited to free-text comments and tags (DockerHub)
+or not metadata at all (Syslabs).
 
@@
 
-\paragraph{Definition}
+\paragraph{Definition of paradata}
 
@@
 the possibily to receive questions regarding that code.
-For a publisher
-For a knowledge archivist,
+Note, however, that not publishing code will always thwart
+the reproduction of incorrect code, at the cost of a scientific
+career \cite{baggerly2009deriving}.
 

### 2022-05-31 article.tex
Better table with recommendations

@@
 
+\begin{table}[h]
+  \begin{tabular}{lp{9cm}}
+    Type of code  & Recommendation \\
+    \hline
+    No code       & Disallow papers that do not supply code publicly \\
+    Raw code      & Don't publish raw code as text, instead, host it on a code hosting website \\
+    Hosted code   & Add CI. Use standarized badges for build status and code coverage. 
+                    Provide for a standarized way to link between paper and code host \\
+    Runnable code & Provide website to upload containers with proper annotation.
+                    Provide for a standarized way to link between container, paper and code repository
+  \end{tabular}
+  \caption{Recommendations described in this paper}
+  \label{tab:recommendations}
+\end{table}
+
 \paragraph{Final word}

### 2022-05-31 article.tex
Add examples, improve table

--- 2022-05-31 article.tex
@@
 An example from the author is the website \cite{bbbqarticleissue157},
-that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}.
+that hosts part of the code for the paper \cite{bilderbeek2022transmembrane},
+as shown in Figure \ref{fig:bbbqarticleissue157}.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{bbbq_article_issue_157.png}
+  \caption{
+    A typical GitHub repository (\cite{bbbqarticleissue157}), 
+    hosting the code for a
+    computational experiment, 
+    as published with \cite{bilderbeek2022transmembrane}.
+    Note that this repository has 72 commits, where a commit is a change
+    to the code.
+  }
+  \label{fig:bbbqarticleissue157}
+\end{figure}
 
@@
 In that way, authors involved in the writing of the software
-are honestly acknowledged.
+are honestly acknowledged. Figure \ref{fig:daisie_contributors}
+shows the number of commits the authors of DAISIE (\cite{etienne2020daisie})
+have made to its code.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{daisie_contributors.png}
+  \caption{
+    An overview of the commits that multiple contributors have
+    made, in this case, to the code of DAISIE (\cite{etienne2020daisie})
+  }
+  \label{fig:daisie_contributors}
+\end{figure}
 
@@
 
+\begin{table}[h]
+  \begin{tabular}{|l|p{9cm}|}
+    \hline
+    \textbf{Type of code} & \textbf{Recommendation} \\
+    \hline
+    No code       & Disallow papers that do not supply code publicly \\
+    \hline
+    Raw code      & Don't publish raw code as text, instead, host it on a code hosting website \\
+    \hline
+    Hosted code   & Add CI. Use standarized badges for build status and code coverage. 
+                    Provide for a standarized way to link between paper and code host \\
+    \hline
+    Runnable code & Provide website to upload containers with proper annotation.
+                    Provide for a standarized way to link between container, paper and code repository \\
+    \hline
+  \end{tabular}
+  \caption{Recommendations described in this paper}
+  \label{tab:recommendations}
+\end{table}
+
 \paragraph{Final word}

### 2022-05-31 article.tex
Pretty table

@@
 
+\rowcolors{2}{gray!10}{gray!40}
 \begin{table}[h]
-  \begin{tabular}{|l|p{9cm}|}
+  \begin{tabular}{|p{2cm}|l|}
     \hline
@@
     \hline
-    Raw code      & Don't publish raw code as text, instead, host it on a code hosting website \\
+    Raw code      & Don't publish raw code as text, host it on a code hosting website \\
     \hline
-    Hosted code   & Add CI. Use standarized badges for build status and code coverage. 
-                    Provide for a standarized way to link between paper and code host \\
+    Hosted code   & Use a standarized badge for build status \\
+                  & Use a standarized badge for code coverage \\
+                  & Standardize to link between paper and code host \\
     \hline
-    Runnable code & Provide website to upload containers with proper annotation.
-                    Provide for a standarized way to link between container, paper and code repository \\
+    Runnable code & Allow to upload containers without limits \\
+                  & Website to host containers must allow for annotation \\
+                  & Use a standarized badge for the container to be functioning  \\
+                  & Standarize to link between container, paper and code repository \\
     \hline

### 2022-05-31 article.tex
Add LDPred2 Issues

@@
 
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{issue_ldpred2.png}
+  \caption{
+    An overview of the issues for the code for
+    LDpred2 (\cite{prive2020ldpred2}). Of the 310 issues, 300 have been closed.
+    The first four issues have been replied to, as indicated by the
+    talk balloon. The fifth issues has a label with the text 'feature request'
+    to signal the issue type.
+  }
+  \label{fig:issue_ldpred2}
+\end{figure}
+
 \paragraph{The problems with hosted code}
@@
 
+\paragraph{Still need the same data to do a reproducible experiment}
+
 \paragraph{Recommendations}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
+\subsection{Funding}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This chapter has not been supported by funding.

### 2022-05-31 article.tex
Remove paragraphs visible in text

@@
 
-\paragraph{Universities produce knowledge}
+% \paragraph{Universities produce knowledge}
 
@@
 
-\paragraph{A paper contains data and should be FAIR data}
+% \paragraph{A paper contains data and should be FAIR data}
 
@@
 
-\paragraph{A paper is data and should be FAIR data}
+% \paragraph{A paper is data and should be FAIR data}
 
@@
 
-\paragraph{For some fields, the experiment is actually run by code}
+% \paragraph{For some fields, the experiment is actually run by code}
 
@@
 
-\paragraph{The experiments within genetic epidemology works are done by code}
+% \paragraph{The experiments within genetic epidemology works are done by code}
 
@@
 
-\paragraph{The code is the paradata for the results of an experiment}
+% \paragraph{The code is the paradata for the results of an experiment}
 
@@
 
-\paragraph{Science should be reproducible}
+% \paragraph{Science should be reproducible}
 
@@
 
-\paragraph{Code should be published}
+% \paragraph{Code should be published}
 
@@
 
-\paragraph{Example of code being the ground truth}
+% \paragraph{Example of code being the ground truth}
 
@@
 
-\paragraph{Raw code is important}
+% \paragraph{Raw code is important}
 
@@
 
-\paragraph{Raw code is an indicator of quality}
+% \paragraph{Raw code is an indicator of quality}
 
@@
 
-\paragraph{The problems with raw code}
+% \paragraph{The problems with raw code}
 
@@
 
-\paragraph{Raw code is already better than no code}
+% \paragraph{Raw code is already better than no code}
 
@@
 
-\paragraph{Example of hosted code}
+% \paragraph{Example of hosted code}
 
@@
 
-\paragraph{Hosted code has important metadata}
+% \paragraph{Hosted code has important metadata}
 
@@
 
-\paragraph{Hosted code has a commit history}
+% \paragraph{Hosted code has a commit history}
 
@@
 
-\paragraph{A commit history helps assess honest reporting}
+% \paragraph{A commit history helps assess honest reporting}
 
@@
 
-\paragraph{Continuous integration helps assess code quality}
+% \paragraph{Continuous integration helps assess code quality}
 
@@
 
-\paragraph{Code coverage helps assess code quality}
+% \paragraph{Code coverage helps assess code quality}
 
@@
 
-\paragraph{Metrics help assess the honest division of labour}
+% \paragraph{Metrics help assess the honest division of labour}
 
@@
 
-\paragraph{A reader can contact the author}
+% \paragraph{A reader can contact the author}
 
@@
 
-\paragraph{The problems with hosted code}
+% \paragraph{The problems with hosted code}
 
@@
 
-\paragraph{Suggestions for knowledge management}
+% \paragraph{Suggestions for knowledge management}
 
@@
 
-\paragraph{Containers allow for runnable code, for longer periods of time}
+% \paragraph{Containers allow for runnable code, for longer periods of time}
 
@@
 
-\paragraph{Runnable code can precisely reproduce an experiment}
+% \paragraph{Runnable code can precisely reproduce an experiment}
 
@@
 
-\paragraph{Runnable code allows others to re-use an experiment}
+% \paragraph{Runnable code allows others to re-use an experiment}
 
@@
 
-\paragraph{The problems with runnable code}
+% \paragraph{The problems with runnable code}
 
@@
 
-\paragraph{Runnable code is the pinnacle of reproducible research}
+% \paragraph{Runnable code is the pinnacle of reproducible research}
 
@@
 
-\paragraph{Definition of paradata}
+% \paragraph{Definition of paradata}
 
@@
 
-\paragraph{The drawbacks of publishing code}
+% \paragraph{The drawbacks of publishing code}
 
@@
 
-\paragraph{The drawbacks of publishing version-controlled code}
+% \paragraph{The drawbacks of publishing version-controlled code}
 
@@
 
-\paragraph{The drawbacks of publishing a running version of code}
+% \paragraph{The drawbacks of publishing a running version of code}
 
@@
 
-\paragraph{Still need the same data to do a reproducible experiment}
+% \paragraph{Still need the same data to do a reproducible experiment}
 
-\paragraph{Recommendations}
+% \paragraph{Recommendations}
 
@@
 
-\paragraph{Final word}
+% \paragraph{Final word}
 

### 2022-06-08 article.tex
Add suggestions, thanks  Wout Dillen

@@
 Findable (e.g. found by web),
-Accessible (e.g. DataDryad and or Zenoda), 
-Interoperable (e.g. comma-separated files) 
+Accessible (e.g. DataDryad and or Zenodo), 
+Interoperable (e.g. plain-text file format, such as comma-separated files) 
 and Reusable (e.g. work on all operating systems)
@@
 i.e.
-Findable (e.g. Google Scholar)
-Accessible (e.g. view it online), 
-Interoperable (e.g. written in English)
-and Reusable (e.g. it is valid to cite and build upon earlier academic papers).
+Findable (e.g. having a DOI)
+Accessible (e.g. open access publications, i.e. published in a public access
+repository), 
+Interoperable (e.g. written in English (i.e. the current lingua franca of sciece))
+and Reusable (e.g. it is valid to cite and build upon earlier academic papers, 
+it is peer-reviewed (i.e. satifies a minimal requirement of quality)).
 This paper itself will later specify the standpoint that a paper

### 2022-06-09 article.tex
Update

@@
 
-Here we define paradata as the data that describes the generation of data.
+Here we define paradata as the data that describes the process of generating data.
 In genetic epidemiology, the data generated is mostly the results 
@@
 as done by computer code.
-In such context, paradata is usually the scientific paper that
-describes what the computer code does.
+In such context, one could argue incorrectly that
+it is the scientific paper describes what the computer code does.
 However, this has the unrealtistic 
@@
     Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
-    Paradata  & Data that describes the generation of data            & The code to conclude that a SNP has a significant association   \\
+    Paradata  & Data that describes the process of generating data    & The code to conclude that a SNP has a significant association   \\
     Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
+    Primary data & Data that is worked with in a research             & The genotype of individuals   \\
     Metadata  & Data that provides information about other data       & The article that describes an experiment                        \\
+    Raw data  & Data directly from the real world, that is not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
     Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
@@
 
-% \paragraph{Universities produce knowledge}
+% \paragraph{Simple example}
 
-The currency that universities generate is knowledge, which in turn is distributed
-in mostly lectures and academic papers.
-Academic papers contain knowledge, either as the
-report of measurements or the outcome of an experiment.
-These reports are described in English, the current lingua franca of science.
+Two different researchers write two different papers,
+that are both accepted. One researcher does not supply the
+computer code that was used to generate the results.
+The experiment of the other researcher can be re-done easily
+and you can verify that, indeed, the results are the same
+as in the paper.
+Do you trust the conclusion of their papers equally?
 
-% \paragraph{A paper contains data and should be FAIR data}
+% \paragraph{Definition of paradata}
 
-A paper contains and/or describes data, where we define
-data as 'individual facts, statistics, or items of information, often 
-numeric' \cite{glossary2003oecd}.
-To ensure unhampered knowledge growth, this data should [TODO: Why? Reference!] 
-follow the FAIR 
-principles.
-The FAIR principles \cite{wilkinson2016fair} are that data should be 
-[TODO: there are probably better examples]
-Findable (e.g. found by web),
-Accessible (e.g. DataDryad and or Zenodo), 
-Interoperable (e.g. plain-text file format, such as comma-separated files) 
-and Reusable (e.g. work on all operating systems)
+In this paper, paradata is defined as information about how data was generated,
+as shown schematically in the left panel of Figure \ref{fig:figure_1}.
+Paradata does not have a clear definition yet, with common elements
+such as that paradata is usually not the primary data, that it describes
+a process, that it describes the way of collecting primary data, or that it
+describes the way the primary data is generating.
 
-% \paragraph{A paper is data and should be FAIR data}
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{figure_1.png}
+  \caption{
+    Left: general relations between data, paradata and metadata.
+    Middle and right: examples of these types of data
+  }
+  \label{fig:figure_1}
+\end{figure}
 
-A paper in itself is data, as it is a collection of text and
-figures that we itself can do measurements upon.
-Also here, a paper should be FAIR, 
-[TODO: there are probably better examples]
-i.e.
-Findable (e.g. having a DOI)
-Accessible (e.g. open access publications, i.e. published in a public access
-repository), 
-Interoperable (e.g. written in English (i.e. the current lingua franca of sciece))
-and Reusable (e.g. it is valid to cite and build upon earlier academic papers, 
-it is peer-reviewed (i.e. satifies a minimal requirement of quality)).
-This paper itself will later specify the standpoint that a paper
-is metadata (i.e. data about data), instead of paradata (i.e.
-data that describes the collection of data)
+% \paragraph{Simple example of paradata}
+
+A simple example would be the data at eBird (\url{https://ebird.org/}),
+where people can submit wher a bird of a certain species was spotted.
+The name of the person that spotted that bird is paradata, 
+as it is that person that generated the data and determined of
+which species the bird was, 
+as shown schematically in the middle panel of Figure \ref{fig:figure_1}.
+
+% \paragraph{Relevance of paradata}
+
+Paradata is relevant, as it allows primary data to be re-used.
+In the eBird example, it helps assess the trustworthiness of the 
+data, as beginner birdwatchers make more mistakes in identifying
+the correct bird species, as compared to experts.
+By using the name and/or experience of the birder as a covariate,
+we can re-use the dataset in other contexts as well.
+
+% \paragraph{This paper is about computer code}
+
+This paper is about paradata in the context of a computational
+experiment in general and using an example from genetic epidemiology.
+A computational experiment is run by computer code
+and usually collects primary data first, then runs an analysis,
+which results in the generation of results.
+This paper takes the standpoint that it is the computer code 
+that describes this generation of results, hence computer code is paradata,
+as shown schematically in the right panel of Figure \ref{fig:figure_1}.
+Reading the computer code allows one to assess the trustworthiness 
+of the experiment, similar to knowing the name or expertise of a bird spotter.
+Having the code allows one to re-use it, in the same or other contexts,
+providing for reproducible research.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Genetic epidemiology}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{Introduction}
+
+This paper uses genetic epidemiology as a specific example,
+to illustrate that the computer code is paradata,
+that this paradata is relevant, as it helps 
+to assess the correctness of the results,
+where the correctness of the results is relevant for healthcare.
 
@@
 
-This paper focusses on research fields that uses computation
-to do experiments and uses genetic epidemiology as an example.
 Genetic epidemiology is a field within biology that, among 
 others, measures the spread (hence 'epidemiology') 
-of heritable (i.e. genetic) traits, as well as the relation
-between having a certain genetic makeup and a certain trait,
+of heritable (i.e. genetic) traits (i.e. a genotype), 
+as well as the relation between having a certain genetic makeup 
+and a certain trait (i.e. a phenotype),
 where the trait can be any human property, such as weight, height,
 the amount of metabolites and having a disease yes/no.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Example study}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Genetic epidemiologists look for associations between genotype
+and phenotype. These associations are subsequently used to 
+create so-called gene panels, where the gene causing 
+an association is measured specifically, to, among others,
+detect people at risk for a phenotype.
 
@@
 
-As an example, we'll use a pseudorandomly selected paper
-from \cite{ahsan2017relative}. The data used by that paper is
+The example study followed is a pseudorandomly selected paper
+from \cite{ahsan2017relative}. The primary data used by that paper is
 from a population study called the Northern Swedish Population
-Health Study (NSPHS) that started in 2010 \cite{igl2010northern}. 
+Health Study (NSPHS) that started in 2010 (\cite{igl2010northern}). 
 The approximately 1000 participants were initially mostly surveyed
-about lifestyle \cite{igl2010northern} and follow-up studies
+about lifestyle (\cite{igl2010northern}) and follow-up studies
 provided the type of data relevant for this paper, 
-which are (1) the genetic data \cite{johansson2013identification},
-(2) the concentration of certain proteins in the 
-blood \cite{enroth2014strong,enroth2015effect}.
+which are (1) the genotypes (\cite{johansson2013identification}),
+(2) the phenotypes, which are protein concentrations, as measured
+from blood (\cite{enroth2014strong,enroth2015effect}).
 
@@
 
-The first type of data, the (final form of the) genetic data,
-consists out of single nucleotide polymorphisms (SNPs).
-SNPs consist out of a name, a position and a nucleotide. 
+The first type of primary data, the genotypes, 
+consists out of single nucleotide polymorphisms (SNPs, pronounced as 'snips').
+A SNP has a name and a location on the DNA, at which there is a certain nucleotide.
 DNA (organized into chromosomes and present in every (nucleated) cell, 
-see figure \ref{fig:eukakyote_dna}) consists out of many nucleotides, of which there are four types,
+see figure \ref{fig:eukakyote_dna}) 
+consists out of many nucleotides, of which there are four types,
 called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
@@
 One SNP example is \verb|rs12133641|, which is a SNP located at position 
-154,428,283,
-where some people have a certain nucleotide. In this case, 
-67 percent of the people within this study have an A,
+154,428,283, where 67 percent of the people within this study have an A,
 and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).
@@
 
-The second type of data are concentrations of certain proteins in the 
-blood. DNA contains the code for building proteins (see Figure 
+The second type of data, the phenotypes, 
+are concentrations of certain proteins as obtained from blood. 
+DNA contains the code for building proteins (see Figure 
 \ref{fig:dna_to_protein}), 
@@
 their presence can be used to assess the health of an individual.
-IL6RA is a protein [TODO: more info]
+IL6RA is one such protein.
 
@@
 correlations between genetic data and biological traits.
-For example, figure \ref{fig:ahsan2017relative_table_2_sub} (from \cite{ahsan2017relative}) shows that
-SNP \verb|rs12133641| is highly correlated (p value is 3.0 * 10e-73, 
-961 individuals) with protein IL6RA. What this results does
+For example, figure \ref{fig:ahsan2017relative_table_2_sub} 
+(from \cite{ahsan2017relative}) shows that
+SNP \verb|rs12133641| is highly correlated (p-value is $3.0^{-73}$, 
+for $n$ = 961 individuals) with protein IL6RA. What this results does
 not yet teach us, is how this correlation works, yet
@@
 the R squared value) is rarely 100 percent, which means that a trait (in
-this case, IL6RA concentrations) cannot be perfectly explained
+this case, the concentration of IL6RA) cannot be perfectly explained
 from the genotype alone. As we can see in figure \ref{fig:ahsan2017relative_s6}, 
@@
 
+These conclusions may end up in the clinic.
+Imagine that the level of IL6RA is associated with a disease,
+we can create a gene panel that specifically measures
+SNP \verb|rs12133641|. In this simple example, it will be easier
+to measure the level of IL6RA in the blood, than using a gene panel.
+However, there are associations published for many diseases,
+in which multiple SNPs contribute to being more likely 
+to develope a disease in the future. Here, the phenotype (having 
+a disease in the future) is impossible to detect at the present
+and associations found in earlier studies are used to create a gene panel.
+As creating a gene panel is costly, those associations better be correct.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To \emph{obtain} the genetic data, yes, there may be fieldwork and/or lab-work
+To \emph{obtain} the raw genetic data, yes, there may be fieldwork and/or lab-work
 involved. For example, a researcher needs to go into the field to take
@@
 
-Programming code is paradata, as it is data that describes how data is collected,
-which is the definition of paradata [TODO: REF].
+Programming code is paradata, 
+as it is data that describes how data is generated.
 Programming code is data that is usually in the form of text, 
@@
 
-The article is metadata, as it is data that describes other
-data, which is the definition of metadata [TODO: REF]:
+The article is metadata, as it is data that describes other data:
 an article describes the
@@
 This chapter has not been supported by funding.
+
+% \paragraph{A paper contains data and should be FAIR data}
+
+A paper contains and/or describes data, where we define
+data as 'individual facts, statistics, or items of information, often 
+numeric' \cite{glossary2003oecd}.
+To ensure unhampered knowledge growth, this data should [TODO: Why? Reference!] 
+follow the FAIR 
+principles.
+The FAIR principles \cite{wilkinson2016fair} are that data should be 
+[TODO: there are probably better examples]
+Findable (e.g. found by web),
+Accessible (e.g. DataDryad and or Zenodo), 
+Interoperable (e.g. plain-text file format, such as comma-separated files) 
+and Reusable (e.g. work on all operating systems)
+
+% \paragraph{A paper is data and should be FAIR data}
+
+A paper in itself is data, as it is a collection of text and
+figures that we itself can do measurements upon.
+Also here, a paper should be FAIR, 
+[TODO: there are probably better examples]
+i.e.
+Findable (e.g. having a DOI)
+Accessible (e.g. open access publications, i.e. published in a public access
+repository), 
+Interoperable (e.g. written in English (i.e. the current lingua franca of sciece))
+and Reusable (e.g. it is valid to cite and build upon earlier academic papers, 
+it is peer-reviewed (i.e. satifies a minimal requirement of quality)).
+This paper itself will later specify the standpoint that a paper
+is metadata (i.e. data about data), instead of paradata (i.e.
+data that describes the collection of data)

### 2022-08-07 article.tex
Add feedback, thanks Patrick Oliver Schenk
Improved

@@
+%
+% 'Data are ...', i.e. use plural
+%
+
 \begin{abstract}
@@
 it is the scientific paper describes what the computer code does.
-However, this has the unrealtistic 
+However, this has the unrealistic 
 assumption that there is a perfect match between the paper and the code.
@@
 
-In this paper, paradata is defined as information about how data was generated,
+In this paper, the definition of paradata
+is 'information about how data was generated',
 as shown schematically in the left panel of Figure \ref{fig:figure_1}.
-Paradata does not have a clear definition yet, with common elements
-such as that paradata is usually not the primary data, that it describes
-a process, that it describes the way of collecting primary data, or that it
-describes the way the primary data is generating.
+Paradata does not have a clear definition yet (hence this book). 
+Common elements in the definitions of paradata are 
+(1) paradata are usually not the primary data, (2) paradata describe
+a process, (3) paradata describe the way of collecting primary data, or 
+(4) paradata describe the way the primary data is generated.
 
@@
 By using the name and/or experience of the birder as a covariate,
-we can re-use the dataset in other contexts as well.
+we can re-use the dataset in other contexts as well,
+as now this example can be generalized to
+a classification problem of unlabeled
+data for learning agents with different amounts 
+of supervised and unsupervised learning.
 
@@
 This paper is about paradata in the context of a computational
-experiment in general and using an example from genetic epidemiology.
-A computational experiment is run by computer code
-and usually collects primary data first, then runs an analysis,
-which results in the generation of results.
+experiment in general and uses an example from genetic 
+epidemiology in particular.
+We'll postpone the introduction to genetic epidemiology now
+and focus on 'computational experiment' first.
+A computational experiment is an experiment run by computer code.
+In chronologic order, it usually collects primary data first, 
+then runs an analysis, which finally results in the generation of results.
+
 This paper takes the standpoint that it is the computer code 
-that describes this generation of results, hence computer code is paradata,
+that describes this generation of results, 
+hence it is the computer code that is paradata (i.e. not the description of the experiment),
 as shown schematically in the right panel of Figure \ref{fig:figure_1}.
@@
 to assess the correctness of the results,
-where the correctness of the results is relevant for healthcare.
+where the correctness of the results is relevant for 
+science in general and healthcare in particular.
 
@@
 Genetic epidemiology is a field within biology that, among 
-others, measures the spread (hence 'epidemiology') 
-of heritable (i.e. genetic) traits (i.e. a genotype), 
-as well as the relation between having a certain genetic makeup 
-and a certain trait (i.e. a phenotype),
-where the trait can be any human property, such as weight, height,
-the amount of metabolites and having a disease yes/no.
-Genetic epidemiologists look for associations between genotype
-and phenotype. These associations are subsequently used to 
-create so-called gene panels, where the gene causing 
+others, measures the spread of heritable traits,
+as well as how these come to be.
+For example, we know that lactose intolerance is, among others,
+caused by producing too few lactose-degrading enzymes,
+and is most commonly found in south-east asia.
+The trait is caused by the genetic make-up, or genotype, of a person.
+The trait, also called phenotype, in this example is lactose intolerance,
+but any human property, such as weight or height can be studied.
+Genetic epidemiology studies the spread of genotypes and phenotypes,
+as well as the relation between those two.
+When an association between genotype and phenotype is found,
+these associations are subsequently used to 
+create gene panels, where the gene causing 
 an association is measured specifically, to, among others,
 detect people at risk for a phenotype.
+
+The rest of this section describes a genetic epidemiology study 
+in more detail, with special focus on the computational experiment.
 
@@
 which are (1) the genotypes (\cite{johansson2013identification}),
-(2) the phenotypes, which are protein concentrations, as measured
-from blood (\cite{enroth2014strong,enroth2015effect}).
+(2) the phenotypes, which are the concentrations of certain proteins
+in the blood (\cite{enroth2014strong,enroth2015effect}).
 
@@
 see figure \ref{fig:eukakyote_dna}) 
-consists out of many nucleotides, of which there are four types,
+consists out of billions of nucleotides.
+There are four types of nucleotides, 
 called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
@@
 The second type of data, the phenotypes, 
-are concentrations of certain proteins as obtained from blood. 
+are concentrations of proteins in the blood. 
 DNA contains the code for building proteins (see Figure 
@@
 their presence can be used to assess the health of an individual.
-IL6RA is one such protein.
+IL6RA is one such protein and, spoiler alert, its concentration
+is associated with the SNP mentioned earlier.
 
@@
 From Figure \ref{fig:ahsan2017relative_s6} we can conclude that, on average,
-the more guanines are inherited at that SNPs location,
+the more guanines are present at that SNPs location,
 the higher concentration of IL6RA can be found in a human's blood.
@@
 43 percent of the variance can be attributed to an individuals'
-genotype and additional factors, 
+genotype. 
+Additional factors, 
 such as the effect
@@
 lifestyle (e.g. smoking yes/no) or having a disease (e.g. diabetes) 
-are needed to explain the variation between individuals better.
+are needed to explain the additional variation.
 
@@
 These conclusions may end up in the clinic.
-Imagine that the level of IL6RA is associated with a disease,
-we can create a gene panel that specifically measures
-SNP \verb|rs12133641|. In this simple example, it will be easier
+For the sake of having a clear example,
+let's assume that a high level of IL6RA 
+is associated with a disease that develops later in life,
+yet is preventable by lifestyle changes.
+Would this be the case, we can create a gene panel that specifically measures
+SNP \verb|rs12133641|. 
+If the gene panel shows an individual has two guanines, 
+we know that this person is likelier to develop higher levels of
+IL6RA and is likelier to benefit from the lifestyle changes.
+
+In this simple example, it will be easier
 to measure the level of IL6RA in the blood, than using a gene panel.
@@
 
-The experiment described above is run by code. There is no fieldwork,
-nor lab-work involved. It is the researcher that writes the code,
-after which the code does the work. A genetic epidemiologist that
-looks for associations between genotype and phenotype,
-one does not need a lab.
+The experiment described above is run by code. 
+It was code that detected the relationship between the genotype
+(in this case, SNP \verb|rs12133641|) 
+and the phenotype (in this case,the concentration of IL6RA).
+To detect this association, 
+there was no fieldwork, nor lab-work involved.
+For this experiment, 
+it has been the researcher that has written the code,
+after which that code does the actual work. 
+A genetic epidemiologist does not need a lab, but computational power instead.
 
@@
 
-Programming code is paradata, 
+Programming code (from now on: 'code') is paradata, 
 as it is data that describes how data is generated.
-Programming code is data that is usually in the form of text, 
+Code is data that is usually in the form of text, 
 spread over one or more files, that describes the experiment.
-The experiment collects the data.
-The data collected is the data we call the results of an experiment.
-
-The article is metadata, as it is data that describes other data:
+The computational experiment (from now on: 'experiment') collects the data
+and then looks for patterns.
+Those patterns (again, a type of data) are the results of an experiment.
+
+The (scientific) article is metadata, as it is data that describes other data:
 an article describes the
@@
 as it has a loose connection with collecting the data.
-Again, if the code and article of an experiment disagree,
+If the code and article of an experiment disagree,
 it is the code that actually let the data be collected.  
@@
 and/or code are unavailable (\cite{peng2021reproducible}).
-One sad example is described in \cite{haibe2020importance}, where
+One saddening example is described in \cite{haibe2020importance}, where
 an algorithm that detects breast cancer better than a human expert,
@@
 data cannot be published as this data is sensitive.
-The code of a computational experiment, however, can be
+The code of a computational experiment, however, \emph{can} be
 distributed without such problems.
@@
 and it is reasonable to assume that this code, when found in the code
-of an article, is assumed to be correct.
+of an article, is correct.
 
@@
 higher than needed, resulting in less significant findings, which
-results in less conclusions drawn from a paper.
+results in (needlessly) less conclusions drawn from a paper.
 

### 2022-08-08 article.tex
Improve

@@
 
+One pseudorandom example of a mismatch between code and article
+was found by me in \cite{labrecque2019interpretation}.
+After the corresponding author of that paper was contacted,
+the code was updated and confirmed to have the same
+results as published in the paper.
+
 It is the paper that accompanies the code,
@@
 
+As it is the code that is the most important actor in an the experiment,
+it should be treated as equally or more important than the scientific
+paper describing what is does.
+Regardless of this, most academic journals, 
+when concerning computational experiments
+only review the (usually) English description of the code
+and make the review of code optional.
+One notable exception is rOpenSci (\cite{ram2018community}),
+which is an organisation that does peer-review academic code 
+and facilitates the optional publication of a paper.
+Although both code and paper are peer-reviewed,
+most effort is spent on reviewing the code,
+as can be seen in, for example, 
+the \verb|mcbette| R package (\cite{bilderbeek2020mcbette}).
+
 There are multiple ways of publishing code,
-that are increasingly hard to do, yet result
-in more reproducible experiments:
+that require increasingly more knowledge to do and result
+in increasingly more reproducible experiments:
 (1) publishing the code as text only, e.g. by printing
@@
 A feature that deserves more attention
-is that code can actively teach what it does.
-A mechanism to do so, which is present in most programming
-languages, is error handling: errors are the mechanism
-by which a programmer can teach the user what to do/expect.
-One example is when calculating the variance of a distribution.
+is that 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
+Error handling is one of the mechanisms to do so.
+For example, let is when calculating the variance of a distribution.
 To calculate the variance of a distribution, at least two values
@@
 To reproduce an experiment from raw code can be painful, 
-yet way better than rewriting the same code from scratch.
+yet it is usually better than rewriting the same code from scratch.
 
@@
 both for the authors of that code, as well as its users.
+This practice has increased,
+with the abstracts from articles in the academic journal
+'Bioinformatics' referring to such a hosted website
+has increased tenfold, going from below 5 percent in 2009 
+to around 50 percent in 2017 (\cite{russell2018large}).
 
@@
 GitHub users can create dedicated websites (called 'repositories')
-to upload code. GitHub hosts that uploaded by and provides
-a user interface for every visiter of such a website.
-The use of code hosting websites, such as GitHub,
+to upload code. Code hosts, such as GitHub, 
+store that uploaded code and provide ways 
+for visiters to download the code and/or interact with it.
+The use of such code hosting websites
 accommodates collaboration (\cite{perez2016ten}).
-\begin{sloppypar}\url{https://github.com/richelbilderbeek/babette},\end{sloppypar}
 and improves transparency (\cite{gorgolewski2016practical}).
@@
 and automatically generated metrics.
+A code hosting website produces and shares this metadata.
+This metadata can help to determine the
+reproducibility, correctness and quality of code
+of computational experiments.
 
@@
 one can go back to an earlier version and continue again from there.
+It is a general recommendation to put version control
+on all human-produced data (\cite{wilson2014best}).
 
@@
 and see if all are reported:
-not reporting a statistical tests is considered a Dubious Research Practice 
-\richel{REF}, as this allows dishonest reporting of p-value hacking.
+failing to reporting a (usually non-significant) statistical tests is 
+one of the many degrees of freedom 
+in p-value hacking (\cite{wicherts2016degrees}).
 

### 2022-08-09 article.tex
Progress

@@
 The cyclomatic complexity correlates with code complexity,
-where more complex code is likelier to contain bugs.
+where more complex code is likelier to contain or give rise to bugs 
+(\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting})
 
 A feature that deserves more attention
-is that 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
+is that 'code must act as a teacher for future developers' (\cite{sadowski2018modern}).
 Error handling is one of the mechanisms to do so.
-For example, let is when calculating the variance of a distribution.
-To calculate the variance of a distribution, at least two values
-are required. When the variance of only one value is requested,
-an error message can helpfully indicate that one needs at least two values.
+A simple example is when one (e.g. a BSc student) 
+needs to calculate the variance of a distribution.
+Commonly one needs to call a function to do so, 
+e.g. \verb|var| in R, that needs the values of the distribution
+as input.
+However, to calculate the variance of a distribution, at least two values
+are required. 
+When the variance of (zero or) only one value is requested,
+an error message can helpfully indicate that the function 
+needs at least two values.
 
@@
 To reproduce an experiment from raw code can be painful, 
-yet it is usually better than rewriting the same code from scratch.
+yet it is usually better than writing the same code from scratch.
 
@@
 GitLab and SourceForge being the most well-known.
-Of these, GitHub is the most popular website to host code in general,
+GitHub is currenly the most popular website to host code in general,
 with millions of users.
@@
 
-A more sophisticated check for the code's correctness is
-to run units tests and determine the percentage of code that has been
-tested, which is called the code coverage. Code coverage
-correlates with code quality (\cite{horgan1994,del1995correlation}). 
-The non-profit organisation rOpenSci, which peer-reviews code,
+Testing, in general, is an important mechanism to ensure
+the correctness of code (an interesting example is \cite{rahman2020exploratory}
+showing bugs in scientific software on the COVID-19 pandemic. 
+It recommends more testing).
+The percentage of (lines of) code tested is called the code coverage.
+Code coverage correlates with code quality (\cite{horgan1994,del1995correlation}). 
+The non-profit organisation rOpenSci (\cite{ram2018community}),
+which peer-reviews code,
 has made it a prerequisite to have a code coverage of 100\%.
@@
 has a 100\% code coverage.
+Hence, code coverage is important metadata, that can automatically
+be acquired when using code hosting website.
 
@@
 have made to its code.
+The distribution of contributed code per author 
+is useful metadata, that can automatically
+be acquired when using code hosting website.
 
@@
 of the code.
+A healthy project has few open bug reports
+and these bug reports should be closed within reasonable time.
+When bug reports start to accumulate, this
+signals either that the code contained many bugs or that
+the developers lost interest in maintaining that code.
+The amount of bug reports is useful metadata, 
+especially when deciding to use (and depend) on a piece of software.
 
@@
 
-Albeith hosted code is vastly superior over raw code
+Albeit hosted code is vastly superior over raw code
 in case of reproduciblity and correctness,
@@
 over time. This process is called bit rot and is a fundamental
+
+HIERO
+
 problem \richel{ref}. A solution that has more 

### 2022-08-10 article.tex
Improve

@@
 will change over time, inevitably causing the experiment to fail
-over time. This process is called bit rot and is a fundamental
-
-HIERO
-
-problem \richel{ref}. A solution that has more 
-longevity is to publishing the code installed in a virtual environment,
-as described below.
-
-A lesser problem, which is only a problem to some, is, that
-the author of the code can be contacted and that bugs can be reported.
-It can be argued that no energy should be wasted on published code.
-This is however, a rather selfish point: a bug should be reported and
-given proper attention, for the sake of the correctness of the 
-results published.
+over time. 
+The degradation of software is a known feature for nearly 
+four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
+here I will use the more modern (and in this context superior)
+definition of software collapse (\cite{hinsen2019dealing}),
+in which software fails due to dependencies on other 
+software.
+Where this paper recommends using older software foundations 
+with multiple implementations (\cite{hinsen2019dealing}),
+there is a modern solution that potentially has more longevity:
+to publish the code installed in a virtual environment
+(including the dependencies used at that time), as described below.
+
+There is another problem that authors of academic code face,
+which is that they will be contacted with a user sending in a bug report.
+One solution for the author is to ignore such emails
+and it can be argued that no energy should be wasted on published code
+and work on something new instead.
+However, upon the find of a new bug, the question is:
+does all the research that use that software still 
+result in the same conclusions?
+A code hosting website usually has a place to report bugs
+that offer transparency in the handling of bug reports.
 
@@
 To make the hosted code findable, commonly a paper refers to its URL.
-At the homepage of hosted code, there is no standardized metadata, to,
+At the homepage of hosted code (e.g. a GitHub repository), 
+there is no standardized metadata, to,
 for example, link to the paper.
@@
 such as in obtaining the history of the code.
-The reusability of code is reasonably easy to assess, by using build badges.
+The reusability of code is reasonably easy to assess, 
+by using so-called build badges.
 However, there is no guarantee that a build badge that claimes success,
@@
 The most reproducible way of submitting the code of an experiment,
-is by providing the code and all it needs 
-installed in a virtual container.
+is by providing the code with its dependencies installed in a virtual container.
 This is close to the golden standard suggested by 
@@
 in time and then run later. Because the container is stand-alone,
-all libraries the code depends on remain in their current/old state.
+all software that the code depends on remain in their current/old state.
 As with hosted code, an experiment may not run with newer versions
-of libraries anymore, but a containerized experiment can run regardless, if
-one refrains from updating.
+of libraries anymore, but a containerized experiment can run regardless, 
+if one refrains from updating (or only does so when there are there problems with fixing).
 
@@
 First, a container is several gigabytes and need to be stored somewhere
-online. Ideally, this would be at Syslabs, as that one is free, yet
-there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
+online. 
+Ideally, for Singularity, this would be at the main container hosting
+site called Syslabs. Using Syslabs is free upon registration, 
+yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
 be entered, such as the URL of the article, nor the URL of hosted code.
@@
 yet there are methods being devised to run code on sensitive
-data with assured privacy \cite{zhang2016review,azencott2018machine}.
+data with assured privacy (\cite{zhang2016review,azencott2018machine}).
 
@@
 sterile is essential to have the correct results,
-unit tests allow a computational biologist to have the same.
+unit tests allow a computational biologist to achieve the same.
 

### 2022-08-12 article.tex
Progress

@@
 
+Code is paradata, as it is describes how the results are generated.
+The part of a scholarly article that describes the experiment is metadata, 
+as it describes what the code is thought to do.
 The code that does the work in a computational experiment should
-always be published together with a scholarly article.
+always be published together with a scholarly article,
+as, if the paper and code differ, it is the code that generated the results.
+
 For research to be reproducible, one ideally has access to
@@
 to 'it does something').
-In academia, to uncover the truth, code correctness is essential.
-To make a comparison with cell biologists, where working
-sterile is essential to have the correct results,
-unit tests allow a computational biologist to achieve the same.
+In academia, to uncover the truth, code correctness is essential,
+similar to cell biologists working sterile to not contaminate their
+cell cultures.
+Unit tests are a vital practice for a computational biologist.
 

### 2022-08-23 article.tex
Remove figures
Quote Linus

@@
 In such context, one could argue incorrectly that
-it is the scientific paper describes what the computer code does.
+it is the scientific paper describes which computation steps 
+have taken place.
 However, this has the unrealistic 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Definitions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[h]
-  \begin{tabular}{lp{5cm}p{5cm}}
-    Term      & Definition                                            & Example                                                         \\
-    \hline
-    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
-    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
-    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
-    Paradata  & Data that describes the process of generating data    & The code to conclude that a SNP has a significant association   \\
-    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
-    Primary data & Data that is worked with in a research             & The genotype of individuals   \\
-    Metadata  & Data that provides information about other data       & The article that describes an experiment                        \\
-    Raw data  & Data directly from the real world, that is not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
-    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
-  \end{tabular}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\epigraph{
+  Talk is cheap. Show me the code. \\
+  \\
+  Linus Torvalds, 2000-08-25. Message to linux-kernel mailing list
+}
 
@@
 
+\iffalse
 \begin{figure}[!htbp]
@@
 \end{figure}
+\fi
 
@@
 \end{figure}
-
+\fi
+
+\iffalse
 \begin{figure}[!htbp]
@@
 
-% \paragraph{A paper contains data and should be FAIR data}
-
-A paper contains and/or describes data, where we define
-data as 'individual facts, statistics, or items of information, often 
-numeric' \cite{glossary2003oecd}.
-To ensure unhampered knowledge growth, this data should [TODO: Why? Reference!] 
-follow the FAIR 
-principles.
-The FAIR principles \cite{wilkinson2016fair} are that data should be 
-[TODO: there are probably better examples]
-Findable (e.g. found by web),
-Accessible (e.g. DataDryad and or Zenodo), 
-Interoperable (e.g. plain-text file format, such as comma-separated files) 
-and Reusable (e.g. work on all operating systems)
-
-% \paragraph{A paper is data and should be FAIR data}
-
-A paper in itself is data, as it is a collection of text and
-figures that we itself can do measurements upon.
-Also here, a paper should be FAIR, 
-[TODO: there are probably better examples]
-i.e.
-Findable (e.g. having a DOI)
-Accessible (e.g. open access publications, i.e. published in a public access
-repository), 
-Interoperable (e.g. written in English (i.e. the current lingua franca of sciece))
-and Reusable (e.g. it is valid to cite and build upon earlier academic papers, 
-it is peer-reviewed (i.e. satifies a minimal requirement of quality)).
-This paper itself will later specify the standpoint that a paper
-is metadata (i.e. data about data), instead of paradata (i.e.
-data that describes the collection of data)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Definitions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[h]
+  \begin{tabular}{lp{5cm}p{5cm}}
+    Term      & Definition                                            & Example                                                         \\
+    \hline
+    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
+    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
+    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
+    Paradata  & Data that describes the process of generating data    & The code to conclude that a SNP has a significant association   \\
+    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
+    Primary data & Data that is worked with in a research             & The genotype of individuals   \\
+    Metadata  & Data that provides information about other data       & The article that describes an experiment                        \\
+    Raw data  & Data directly from the real world, that is not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
+    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
+  \end{tabular}
+\end{table}

### 2022-08-23 article.tex
Improve quote
Improve

@@
 In such context, one could argue incorrectly that
-it is the scientific paper describes which computation steps 
+it is the -usually English- scientific paper that best describes which computation steps 
 have taken place.
@@
 \epigraph{
-  Talk is cheap. Show me the code. \\
-  \\
-  Linus Torvalds, 2000-08-25. Message to linux-kernel mailing list
+  Talk is cheap. Show me the code.
+}{
+  Linus Torvalds, 2000-08-25
 }
@@
   \caption{
-    Left: general relations between data, paradata and metadata.
-    Middle and right: examples of these types of data
+    Left: general relation between data, paradata and metadata.
+    Middle: the same relations specified for eBird.
+    Right: the same relations specified for genetic epidemiology.
   }
@@
 In the eBird example, it helps assess the trustworthiness of the 
-data, as beginner birdwatchers make more mistakes in identifying
+data, as beginner birdwatchers tend to make more mistakes in identifying
 the correct bird species, as compared to experts.
@@
 data for learning agents with different amounts 
-of supervised and unsupervised learning.
+of experience.
 
@@
 that describes this generation of results, 
-hence it is the computer code that is paradata (i.e. not the description of the experiment),
+hence it is the computer code that is paradata 
+(i.e. not the description of the experiment as done in a natural language),
 as shown schematically in the right panel of Figure \ref{fig:figure_1}.
@@
 of the experiment, similar to knowing the name or expertise of a bird spotter.
-Having the code allows one to re-use it, in the same or other contexts,
-providing for reproducible research.
+Having the code allows one to 
+verify, re-use and extend the experiment, 
+in the same or other contexts, providing for reproducible research.
+Additionnally, this paper suggests some best practices on how to make 
+an experiment FAIR data (findability, accessibility, interoperability, and reusability) \cite{wilkinson2016fair} better.
 
@@
 science in general and healthcare in particular.
+However, any field that uses computation in its experiments
+could be used as an example.
 
@@
 and is most commonly found in south-east asia.
-The trait is caused by the genetic make-up, or genotype, of a person.
+The trait is caused by the genetic make-up, 
+or genotype (see Table \ref{tab:definitions}), of a person.
 The trait, also called phenotype, in this example is lactose intolerance,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Code is more important than the (English) paper}
+\section{Code is more important than the paper}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 an article describes the
-experiment and hence the programming
-code in English. However, it is not the best
+experiment (and hence the programming
+code) in a natural language, which is usually English (at this point in time). 
+However, a natural language is not the best
 candidate to describe how the data is collected,
@@
   \end{tabular}
+  \caption{Terms used in this paper, and their definitions}
+  \label{tab:definitions}
 \end{table}

### 2022-08-23 article.tex
Citations are surrounded by blocky brackets only

@@
 The trait is caused by the genetic make-up, 
-or genotype (see Table \ref{tab:definitions}), of a person.
-The trait, also called phenotype, in this example is lactose intolerance,
-but any human property, such as weight or height can be studied.
+or genotype 
+(see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
+of a person.
+The trait, also called phenotype (again, see Table \ref{tab:definitions}), 
+in this example is lactose intolerance,
+yet any human property, such as weight or height can be studied.
 Genetic epidemiology studies the spread of genotypes and phenotypes,
@@
 an association is measured specifically, to, among others,
-detect people at risk for a phenotype.
+detect people at risk for the associated phenotype.
 
@@
 from a population study called the Northern Swedish Population
-Health Study (NSPHS) that started in 2010 (\cite{igl2010northern}). 
+Health Study (NSPHS) that started in 2010 \cite{igl2010northern}. 
 The approximately 1000 participants were initially mostly surveyed
-about lifestyle (\cite{igl2010northern}) and follow-up studies
+about lifestyle \cite{igl2010northern} and follow-up studies
 provided the type of data relevant for this paper, 
-which are (1) the genotypes (\cite{johansson2013identification}),
+which are (1) the genotypes \cite{johansson2013identification},
 (2) the phenotypes, which are the concentrations of certain proteins
-in the blood (\cite{enroth2014strong,enroth2015effect}).
+in the blood \cite{enroth2014strong,enroth2015effect}.
 
@@
 and replicability (i.e. to reconclude a conclusion)
-are fundamental characteristics of scientific studies (\cite{patil2019visual}).
+are fundamental characteristics of scientific studies \cite{patil2019visual}.
 For computational science, it is relatively easy to 
@@
 reproducible research a minimal requirement for 
-publication (\cite{peng2011reproducible}).
+publication \cite{peng2011reproducible}.
 
@@
 It is however (too) common that the original data
-and/or code are unavailable (\cite{peng2021reproducible}).
+and/or code are unavailable \cite{peng2021reproducible}.
 One saddening example is described in \cite{haibe2020importance}, where
@@
 simple (and may legitimately fear that those simple errors are common) 
-(\cite{baggerly2009deriving}),
+\cite{baggerly2009deriving},
 and this example is not too far off from the
@@
 and make the review of code optional.
-One notable exception is rOpenSci (\cite{ram2018community}),
+One notable exception is rOpenSci \cite{ram2018community},
 which is an organisation that does peer-review academic code 
@@
 as can be seen in, for example, 
-the \verb|mcbette| R package (\cite{bilderbeek2020mcbette}).
+the \verb|mcbette| R package \cite{bilderbeek2020mcbette}.
 
@@
 Additionally, following a style guide (e.g. \cite{style_guide}), 
-improves software quality (\cite{fang2001}).
+improves software quality \cite{fang2001}.
 
@@
 where more complex code is likelier to contain or give rise to bugs 
-(\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting})
+\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}
 
 A feature that deserves more attention
-is that 'code must act as a teacher for future developers' (\cite{sadowski2018modern}).
+is that 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
 Error handling is one of the mechanisms to do so.
@@
 has increased tenfold, going from below 5 percent in 2009 
-to around 50 percent in 2017 (\cite{russell2018large}).
+to around 50 percent in 2017 \cite{russell2018large}.
 
@@
 The use of such code hosting websites
-accommodates collaboration (\cite{perez2016ten}).
-and improves transparency (\cite{gorgolewski2016practical}).
+accommodates collaboration \cite{perez2016ten}.
+and improves transparency \cite{gorgolewski2016practical}.
 An example from the author is the website \cite{bbbqarticleissue157},
@@
   \caption{
-    A typical GitHub repository (\cite{bbbqarticleissue157}), 
+    A typical GitHub repository \cite{bbbqarticleissue157}, 
     hosting the code for a
@@
 It is a general recommendation to put version control
-on all human-produced data (\cite{wilson2014best}).
+on all human-produced data \cite{wilson2014best}.
 
@@
 
-Hosted code improves transparency (\cite{gorgolewski2016practical}).
+Hosted code improves transparency \cite{gorgolewski2016practical}.
 The history of changes, called the commit history, 
@@
 one of the many degrees of freedom 
-in p-value hacking (\cite{wicherts2016degrees}).
+in p-value hacking \cite{wicherts2016degrees}.
 
@@
 Using a CI service (such as GitHub Actions), is known to significantly 
-increase the number of bugs exposed (\cite{vasilescu2015}) and increases
-the speed at which new features are added (\cite{vasilescu2015}).
+increase the number of bugs exposed \cite{vasilescu2015} and increases
+the speed at which new features are added \cite{vasilescu2015}.
 Using CI opens up the possibility to informally annotate features of
@@
 The percentage of (lines of) code tested is called the code coverage.
-Code coverage correlates with code quality (\cite{horgan1994,del1995correlation}). 
-The non-profit organisation rOpenSci (\cite{ram2018community}),
+Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
+The non-profit organisation rOpenSci \cite{ram2018community},
 which peer-reviews code,
@@
 are honestly acknowledged. Figure \ref{fig:daisie_contributors}
-shows the number of commits the authors of DAISIE (\cite{etienne2020daisie})
+shows the number of commits the authors of DAISIE \cite{etienne2020daisie}
 have made to its code.
@@
     An overview of the commits that multiple contributors have
-    made, in this case, to the code of DAISIE (\cite{etienne2020daisie})
+    made, in this case, to the code of DAISIE \cite{etienne2020daisie}
   }
@@
     An overview of the issues for the code for
-    LDpred2 (\cite{prive2020ldpred2}). Of the 310 issues, 300 have been closed.
+    LDpred2 \cite{prive2020ldpred2}. Of the 310 issues, 300 have been closed.
     The first four issues have been replied to, as indicated by the
@@
 here I will use the more modern (and in this context superior)
-definition of software collapse (\cite{hinsen2019dealing}),
+definition of software collapse \cite{hinsen2019dealing},
 in which software fails due to dependencies on other 
@@
 Where this paper recommends using older software foundations 
-with multiple implementations (\cite{hinsen2019dealing}),
+with multiple implementations \cite{hinsen2019dealing},
 there is a modern solution that potentially has more longevity:
@@
 yet there are methods being devised to run code on sensitive
-data with assured privacy (\cite{zhang2016review,azencott2018machine}).
+data with assured privacy \cite{zhang2016review,azencott2018machine}.
 

### 2022-08-23 article.tex
Use a semicolon instead of a plus before ggplot2::ggsave, as the plus now gives a rightful error

@@
 which are (1) the genotypes \cite{johansson2013identification},
-(2) the phenotypes, which are the concentrations of certain proteins
-in the blood \cite{enroth2014strong,enroth2015effect}.
+(2) the phenotypes, in this case, concentrations of certain proteins in the blood \cite{enroth2014strong,enroth2015effect}.
 
@@
 DNA (organized into chromosomes and present in every (nucleated) cell, 
-see figure \ref{fig:eukakyote_dna}) 
+\iffalse
+(see figure \ref{fig:eukakyote_dna}))
+\fi
 consists out of billions of nucleotides.
@@
 are concentrations of proteins in the blood. 
-DNA contains the code for building proteins (see Figure 
-\ref{fig:dna_to_protein}), 
-as well as the rate
+DNA contains the code for building proteins
+\iffalse
+(see Figure \ref{fig:dna_to_protein})
+\fi
+, as well as the rate
 at which a protein is created. Some proteins end up in the blood and
@@
 correlations between genetic data and biological traits.
-For example, figure \ref{fig:ahsan2017relative_table_2_sub} 
-(from \cite{ahsan2017relative}) shows that
-SNP \verb|rs12133641| is highly correlated (p-value is $3.0^{-73}$, 
-for $n$ = 961 individuals) with protein IL6RA. What this results does
-not yet teach us, is how this correlation works, yet
-figure \ref{fig:ahsan2017relative_s6} shows us the direction of the association:
-the X axis shows the possible genetic make-ups (aka the genotype) of the individuals,
-where 0 denotes the individuals with genotype
-AA (the individual inherited one adenosine 
-from his/here mother and one adenosine from his/her father), 
-1.0 denotes AG (one A is inherited from one parent, 
-where the G is inherited from the other parent) and 2.0 denotes GG.
-From Figure \ref{fig:ahsan2017relative_s6} we can conclude that, on average,
-the more guanines are present at that SNPs location,
-the higher concentration of IL6RA can be found in a human's blood.
-
-The amount of variance that can be explained by an association (i.e.
+For example, \cite{ahsan2017relative} shows that
+SNP \verb|rs12133641| is highly correlated (p-value is $3.0^{-73}$
+\iffalse
+(see figure \ref{fig:ahsan2017relative_table_2_sub})
+\fi
+,
+for $n$ = 961 individuals) with protein IL6RA.
+The direction of the association is also concluded:
+the more guanines (as opposited to adenines) are present at that SNPs location,
+the higher concentration of IL6RA can be found in a human's blood
+\iffalse
+(see figure \ref{fig:ahsan2017relative_table_2_sub})
+\fi
+. The amount of variance that can be explained by an association (i.e.
 the R squared value) is rarely 100 percent, which means that a trait (in
 this case, the concentration of IL6RA) cannot be perfectly explained
-from the genotype alone. As we can see in figure \ref{fig:ahsan2017relative_s6}, 
-43 percent of the variance can be attributed to an individuals'
-genotype. 
-Additional factors, 
+from the genotype (in this case, SNP \verb|rs12133641|) alone. 
+In this example, 43 percent of the variance 
+can be attributed to an individuals' genotype
+\iffalse
+(see figure \ref{fig:ahsan2017relative_table_2_sub})
+\fi
+. Additional factors, 
 such as the effect
@@
 
-These conclusions may end up in the clinic.
-For the sake of having a clear example,
+The conclusions drawn from this paper,
+may end up in the clinic.
+For the sake of having a clear (yet fictuous) example,
 let's assume that a high level of IL6RA 
@@
 
-In this simple example, it will be easier
+From this simple example, it will be easier
 to measure the level of IL6RA in the blood, than using a gene panel.

### 2022-08-24 article.tex
Progress

@@
 From this simple example, it will be easier
-to measure the level of IL6RA in the blood, than using a gene panel.
+to measure the level of IL6RA in the blood, than using a gene panel,
+as blood tests are easier and cheaper.
 However, there are associations published for many diseases,
-in which multiple SNPs contribute to being more likely 
-to develope a disease in the future. Here, the phenotype (having 
-a disease in the future) is impossible to detect at the present
+in which one (e.g. phenylketonuria) or many (e.g. \cite{bruce2009metabolic}) 
+SNPs contribute to being more likely to develope a disease in the future. 
+Here, the phenotype (having a disease in the future) 
+is impossible to detect at the present
 and associations found in earlier studies are used to create a gene panel.
@@
 (in this case, SNP \verb|rs12133641|) 
-and the phenotype (in this case,the concentration of IL6RA).
+and the phenotype (in this case, the concentration of IL6RA).
 To detect this association, 
-there was no fieldwork, nor lab-work involved.
+there was no fieldwork nor lab-work involved.
 For this experiment, 
@@
 
-To \emph{obtain} the raw genetic data, yes, there may be fieldwork and/or lab-work
+To \emph{obtain} the raw genetic data, there may have been fieldwork and/or lab-work
 involved. For example, a researcher needs to go into the field to take
-samples (e.g. blood) of humans or other species. To analyse these samples,
-a researcher needs a lab. Depending on the technique being used on
-how to analyse the samples, there may be a big bioinformatics step
-to aggregate the measurements into useful genetic data. 
+samples (e.g. blood) of humans or other species. 
+To analyse these samples,
+a researcher needs a lab. 
+Depending on the analysis, there may be a big bioinformatics step
+to aggregate the measurements into usable data. 
 Although this paper does not focus on the code run to 

### 2022-08-24 article.tex
Progress

@@
 Although this paper does not focus on the code run to 
-aggragate the raw data into useful genetic data, the same arguments
+aggragate raw and noisy measurements into usable data, the same arguments
 can be made for that code as in the example: that code is 
-the ground truth of how the raw data is collected into useful genetic data.
+the ground truth of how the raw data is converted into useful genetic data.
 
@@
 Code is data that is usually in the form of text, 
-spread over one or more files, that describes the experiment.
+spread over one or more files, that describes an experiment/annalysis.
 The computational experiment (from now on: 'experiment') collects the data
@@
 an article describes the
-experiment (and hence the programming
-code) in a natural language, which is usually English (at this point in time). 
-However, a natural language is not the best
+experiment; an article describes what the code does in a natural 
+language, such as English (which I'll use from now on).
+However, English (or any other natural language) is not the best
 candidate to describe how the data is collected,
-as it has a loose connection with collecting the data.
+as it only has a loose connection with collecting the data.
 If the code and article of an experiment disagree,
@@
 are fundamental characteristics of scientific studies \cite{patil2019visual}.
-For computational science, it is relatively easy to 
+For computational science, theoretically, it should be relatively easy to 
 reproduce an experiment, as all it takes is a computer, electricity,
 an optional internet connection, the code and the data.
-In computational science, however, it appears that 
-culture to reproduce results has been lost 
-and to counter this, it has been suggested to make
-reproducible research a minimal requirement for 
+In practice, however, it appears that 
+the academic culture to reproduce results 
+has been lost over time \cite{peng2011reproducible}.
+That same study recommended, as a solution to counter this, 
+to make reproducible research a minimal requirement for 
 publication \cite{peng2011reproducible}.
@@
 One saddening example is described in \cite{haibe2020importance}, where
-an algorithm that detects breast cancer better than a human expert,
+an algorithm that detects breast cancer from images better than a human expert,
 was irreproducible.
-In the case of genetic epidemiology, it is a given that the original
-data cannot be published as this data is sensitive.
+In the case of genetic epidemiology, 
+the original data usually cannot be published, 
+due to the guaranteed privacy of the people having shared their characteristics.
 The code of a computational experiment, however, \emph{can} be
 distributed without such problems.
-It is, however, not code that is published, but an English
-description of what that code does instead.
+In academic journals, however, its is not code that is published, 
+but an English description of what it does instead.
+
 From a knowledge management's perspective,
+it is the code that must be archived 
 to conserve an experiment,
-it is the code that must be archived and not the
+and not the
 English description of what it does:
 it is the code does the actual work.
-Hence, from a knowledge management perspective, emphasis should be
+Hence, from that perspective, emphasis should be
 put more on the preservation of code, as it is the most important actor
@@
 
-Code should be published, as it holds the
+From a broader perspective, code should be published, as it holds the
 ground truth of an experiment; it does the actual work.
@@
 To illustrate how easy it is to get a mismatch between a paper
-and the code, consider this fictional example
-of a text in a paper:
+and the code, consider this example
+of a text in a fictional paper:
 

### 2022-08-24 article.tex
Progress

@@
 \verb|t.test| is indeed the name of an R function to do a T-test
-and it is reasonable to assume that this code, when found in the code
+and it is reasonable to assume that this code, when glancing over the code
 of an article, is correct.
@@
 higher than needed, resulting in less significant findings, which
-results in (needlessly) less conclusions drawn from a paper.
+results in (needlessly) less conclusions drawn in this fictional paper.
 
@@
 results as published in the paper.
+This sequence of events could only have happened, 
+because the authors chose to publish their code.
 
@@
 But regardless of the size of the code, 
-it is the code that is the ground truth.
+it is the code that is the ground truth
+and should be treated as a first-class citizen
+in a computational experiment.
 
@@
 for example, using a Singularity container.
-Here I will go into details of each of these ways.
+Here I will go into details of each of these ways,
+by describing some advantages, disadvantages, as well as how to
+make the code more FAIR data \cite{wilkinson2016fair}.
 
@@
 correctly and reproducibly, 
-as it has/should been that code that generated the results of an experiment.
-Again: if the paper and code differ, it is the code that did
-the actual work with the actual/true results.
+as it the code that can reproduce the results.
 
@@
 is usually regarded as having a lower chance of containing bugs.
-Additionally, following a style guide (e.g. \cite{style_guide}), 
+Additionally, following a style guide 
+(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
+or PEP 8 \cite{van2001pep} for Python), 
 improves software quality \cite{fang2001}.
@@
 A more interesting metric is the cyclomatic complexity of the code.
-The cyclomatic complexity is defined as the number of ways that
+The cyclomatic complexity is approximately defined 
+as the number of independent paths that
 the code can be executed. 
@@
 has a cyclomatic complexity of 2, as the condition within the \verb|if|
-statement can be true or false.
+statement can be true or false,
+resulting in the body of the \verb|if| statement being either
+exectuted or ignored.
 The cyclomatic complexity correlates with code complexity,
@@
 However, to calculate the variance of a distribution, at least two values
-are required. 
+are required (else a division by zero occurs in the calcation of the variance). 
 When the variance of (zero or) only one value is requested,
 an error message can helpfully indicate that the function 
-needs at least two values.
+needs at least two values to work as expected.
 
@@
 Publishing raw code is already better than not publishing code at all,
-as that raw code can be assumed to have run successfully at least once.
+as that raw code can reasonably be assumed to have run successfully at least once.
 To reproduce an experiment from raw code can be painful, 
 yet it is usually better than writing the same code from scratch.
+In general, it appears that sharing code is rare \cite{stodden2011trust,read2015sizing}.
+
+% \paragraph{Raw code can be FAIR data}
+
+It is hard to make raw code be FAIR data. It is hard for computers to find
+copy-pasted code in a Supplementary Materials. 
+That code is usually only accessible
+if the paper has been published under an open access license.
+When code can be found and accessed, it is an open question if the
+code can \emph{still} interoperate with current datasets, software
+and hardware. Raw code usually lacks the metadata to describe its needs.
+I would say that raw code, albeit better than no code, fits
+FAIR data only poorly.
 
@@
 Code is sometimes published alongside a paper, 
-in the form of a hosted website.
+in the form of a code hosting website.
 This way of publishing code gives additional options,
@@
 GitHub is currenly the most popular website to host code in general,
-with millions of users.
+with millions of users (a simple user search can confirm this).
 GitHub users can create dedicated websites (called 'repositories')
@@
 An example from the author is the website \cite{bbbqarticleissue157},
-that hosts part of the code for the paper \cite{bilderbeek2022transmembrane},
-as shown in Figure \ref{fig:bbbqarticleissue157}.
+that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}
+\iffalse
+, as shown in Figure \ref{fig:bbbqarticleissue157}
+\fi
+.
 
@@
 of computational experiments.
+Here, I discuss those features and why these are important.
 
@@
 
-Hosted code improves transparency \cite{gorgolewski2016practical}.
-The history of changes, called the commit history, 
+Hosted code improves transparency \cite{gorgolewski2016practical},
+as, among others, the history of changes (i.e. the git commit history), 
 helps to assess the honesty in the reporting of the code. 
 Ideally, the full version history of the code 
-has been generated, by hosting the code directly at its inception.
+has been generated, by hosting the code directly at its inception,
+so that the ontology of the experiment can be viewed.
 However, in some repositories, only the final version of the code has been
 posted, which means all historical information is lost.
-From a code history, it can be observed, for example, 
-how much statistical tests have really/actually been done 
+The code history has many important properties,
+as, for example, 
+it can be observed how much statistical tests have really/actually been done 
 and see if all are reported:
@@
 Some code hosts, among others, GitHub, 
-allows the researcher to start software scripts when uploading a new version.
-To determine the reproducibility of the code, one such script should
-run the code (on trivial data) to verify if the code works at all.
+allow the researcher/user to start software scripts upon uploading a change.
+This is useful to determine the reproducibility of the code: 
+one such script should run the code (on trivial data) 
+to verify if the code works at all.
 If the functionality of the code is lost, the researcher
@@
 A formalized version of this workflow is called continuous integration (CI).
-Using a CI service (such as GitHub Actions), is known to significantly 
+Using a CI service (such as GitHub Actions for GitHub), 
+is known to significantly 
 increase the number of bugs exposed \cite{vasilescu2015} and increases
@@
 a repository. For example, Figure \ref{fig:build_badge} shows a 
-build badge that indicates that a LaTeX document (i.e. this paper)
+build badge that indicates that a \LaTeX document (i.e. this paper)
 could be built.
@@
 the correctness of code (an interesting example is \cite{rahman2020exploratory}
-showing bugs in scientific software on the COVID-19 pandemic. 
-It recommends more testing).
+showing bugs in scientific software on the COVID-19 pandemic:
+it recommends more testing).
 The percentage of (lines of) code tested is called the code coverage.
@@
 Also here, CI opens up the possibility to informally annotate the
-code coverage. For example, Figure \ref{fig:badge_codecov} shows a 
-build badge that indicates that certain R package's code 
-has a 100\% code coverage.
+code coverage. 
+For example, build badges
+\iffalse
+(e.g. as shown in Figure \ref{fig:badge_codecov})
+\fi
+are visual indications on a website that can, among others,
+show a certain R package's code has a 100\% code coverage.
+
 Hence, code coverage is important metadata, that can automatically
@@
 In that way, authors involved in the writing of the software
-are honestly acknowledged. Figure \ref{fig:daisie_contributors}
+are honestly acknowledged. 
+\iffalse
+Figure \ref{fig:daisie_contributors}
 shows the number of commits the authors of DAISIE \cite{etienne2020daisie}
 have made to its code.
+\fi
 The distribution of contributed code per author 
@@
 software.
-Where this paper recommends using older software foundations 
+Where that paper recommends using older software foundations 
 with multiple implementations \cite{hinsen2019dealing},
-there is a modern solution that potentially has more longevity:
+this paper recommends a modern solution that potentially has more longevity:
 to publish the code installed in a virtual environment
@@
 There is another problem that authors of academic code face,
-which is that they will be contacted with a user sending in a bug report.
+which is that they may be contacted with a user sending in a bug report.
+HIERO
 One solution for the author is to ignore such emails

### 2022-08-24 article.tex
Progress

@@
 yet it is usually better than writing the same code from scratch.
-In general, it appears that sharing code is rare \cite{stodden2011trust,read2015sizing}.
+In general, it appears that sharing code is 
+rare \cite{stodden2011trust,read2015sizing},
+with a pleasant exception being \cite{conesa2019making}.
 
@@
 There is another problem that authors of academic code face,
-which is that they may be contacted with a user sending in a bug report.
-HIERO
+which is that they may be contacted with a user 
+asking for technical report 
+(against which a 'no' is just fine \cite{barnes2010publish})
+or -more relevant to the research- sending in a bug report.
 One solution for the author is to ignore such emails
 and it can be argued that no energy should be wasted on published code
-and work on something new instead.
+and work on something new instead 
+(see \cite{barnes2010publish} for a better way to deal with the problem).
 However, upon the find of a new bug, the question is:
@@
 
-Code is harder to preserve than an English text.
+Code is harder to preserve than an English text
+and preserving code is rarely done \cite{barnes2010publish}.
 

### 2022-08-24 article.tex
EOD

@@
 result in the same conclusions?
-A code hosting website usually has a place to report bugs
-that offer transparency in the handling of bug reports.
+To prevent authors chosing to hide bug reports, 
+a code hosting website usually has a public place to report bugs,
+to increase transparency in the handling of bug reports.
 
@@
 The most reproducible way of submitting the code of an experiment,
-is by providing the code with its dependencies installed in a virtual container.
+is by providing the code with all the (software) dependencies 
+it needs installed in a virtual container.
 This is close to the golden standard suggested by 
-\cite{peng2011reproducible} (see also Figure \ref{fig:peng2011reproducible}),
-however, that was before virtualization existed.
+\cite{peng2011reproducible} 
+\iffalse
+(see also Figure \ref{fig:peng2011reproducible})
+\fi
+,
+however, that article was written before virtualization existed.
 Virtualization allows one to provide a standard environment,
 that can be used on any other computer.
-The best known examples of container software are Docker and Singularity,
-where the latter is the only viable option to be able to run
-an experiment on a high performance computing (HPC) environment,
-such as a computer cluster.
+The best known examples of container software are Docker and Singularity.
+Due to access rights, Singularity is the only viable option to run
+an experiment on a 
+high performance computing (HPC) environment (e.g. a computer cluster).
 

### 2022-08-25 article.tex
Add sensitive data

@@
 of an experiment into a container is a good solution, as it 
-can stand the test of time well. Rewarding this, however, is not
-done yet by journals.
+can stand the test of time better.
+Creating such a container, however, requires more skill
+that -as of today- is not rewarded,
+partly as containers follow FAIR data principles only poorly.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sensitive data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For a computational experiment to be reproducible,
+an independent person should be able to
+reproduce the same results.
+The reproduce the same results, one needs the same
+input data and the same code (and the same 
+dependencies and environment).
+When the input data is sensitive, however, and thus cannot be shared with
+an independent researcher, he/she thus cannot reproduce the
+experiment.
+
+The solution for this problem is simple:
+supply a simulated dataset with the code.
+In the case of genetic epidemiology this would mean
+simulated genotypes and associated phenotypes,
+as can be done with the \verb|plinkr| R package \cite{plinkr}.
+This solution, however, is known to all developer that follow
+good practices when writing code, as public (i.e. simulated) 
+datasets are needed for automatic testing of code.
+One extra benefit of simulated data is that these can be used
+as a benchmark, as slightly different analyses should give 
+similar conclusions.
 
@@
 
-Code is paradata, as it is describes how the results are generated.
-The part of a scholarly article that describes the experiment is metadata, 
-as it describes what the code is thought to do.
+Code is paradata, as it directly describes how the results are generated.
+The part of a scholarly article 
+that describes the experiment in a natural language is metadata, 
+as it describes indirectly what the code does (or should be doing).
 The code that does the work in a computational experiment should
@@
 data with assured privacy \cite{zhang2016review,azencott2018machine}.
+Additionally, to ensure that
 

### 2022-08-25 article.tex
Went through it

@@
 
+Code has additional useful information, similar to confidence intervals,
+that allow a reader to gauge how much he/she trusts the results.
+The more of the best practices are followed, 
+the more trustworthy are the results.
+The most important best practices discussed in this paper are
+automatic testing, having a high code coverage and code
+of low (cyclomatic) complexity.
+In academia, to uncover the truth, code correctness is essential,
+similar to cell biologists working sterile to not contaminate their
+cell cultures.
+These best practices are vital for a computational biologist.
+
 For research to be reproducible, one ideally has access to
@@
 data with assured privacy \cite{zhang2016review,azencott2018machine}.
-Additionally, to ensure that
-
-Code has additional useful information, similar to confidence intervals,
-that allow a reader to gauge how much he/she trusts the results.
-The most important way to determine the quality of code
-is the amount of unit tests.
-When following a set of best practices, such as DevOps, TDD, Agile,
-writing unit tests is an essential 
-part in writing code.
-The amount of unit tests is an honest signal 
-for code correctness (i.e. it does what it is supposed to do, as opposed
-to 'it does something').
-In academia, to uncover the truth, code correctness is essential,
-similar to cell biologists working sterile to not contaminate their
-cell cultures.
-Unit tests are a vital practice for a computational biologist.
+Additionally, it follows software development's best practices 
+(especially the practice to test code automatically) 
+to release a simulated/public dataset, with
+the additional benefit that this dataset can be used for comparisons.
 
@@
 and preserving code is rarely done \cite{barnes2010publish}.
-
 Although code is the primary actor in computational experiments,
@@
 Most academic journal do not require authors to submit their code,
-nor it the submitted code peer reviewed.
-
-Although the code of computation experiments can be archived well, 
-there is no incentive to do so.
-
-Although a runnable version of a computation experiment can be archived well, 
-there is no incentive to do so.
+nor is the submitted code peer reviewed, 
+rOpenSci \cite{ram2018community} being the pleasant exception.
+
+Code used in academic research complies badly with FAIR principles,
+where unpublished code is the worst offender.
+However, any idealistic researcher does not have the
+tools to following FAIR principles in an exemplary way:
+there are no standards in how a scholarly article
+refers to academic code (i.e. in a machine-friendly way).
+Creating a (Singularity) container is the best a researcher
+can do to ensure his/her code is reusable,
+yet, there are no incentives to make code accessible, interoperable
+or re-usable, with standarized metadata attributes being absent.
 
@@
 Here we repeat the definition of 'paradata' and discuss 
-its (numbered) constituents.
-This paper defined paradata as 'data (1) about the collecting (2) of the data (3)',
-where (1) code must be seen as data, (2) downloading raw data
-and doing calculations must be seen as collecting, and (3) the
-results of an experiment must be seen as data.
-This paper argues, that (1) code is data in the form of text spread
-over one or more files, that has useful measurable properties, 
-(2) downloading raw datas and doing calculations, such as a T-test,
-does describe how the bits and pieces of an end result is collected,
-and (3) an experimental results is data, as it can be measured and
-used as the raw data of a next experiment.
+its constituents.
+This paper defined paradata as 'data about the collecting of the data'.
+This implies that (1) code must be seen as data.
+This paper argues, that code is data in the form of text spread
+over one or more files with useful measurable properties.
+(2) downloading raw data and doing calculations must be seen as collecting.
+This paper argues that downloading raw datas and doing calculations, 
+such as a T-test,
+does describe how the bits and pieces of an end result are collected.
+(3) the results of an experiment must be seen as data.
+This paper argues that an experimental results is data, 
+as it can be measured and can be used as the raw data of a next experiment.
 
@@
 Publishing code may be disadvantageous for an author.
-For science, yes, as this allows reproducible research.
+For science, code should be published, 
+this allows reproducible research 
+(again, see \cite{haibe2020importance} for a tragic example).
 For an author, publishing code alongside an experiment opens up
 the possibily to receive questions regarding that code.
-Note, however, that not publishing code will always thwart
-the reproduction of incorrect code, at the cost of a scientific
-career \cite{baggerly2009deriving}.
+Note, however, that not publishing code may put 
+oneself in the focus of attention
+and -after much effect by others reproducing an incorrect result-
+at the cost of a scientific career \cite{baggerly2009deriving}.
 
@@
 publishing code alongside an experiment opens up
-the possibily to receive questions regarding that code.
+the possibily to receive questions regarding that code,
+as well as other interactions, including helpful contributions.
 
@@
 be treated like a first class citizen. As doing so in an exemplary way
-in yet to be rewarded, hence it has to be the idealististic scientists
+is yet to be rewarded, hence it has to be the idealististic scientists
 to wage this battle. I feel the truth and science are worth fighting for

### 2023-01-24 article.tex
Remove all usages of 'I', except in the last sentence

@@
 for example, using a Singularity container.
-Here I will go into details of each of these ways,
-by describing some advantages, disadvantages, as well as how to
+In this chapter, details of each of these ways will be described,
+including advantages, disadvantages, as well as how to
 make the code more FAIR data \cite{wilkinson2016fair}.
@@
 and hardware. Raw code usually lacks the metadata to describe its needs.
-I would say that raw code, albeit better than no code, fits
+One could say that raw code, albeit better than no code, fits
 FAIR data only poorly.
@@
 of computational experiments.
-Here, I discuss those features and why these are important.
+Below, those features and why these are important are discussed.
 
@@
 four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
-here I will use the more modern (and in this context superior)
-definition of software collapse \cite{hinsen2019dealing},
+here a more modern (and in this context superior)
+definition is used: software collapse \cite{hinsen2019dealing},
 in which software fails due to dependencies on other 

### 2023-01-24 article.tex
Remove all usages of `'ll`, except in the last sentence

@@
 epidemiology in particular.
-We'll postpone the introduction to genetic epidemiology now
-and focus on 'computational experiment' first.
+The introduction to genetic epidemiology will be postponed for now,
+insteaf focussing on 'computational experiment' first.
 A computational experiment is an experiment run by computer code.
@@
 experiment; an article describes what the code does in a natural 
-language, such as English (which I'll use from now on).
+language, such as English.
 However, English (or any other natural language) is not the best

### 2023-01-24 article.tex
Correct singular/plurality

@@
 Common elements in the definitions of paradata are 
-(1) paradata are usually not the primary data, (2) paradata describe
-a process, (3) paradata describe the way of collecting primary data, or 
-(4) paradata describe the way the primary data is generated.
+(1) paradata is usually not the primary data, (2) paradata describes
+a process, (3) paradata describes the way of collecting primary data, or 
+(4) paradata describes the way the primary data is generated.
 
@@
 similar website to upload containers, as the metadata for these
-sites are limited to free-text comments and tags (DockerHub)
+sites is limited to free-text comments and tags (DockerHub)
 or not metadata at all (Syslabs).

### 2023-01-24 article.tex
Data is plural

--- 2023-01-24 article.tex
@@
-%
-% 'Data are ...', i.e. use plural
-%
-
 \begin{abstract}
 
-Here we define paradata as the data that describes the process of generating data.
-In genetic epidemiology, the data generated is mostly the results 
+Here we define paradata as the data that describe the process of generating data.
+In genetic epidemiology, the data generated are mostly the results 
 of an analysis (e.g. predicting a person having a disease),
@@
 In this chapter it is argued that the source code should should be supplied,
-as this is the true paradata: if the paper and code disagree, it is the
+as these are the true paradata: if the paper and code disagree, it is the
 code that has generated the results.
@@
 In this paper, the definition of paradata
-is 'information about how data was generated',
+is 'information about how data were generated',
 as shown schematically in the left panel of Figure \ref{fig:figure_1}.
-Paradata does not have a clear definition yet (hence this book). 
+The term 'paradata' does not have a clear definition yet (hence this book). 
 Common elements in the definitions of paradata are 
@@
 a process, (3) paradata describe the way of collecting primary data, or 
-(4) paradata describe the way the primary data is generated.
+(4) paradata describe the way the primary data are generated.
 
@@
 
-A simple example would be the data at eBird (\url{https://ebird.org/}),
+A simple example would be data at eBird (\url{https://ebird.org/}),
 where people can submit wher a bird of a certain species was spotted.
@@
 
-Paradata is relevant, as it allows primary data to be re-used.
+Paradata are relevant, as it allows primary data to be re-used.
 In the eBird example, it helps assess the trustworthiness of the 
@@
 to illustrate that the computer code is paradata,
-that this paradata is relevant, as it helps 
+that this type of paradata is relevant, as it helps 
 to assess the correctness of the results,
@@
 can be made for that code as in the example: that code is 
-the ground truth of how the raw data is converted into useful genetic data.
+the ground truth of how the raw data are converted into useful genetic data.
 
@@
 Programming code (from now on: 'code') is paradata, 
-as it is data that describes how data is generated.
+as it is data that describes how data are generated.
 Code is data that is usually in the form of text, 
@@
 However, English (or any other natural language) is not the best
-candidate to describe how the data is collected,
+candidate to describe how the data are collected,
 as it only has a loose connection with collecting the data.
@@
 and automatically generated metrics.
-A code hosting website produces and shares this metadata.
-This metadata can help to determine the
+A code hosting website produces and shares these metadata.
+These metadata can help to determine the
 reproducibility, correctness and quality of code
@@
 are different, such as the random numbers as generated by an operating
-system, or data that is downloaded from additional online sources.
+system, or data that are downloaded from additional online sources.
 
@@
 dependencies and environment).
-When the input data is sensitive, however, and thus cannot be shared with
-an independent researcher, he/she thus cannot reproduce the
+When the input data are sensitive, however, and thus cannot be shared with
+an independent researcher, one thus cannot reproduce the
 experiment.
@@
 both the data used and the code.
-In some fields, such as genetic epidemiology, the data is
+In some fields, such as genetic epidemiology, the data are
 sensitive, hence cannot be released,
@@
     Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
-    Paradata  & Data that describes the process of generating data    & The code to conclude that a SNP has a significant association   \\
+    Paradata  & Data that describe the process of generating data     & The code to conclude that a SNP has a significant association   \\
     Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
-    Primary data & Data that is worked with in a research             & The genotype of individuals   \\
-    Metadata  & Data that provides information about other data       & The article that describes an experiment                        \\
-    Raw data  & Data directly from the real world, that is not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
+    Primary data & Data that are worked with in a research            & The genotype of individuals   \\
+    Metadata  & Data that provide information about other data        & The article that describes an experiment                        \\
+    Raw data  & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
     Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         

### 2023-01-24 article.tex
No he/she
Fix typo: 'fictuous' -> 'fictitious'

@@
 may end up in the clinic.
-For the sake of having a clear (yet fictuous) example,
+For the sake of having a clear (yet fictitious) example,
 let's assume that a high level of IL6RA 
@@
 Code has additional useful information, similar to confidence intervals,
-that allow a reader to gauge how much he/she trusts the results.
+that allow a reader to gauge how much one trusts the results.
 The more of the best practices are followed, 

### 2023-01-25 scripts/edit.sh
Remove TOC

@@ -1,2 +1,2 @@
 #!/bin/bash
-mousepad header.tex article.bib article.tex
+mousepad header.tex article.bib article.tex feedback/feedback_20221209.md

### 2023-01-25 article.tex
Read all feedback
Use Springer style more
Follow Springer guidelines
Definitions below abstract
Remove table of recommendations
EOD

@@
-\begin{abstract}
+\setcounter{chapter}{7}
+\chapter{Research code is as vital paradata as the article itself}
+
+\abstract{
 
@@
 
-\end{abstract}
+} % abstract
 
@@
 genetic epidemiology
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Definitions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[h]
+  \begin{tabular}{lp{5cm}p{5cm}}
+    Term      & Definition                                            & Example                                                         \\
+    \hline
+    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
+    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
+    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
+    Paradata  & Data that describe the process of generating data     & The code to conclude that a SNP has a significant association   \\
+    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
+    Primary data & Data that are worked with in a research            & The genotype of individuals   \\
+    Metadata  & Data that provide information about other data        & The article that describes an experiment                        \\
+    Raw data  & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
+    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
+  \end{tabular}
+  \caption{Terms used in this paper, and their definitions}
+  \label{tab:definitions}
+\end{table}
 
@@
 
-Two different researchers write two different papers,
-that are both accepted. One researcher does not supply the
-computer code that was used to generate the results.
-The experiment of the other researcher can be re-done easily
-and you can verify that, indeed, the results are the same
-as in the paper.
-Do you trust the conclusion of their papers equally?
-
-% \paragraph{Definition of paradata}
-
-In this paper, the definition of paradata
-is 'information about how data were generated',
-as shown schematically in the left panel of Figure \ref{fig:figure_1}.
-The term 'paradata' does not have a clear definition yet (hence this book). 
-Common elements in the definitions of paradata are 
-(1) paradata are usually not the primary data, (2) paradata describe
-a process, (3) paradata describe the way of collecting primary data, or 
-(4) paradata describe the way the primary data are generated.
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\linewidth]{figure_1.png}
-  \caption{
-    Left: general relation between data, paradata and metadata.
-    Middle: the same relations specified for eBird.
-    Right: the same relations specified for genetic epidemiology.
-  }
-  \label{fig:figure_1}
-\end{figure}
-
-% \paragraph{Simple example of paradata}
-
-A simple example would be data at eBird (\url{https://ebird.org/}),
-where people can submit wher a bird of a certain species was spotted.
-The name of the person that spotted that bird is paradata, 
-as it is that person that generated the data and determined of
-which species the bird was, 
-as shown schematically in the middle panel of Figure \ref{fig:figure_1}.
-
-% \paragraph{Relevance of paradata}
-
-Paradata are relevant, as it allows primary data to be re-used.
-In the eBird example, it helps assess the trustworthiness of the 
-data, as beginner birdwatchers tend to make more mistakes in identifying
-the correct bird species, as compared to experts.
-By using the name and/or experience of the birder as a covariate,
-we can re-use the dataset in other contexts as well,
-as now this example can be generalized to
-a classification problem of unlabeled
-data for learning agents with different amounts 
-of experience.
+Two different researchers in genetic epideomology (more on that field later), 
+write two equally good manuscripts
+that involve computation.
+Both manuscripts are accepted by an equally prestigous journal after peer view. 
+One researcher, however, does not supply the
+computer code that was used to generate the results,
+where the other does.
+Are the conclusion of these papers to be trusted equally?
+Is this difference relevant and worth the effort?
+This chapter discusses the effect of supplying the computer 
+code (from now on: 'code') 
+and claims that code is paradata and that this paradata 
+is just as important as the paper itself.
+
+% \paragraph{The earliest definition of paradata}
+
+Paradata are data about the data collection process \cite{choumert2019using}.
+The concept of paradata (although not named as such yet) 
+was introduced by \cite{couper1998measuring},
+who developed a computer-assisted interview program
+that, among others, records all key strokes,
+measures the time used to answer each question 
+and even the time the monitor is turned off.
+The goal in that context was to assess and improve survey quality.
+
+% \paragraph{Modern forms of paradata}
+
+Nowadays, the lack of paradata can be seen as 'a drastic constraint'
+in data use \cite{huvila2022improving}.
+Given the stance that code to do a computational experiment is paradata, 
+the effect of its absence is 
+that a computational experiment cannot be redone easily.
 
@@
 The introduction to genetic epidemiology will be postponed for now,
-insteaf focussing on 'computational experiment' first.
+instead focussing on 'computational experiment' first.
 A computational experiment is an experiment run by computer code.
@@
 an experiment FAIR data (findability, accessibility, interoperability, and reusability) \cite{wilkinson2016fair} better.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{figure_1.png}
+  \caption{
+    Left: general relation between data, paradata and metadata.
+    Middle: the same relations specified for eBird.
+    Right: the same relations specified for genetic epidemiology.
+  }
+  \label{fig:figure_1}
+\end{figure}
 
@@
 Using CI opens up the possibility to informally annotate features of
-a repository. For example, Figure \ref{fig:build_badge} shows a 
+a repository. 
+\iffalse
+For example, Figure \ref{fig:build_badge} shows a 
 build badge that indicates that a \LaTeX document (i.e. this paper)
 could be built.
+\fi
 
@@
 
+% Reviewer 1: this section starts with a sentence that appears to undercut the whole theme of this chapter!
 Code may or may not be paradata, depending on how the definition
@@
 
+\iffalse
 \rowcolors{2}{gray!10}{gray!40}
@@
 \end{table}
+\fi
 
@@
 This chapter has not been supported by funding.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Definitions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[h]
-  \begin{tabular}{lp{5cm}p{5cm}}
-    Term      & Definition                                            & Example                                                         \\
-    \hline
-    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
-    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
-    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
-    Paradata  & Data that describe the process of generating data     & The code to conclude that a SNP has a significant association   \\
-    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
-    Primary data & Data that are worked with in a research            & The genotype of individuals   \\
-    Metadata  & Data that provide information about other data        & The article that describes an experiment                        \\
-    Raw data  & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
-    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
-  \end{tabular}
-  \caption{Terms used in this paper, and their definitions}
-  \label{tab:definitions}
-\end{table}

### 2023-01-26 article.tex
Introduction is cleaner

@@
 
+% Alexa and Blake?
 Two different researchers in genetic epideomology (more on that field later), 
@@
 
-% \paragraph{The earliest definition of paradata}
-
-Paradata are data about the data collection process \cite{choumert2019using}.
+% \paragraph{The first definition paradata}
+
 The concept of paradata (although not named as such yet) 
@@
 
+% \paragraph{Definition of paradata}
+
+This paper defines paradata as 'data about the data collection 
+process' \cite{choumert2019using}.
+This definition is chosen for its simplicity, 
+yet there are multiple related and/or more nuanced 
+definitions of paradata (see \cite{huvila2022improving} for an overview).
+
+% Code is paradata
+
+The code used in experiments is just that, 'data about the data collection 
+process', as it commonly downloads data, selects relevant subsets in those
+downloaded data and performs statistical tests to generate results.
+Code, hence, is paradata and this chapter explores and illustrates the
+consequence of that premise.
+
 % \paragraph{Modern forms of paradata}
@@
 in data use \cite{huvila2022improving}.
-Given the stance that code to do a computational experiment is paradata, 
-the effect of its absence is 
+Computer code, being paradata, the effect of its absence is, most obviously,
 that a computational experiment cannot be redone easily.
+yet there are wider implications as well.
 
@@
 epidemiology in particular.
-The introduction to genetic epidemiology will be postponed for now,
-instead focussing on 'computational experiment' first.
-A computational experiment is an experiment run by computer code.
-In chronologic order, it usually collects primary data first, 
-then runs an analysis, which finally results in the generation of results.
-
-This paper takes the standpoint that it is the computer code 
-that describes this generation of results, 
-hence it is the computer code that is paradata 
-(i.e. not the description of the experiment as done in a natural language),
-as shown schematically in the right panel of Figure \ref{fig:figure_1}.
-Reading the computer code allows one to assess the trustworthiness 
-of the experiment, similar to knowing the name or expertise of a bird spotter.
-Having the code allows one to 
-verify, re-use and extend the experiment, 
-in the same or other contexts, providing for reproducible research.
-Additionnally, this paper suggests some best practices on how to make 
-an experiment FAIR data (findability, accessibility, interoperability, and reusability) \cite{wilkinson2016fair} better.
+It is argued that code, being paradata, is as vital as the academic
+paper itself, illustrated by the two scientists at the start of 
+this chapter. After that, the implications and recommendations
+follow.
 
@@
 could be used as an example.
+
+% \paragraph{Publication of code}
+
+Going back to the example scientists in the introduction,
+where one does and the other does not publish the code,
+it is not common to publish the code of an experiment or analysis.
+For example, in computer graphics, 
+a field intimately familiar with computer code,
+only 192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
+Another study analysed the reproducibility of registered reports
+in the field of psychology, 
+where still only 37 out of 62 studies supplied the code 
+to redo the analysis \cite{obels2020analysis}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{This paper}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This paper has been produced following all recommendations in it:
+it's \latex code is hosted on GitHub 
+at \url{https://github.com/richelbilderbeek/chapter_paradata}.
+Due this, one can, among others, see the complete version history of this paper.
+This alone could allow research to be done on, for example,
+the effect of reviewers' feedback on a first manuscript.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Funding}

### 2023-01-27 article.tex
Add Barba

@@
 yet there are multiple related and/or more nuanced 
-definitions of paradata (see \cite{huvila2022improving} for an overview).
+definitions of paradata (see for example \cite{huvila2022improving} 
+and \cite{skold2022interrogating}).
 
@@
 Computer code, being paradata, the effect of its absence is, most obviously,
-that a computational experiment cannot be redone easily.
+that a computational experiment cannot be redone 
+easily,
 yet there are wider implications as well.
@@
 the academic culture to reproduce results 
-has been lost over time \cite{peng2011reproducible}.
+has been lost over time in general \cite{peng2011reproducible},
+with labs that embrace reproducibility (for example, \cite{barba2016hard})
+are the exception.
+
 That same study recommended, as a solution to counter this, 

### 2023-01-27 article.tex
Add weissgerber2018we

@@
 
-Here we define paradata as the data that describe the process of generating data.
+Here we define paradata as data about the data collection process.
 In genetic epidemiology, the data generated are mostly the results 
@@
 In such context, one could argue incorrectly that
-it is the -usually English- scientific paper that best describes which computation steps 
+it is the scientific paper that best describes which computation steps 
 have taken place.
@@
 assumption that there is a perfect match between the paper and the code.
-In this chapter it is argued that the source code should should be supplied,
+In this chapter it is argued that the source code should be supplied,
 as these are the true paradata: if the paper and code disagree, it is the
@@
 
-% Alexa and Blake?
+% Alex and Blake?
 Two different researchers in genetic epideomology (more on that field later), 
@@
 \begin{verbatim}
-We compared the values of x and y using a one-tailed Student's T-test,
+We compared the values of x and y using an unpaired one-tailed Student's T-test,
 as we expect the average value of x to be less than the average value of y.
@@
 We ignore the choices of words and style of this sentence: what is
-important is the content: a one-tailed T test is performed.
+important is the content: an unpaired one-tailed T test is performed.
 Taking a look at the (in this case, the programming language R) code, 
@@
 but the code is correct:
-\verb|t.test| is indeed the name of an R function to do a T-test
+\verb|t.test| is indeed the name of an R function to do an unpaired T-test
 and it is reasonable to assume that this code, when glancing over the code
@@
 results in (needlessly) less conclusions drawn in this fictional paper.
+
+To approach the same problem from a different angle:
+correctly reporting the T-test in a reproducable ways
+is already a problem, with only 80 out of a selected 179 papers
+fully reporting the parity of T-test (i.e. paired or unpaired) 
+\cite{weissgerber2018we}.
+In those unresolved 99 papers, 
+the code (if published) would clarify this.
 
@@
 This chapter has not been supported by funding.
+
+% Duran et al., grounds for trust
+% [A] process is epistemically opaque relative to a cognitive agent X at time t
+% just in case X does not know at t all of the epistemically relevant elements of
+% the process (Humphreys 2009, 618)
+% 
+% On Computational reliability
+% 
+% (CR) if S’s believing p at t results from m, then S’s belief in p at t is justified.
+% where S is a cognitive agent, p is any truth-valued proposition related to the
+% results of a computer simulation, t is any given time, and m is a reliable com‑
+% puter simulation.
+%
+%
+% 4 Sources for Computational Reliabilism
+%
+% 1. Verification and validation methods
+% 2. Robustness analysis for computer simulations
+% 3. A history of (un)successful implementations
+% 4. Expert knowledge

### 2023-01-27 article.tex
T test example is genetic epidemiological

@@
 To illustrate how easy it is to get a mismatch between a paper
-and the code, consider this example
-of a text in a fictional paper:
+and the code, 
+consider this example of a possible text in the fictional 
+genetic epidemiology paper:
 
 \begin{verbatim}
-We compared the values of x and y using an unpaired one-tailed Student's T-test,
-as we expect the average value of x to be less than the average value of y.
+We compared the in-blood concentrations of IL6RA 
+between subjects having AA versus subjects having GG at SNP rs12133641,
+using an unpaired one-tailed T-test,
+as we expect the average concentration of IL6RA for those with AA 
+to be less than the average concentration in those with GG.
 \end{verbatim}
@@
 \begin{verbatim}
-t.test(x, y)
+t.test(subjects_with_aa, subjects_with_gg)
 \end{verbatim}
@@
 
-To approach the same problem from a different angle:
+Note that there is another side of this same coin:
 correctly reporting the T-test in a reproducable ways
-is already a problem, with only 80 out of a selected 179 papers
+is known to be a problem, with only 80 out of a selected 179 papers
 fully reporting the parity of T-test (i.e. paired or unpaired) 

### 2023-01-27 article.tex
Progress

@@
 
-Nowadays, the lack of paradata can be seen as 'a drastic constraint'
-in data use \cite{huvila2022improving}.
-Computer code, being paradata, the effect of its absence is, most obviously,
+A recent paper explores the use of paradata to increase the impact of data,
+stating that lack of paradata can be seen as 'a drastic constraint'
+in the use of data and offer some suggestions to 
+make paradata useful for data re(use):
+paradata should be comprehensive, comprehensively documentated
+and be friendly to computers \cite{huvila2022improving}.
+These suggestions appear to work with code as well,
+where these suggestions show up as best practices (such
+as \cite{stodden2013best}), some of which
+described below.
+The lack of computer code is just as much a drastic constraint
+in the use of a scientific experiment.
+The clearest effect of its absence is, most obviously,
 that a computational experiment cannot be redone 
-easily,
-yet there are wider implications as well.
+easily, yet there are wider implications, described below, as well.
 

### 2023-02-02 article.tex
Improve

@@
 make paradata useful for data re(use):
-paradata should be comprehensive, comprehensively documentated
-and be friendly to computers \cite{huvila2022improving}.
+paradata should be comprehensive, comprehensively documented
+and friendly to computers \cite{huvila2022improving}.
 These suggestions appear to work with code as well,
@@
 described below.
-The lack of computer code is just as much a drastic constraint
+The lack of computer code is just as much 'a drastic constraint'
 in the use of a scientific experiment.
@@
 which are (1) the genotypes \cite{johansson2013identification},
-(2) the phenotypes, in this case, concentrations of certain proteins in the blood \cite{enroth2014strong,enroth2015effect}.
+(2) the phenotypes, in this case, concentrations of certain proteins in the 
+blood \cite{enroth2014strong,enroth2015effect}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Code is more important than the paper}
+\section{Code is just as important as the paper}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and the phenotype (in this case, the concentration of IL6RA).
-To detect this association, 
-there was no fieldwork nor lab-work involved.
-For this experiment, 
-it has been the researcher that has written the code,
-after which that code does the actual work. 
-A genetic epidemiologist does not need a lab, but computational power instead.
-
-To \emph{obtain} the raw genetic data, there may have been fieldwork and/or lab-work
-involved. For example, a researcher needs to go into the field to take
-samples (e.g. blood) of humans or other species. 
-To analyse these samples,
-a researcher needs a lab. 
-Depending on the analysis, there may be a big bioinformatics step
-to aggregate the measurements into usable data. 
-Although this paper does not focus on the code run to 
-aggragate raw and noisy measurements into usable data, the same arguments
-can be made for that code as in the example: that code is 
-the ground truth of how the raw data are converted into useful genetic data.
-
-% \paragraph{The code is the paradata for the results of an experiment}
-
-Programming code (from now on: 'code') is paradata, 
-as it is data that describes how data are generated.
-Code is data that is usually in the form of text, 
-spread over one or more files, that describes an experiment/annalysis.
-The computational experiment (from now on: 'experiment') collects the data
-and then looks for patterns.
-Those patterns (again, a type of data) are the results of an experiment.
-
-The (scientific) article is metadata, as it is data that describes other data:
-an article describes the
-experiment; an article describes what the code does in a natural 
-language, such as English.
-However, English (or any other natural language) is not the best
-candidate to describe how the data are collected,
-as it only has a loose connection with collecting the data.
-If the code and article of an experiment disagree,
-it is the code that actually let the data be collected.  
-Instead, an article is metadata about a research.
+
+The code for this experiment is vital paradata.
+Like all paradata, its omission would be a constraint in the
+use of the conclusion reached by it:
+we can use it to assess the trustworthiness of the finding,
+for example, by running the code again.
+Ideally, the code is comprehensive, comprehensively documented
+and be friendly to computers \cite{huvila2022improving}.
+Code being friendly to computer seems like an obvious 'yes',
+yet this is naive, as commonly it is hard to run code in practice. 
+% TODO: Link to correct paragraphs
+For code to be called computer-friendly, it must be run on an independent entity, 
+as is discussed in the paragraphs about hosting code and/or putting the code in a container.
+% TODO: Link to correct paragraphs
+Also note that there is a problem due to the sensitivity of the data,
+which is discussed in paragraph.
+Lastly, note that the researcher used data that was itself generated by code:
+complex bioinformatics pipelines are used to obtain reliable DNA sequences 
+from the biological sample.
 
@@
 
-One pseudorandom example of a mismatch between code and article
-was found by me in \cite{labrecque2019interpretation}.
-After the corresponding author of that paper was contacted,
-the code was updated and confirmed to have the same
-results as published in the paper.
-This sequence of events could only have happened, 
-because the authors chose to publish their code.
-
 It is the paper that accompanies the code,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{There are multiple way to publish code}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\subsection{Making code useful paradata}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Again, paradata should be comprehensive, comprehensively documented
+and friendly to computers \cite{huvila2022improving}.
+In this section, these ideal properties of paradata are applied to code.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Code must be extensively documented}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Code must be extensive}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Code must be reviewed
 As it is the code that is the most important actor in an the experiment,
@@
 
-There are multiple ways of publishing code,
-that require increasingly more knowledge to do and result
-in increasingly more reproducible experiments:
-(1) publishing the code as text only, e.g. by printing
-it in a Supplementary Materials, (2) publishing
-the code on a hosted website, e.g on GitHub, with
-added features (3) publishing 
-the installed code within a virtual environment,
-for example, using a Singularity container.
-In this chapter, details of each of these ways will be described,
-including advantages, disadvantages, as well as how to
-make the code more FAIR data \cite{wilkinson2016fair}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Publishing raw code}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Code is commonly published alongside of a 
-paper, pasted as text in the Supplementary Materials 
-(for example, in \cite{labrecque2019interpretation}).
-This is the rawest form of publishing code,
-as it is just text.
-
-% \paragraph{Raw code is important}
-
-The raw code of a computational experiment is
-more important to preserve than the article describing
-the experiment, even when both describe the experiment identically,
-correctly and reproducibly, 
-as it the code that can reproduce the results.
-
+% Code must be clean
 % \paragraph{Raw code is an indicator of quality}
@@
 
-% \paragraph{The problems with raw code}
-
-There are multiple problems with publishing raw code.
-First, there is no guarantee that that code \emph{can} actually
-run, for example, due to a typo added after the code was pasted.
-Secondly, there is no way to determine if code that worked yesterday
-works today. These problems can be prevented by, instead of
-pasting raw code, using a website to host that code (see below).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Code must be computer friendly}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+There are multiple ways of publishing code,
+that require increasingly more knowledge to do and result
+in increasingly more reproducible experiments:
+(1) publishing the code as text only, e.g. by printing
+it in a Supplementary Materials, (2) publishing
+the code on a hosted website, e.g on GitHub, with
+added features (3) publishing 
+the installed code within a virtual environment,
+for example, using a Singularity container.
+In this chapter, details of each of these ways will be described,
+including advantages, disadvantages, as well as how to
+make the code more FAIR data \cite{wilkinson2016fair}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Publishing raw code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Publishing hosted code}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% \paragraph{Code must be hosted to be moderately computer-friendly}
 Code is sometimes published alongside a paper, 
@@
 similar conclusions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Code is paradata, as it directly describes how the results are generated.
-The part of a scholarly article 
-that describes the experiment in a natural language is metadata, 
-as it describes indirectly what the code does (or should be doing).
-The code that does the work in a computational experiment should
-always be published together with a scholarly article,
-as, if the paper and code differ, it is the code that generated the results.
-
-Code has additional useful information, similar to confidence intervals,
-that allow a reader to gauge how much one trusts the results.
-The more of the best practices are followed, 
-the more trustworthy are the results.
-The most important best practices discussed in this paper are
-automatic testing, having a high code coverage and code
-of low (cyclomatic) complexity.
-In academia, to uncover the truth, code correctness is essential,
-similar to cell biologists working sterile to not contaminate their
-cell cultures.
-These best practices are vital for a computational biologist.
-
-For research to be reproducible, one ideally has access to
-both the data used and the code.
-In some fields, such as genetic epidemiology, the data are
-sensitive, hence cannot be released,
-yet there are methods being devised to run code on sensitive
-data with assured privacy \cite{zhang2016review,azencott2018machine}.
-Additionally, it follows software development's best practices 
-(especially the practice to test code automatically) 
-to release a simulated/public dataset, with
-the additional benefit that this dataset can be used for comparisons.
-
-Code is harder to preserve than an English text
-and preserving code is rarely done \cite{barnes2010publish}.
-Although code is the primary actor in computational experiments,
-there is no incentive to submit code alongside a publication.
-Most academic journal do not require authors to submit their code,
-nor is the submitted code peer reviewed, 
-rOpenSci \cite{ram2018community} being the pleasant exception.
-
-Code used in academic research complies badly with FAIR principles,
-where unpublished code is the worst offender.
-However, any idealistic researcher does not have the
-tools to following FAIR principles in an exemplary way:
-there are no standards in how a scholarly article
-refers to academic code (i.e. in a machine-friendly way).
-Creating a (Singularity) container is the best a researcher
-can do to ensure his/her code is reusable,
-yet, there are no incentives to make code accessible, interoperable
-or re-usable, with standarized metadata attributes being absent.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Code is paradata, as it directly describes how the results are generated.
+The part of a scholarly article 
+that describes the experiment in a natural language is metadata, 
+as it describes indirectly what the code does (or should be doing).
+The code that does the work in a computational experiment should
+always be published together with a scholarly article,
+as, if the paper and code differ, it is the code that generated the results.
+
+Code has additional useful information, similar to confidence intervals,
+that allow a reader to gauge how much one trusts the results.
+The more of the best practices are followed, 
+the more trustworthy are the results.
+The most important best practices discussed in this paper are
+automatic testing, having a high code coverage and code
+of low (cyclomatic) complexity.
+In academia, to uncover the truth, code correctness is essential,
+similar to cell biologists working sterile to not contaminate their
+cell cultures.
+These best practices are vital for a computational biologist.
+
+For research to be reproducible, one ideally has access to
+both the data used and the code.
+In some fields, such as genetic epidemiology, the data are
+sensitive, hence cannot be released,
+yet there are methods being devised to run code on sensitive
+data with assured privacy \cite{zhang2016review,azencott2018machine}.
+Additionally, it follows software development's best practices 
+(especially the practice to test code automatically) 
+to release a simulated/public dataset, with
+the additional benefit that this dataset can be used for comparisons.
+
+Code is harder to preserve than an English text
+and preserving code is rarely done \cite{barnes2010publish}.
+Although code is the primary actor in computational experiments,
+there is no incentive to submit code alongside a publication.
+Most academic journal do not require authors to submit their code,
+nor is the submitted code peer reviewed, 
+rOpenSci \cite{ram2018community} being the pleasant exception.
+
+Code used in academic research complies badly with FAIR principles,
+where unpublished code is the worst offender.
+However, any idealistic researcher does not have the
+tools to following FAIR principles in an exemplary way:
+there are no standards in how a scholarly article
+refers to academic code (i.e. in a machine-friendly way).
+Creating a (Singularity) container is the best a researcher
+can do to ensure his/her code is reusable,
+yet, there are no incentives to make code accessible, interoperable
+or re-usable, with standarized metadata attributes being absent.
+
 % \paragraph{Final word}

### 2023-02-06 article.tex
Improve structure

@@
 \setcounter{chapter}{7}
-\chapter{Research code is as vital paradata as the article itself}
+\chapter{Research code is paradata and as vital as the article itself}
 
@@
 Here we define paradata as data about the data collection process.
-In genetic epidemiology, the data generated are mostly the results 
-of an analysis (e.g. predicting a person having a disease),
-as done by computer code.
-In such context, one could argue incorrectly that
-it is the scientific paper that best describes which computation steps 
-have taken place.
-However, this has the unrealistic 
-assumption that there is a perfect match between the paper and the code.
-In this chapter it is argued that the source code should be supplied,
-as these are the true paradata: if the paper and code disagree, it is the
-code that has generated the results.
+Computer code is the paradata for a computatioal/academic experiment,
+allowing the experiment to be reproduced.
+However, code is published alongside an academic
+paper in a minority of all cases.
+As reproducibility is a cornerstone of science,
+research code should be published.
+Earlier general recommendations about making paradata (more) useful
+is tailored to code, using the field of genetic epidemiology
+as an example.
 The chapter concludes by some rules how to better code to serve as paradata,
@@
 One researcher, however, does not supply the
-computer code that was used to generate the results,
+computer code (from now on: 'code') that was used to generate the results,
 where the other does.
@@
 This chapter discusses the effect of supplying the computer 
-code (from now on: 'code') 
+code  
 and claims that code is paradata and that this paradata 
@@
 The lack of computer code is just as much 'a drastic constraint'
-in the use of a scientific experiment.
+in the use of a (computational) scientific experiment.
 The clearest effect of its absence is, most obviously,
@@
 This paper uses genetic epidemiology as a specific example,
-to illustrate that the computer code is paradata,
+to illustrate that the code is paradata,
 that this type of paradata is relevant, as it helps 
@@
 where one does and the other does not publish the code,
-it is not common to publish the code of an experiment or analysis.
+it is not common to publish the code of an experiment or analysis 
+\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
+being \cite{conesa2019making}).
 For example, in computer graphics, 
@@
 as well as how these come to be.
-For example, we know that lactose intolerance is, among others,
-caused by producing too few lactose-degrading enzymes,
-and is most commonly found in south-east asia.
+For example, we know that lactose intolerance in adults is
+caused by a decline in the production of lactose-degrading enzymes,
+and is most commonly found in south-east asia and south africa \cite{storhaug2017country}.
 The trait is caused by the genetic make-up, 
@@
 The trait, also called phenotype (again, see Table \ref{tab:definitions}), 
-in this example is lactose intolerance,
+in this example is lactose intolerance at adult age,
 yet any human property, such as weight or height can be studied.
@@
 yet is preventable by lifestyle changes.
-Would this be the case, we can create a gene panel that specifically measures
+Would this be the case, we can create a tailored 
+experiment, called a gene panel, that specifically measures
 SNP \verb|rs12133641|. 
@@
 % TODO: Link to correct paragraphs
-Also note that there is a problem due to the sensitivity of the data,
+Also note that there is a problem due to the inherent sensitivity of the genetic data,
 which is discussed in paragraph.
@@
 The code of a computational experiment must be published
-for an experiment to reproducible.
+for an experiment to be easily reproducible.
 It is however (too) common that the original data
@@
 
-From a knowledge management's perspective,
-it is the code that must be archived 
-to conserve an experiment,
-and not the
-English description of what it does:
-it is the code does the actual work.
-Hence, from that perspective, emphasis should be
-put more on the preservation of code, as it is the most important actor
-in an experiment.
-
 % \paragraph{Example of code being the ground truth}
@@
 \begin{verbatim}
-t.test(subjects_with_aa, subjects_with_gg)
+t_test_result <- t.test(subjects_with_aa, subjects_with_gg)
 \end{verbatim}
 
-The line of code is simplified (as the result of the T test is never stored),
-but the code is correct:
+The code is correct:
 \verb|t.test| is indeed the name of an R function to do an unpaired T-test
@@
 Here, however, we see a mismatch between the English description and the code:
-by default, the R function \verb|t.test| does a \textbf{two}-tailed T-test.
+by default, the R function \verb|t.test| does 
+a \textbf{two}-tailed unpaired T-test.
 A consequence of this fictional example is that the published p-values are
@@
 
-But regardless of the size of the code, 
-it is the code that is the ground truth
-and should be treated as a first-class citizen
-in a computational experiment.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Making code useful paradata}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Making code useful paradata}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In this section, these ideal properties of paradata are applied to code.
+The resulting recommendations are compared with
+existing best practices in writing code.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Code must be extensive}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Code must be reviewed
+For code to be ideal paradata, it must be documented.
+However, code does explain itself to some extent,
+for example, the function name \verb|t.test| as used above.
+
+\begin{itemize}
+  \item Code must be documented to the right amount
+  \item Code must be reviewed
+\end{itemize}
+
+\paragraph{Code must be documented to the right amount}
+
+Code should be documented to the right amount:
+too little documentation is unhelpful, where too much is a needless burden
+on the programmer.
+The first recommended practice is to write the code in such a way 
+that it becomes self-explanatory \cite{wilson2014best}.
+What is left to document are the reasons behind the code, 
+its design and its purpose \cite{wilson2014best}.
+
+\paragraph{Code must be reviewed}
+
 As it is the code that is the most important actor in an the experiment,
@@
 the \verb|mcbette| R package \cite{bilderbeek2020mcbette}.
-
-% Code must be clean
-% \paragraph{Raw code is an indicator of quality}
+Within software development best-practices, 
+code reviews are recommended as well \cite{wilson2014best}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Code must be extensive}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For code to be ideal paradata, it must be extensive.
+As code has many properties, there are many recommendations on this aspect.
+
+\begin{itemize}
+  \item Code must be clean
+  \item Code must be hosted
+  \begin{itemize}
+    \item Hosted code generated important metadata
+    \item Code must be tested
+    \item ...
+  \end{itemize}
+  \item Code must be tested
+  \begin{itemize}
+    \item ...
+  \end{itemize}
+\end{itemize}
+
+\paragraph{Code must be clean}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Code must be computer friendly}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-There are multiple ways of publishing code,
-that require increasingly more knowledge to do and result
-in increasingly more reproducible experiments:
-(1) publishing the code as text only, e.g. by printing
-it in a Supplementary Materials, (2) publishing
-the code on a hosted website, e.g on GitHub, with
-added features (3) publishing 
-the installed code within a virtual environment,
-for example, using a Singularity container.
-In this chapter, details of each of these ways will be described,
-including advantages, disadvantages, as well as how to
-make the code more FAIR data \cite{wilkinson2016fair}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Publishing raw code}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{Raw code is already better than no code}
-
-Publishing raw code is already better than not publishing code at all,
-as that raw code can reasonably be assumed to have run successfully at least once.
-To reproduce an experiment from raw code can be painful, 
-yet it is usually better than writing the same code from scratch.
-In general, it appears that sharing code is 
-rare \cite{stodden2011trust,read2015sizing},
-with a pleasant exception being \cite{conesa2019making}.
-
-% \paragraph{Raw code can be FAIR data}
-
-It is hard to make raw code be FAIR data. It is hard for computers to find
-copy-pasted code in a Supplementary Materials. 
-That code is usually only accessible
-if the paper has been published under an open access license.
-When code can be found and accessed, it is an open question if the
-code can \emph{still} interoperate with current datasets, software
-and hardware. Raw code usually lacks the metadata to describe its needs.
-One could say that raw code, albeit better than no code, fits
-FAIR data only poorly.
+\paragraph{Code must be hosted}
 
@@
 
-% \paragraph{Example of hosted code}
+\paragraph{Example of hosted code}
 
@@
 
-% \paragraph{Hosted code has important metadata}
+\paragraph{Hosted code has important metadata}
 
@@
 
-% \paragraph{Hosted code has a commit history}
+\paragraph{Hosted code keeps a history of changes}
 
@@
 
-% \paragraph{A commit history helps assess honest reporting}
+\paragraph{A commit history helps assess honest reporting}
 
@@
 
-% \paragraph{Continuous integration helps assess code quality}
+\paragraph{Code must be tested}
+
+\paragraph{Continuous integration helps assess code quality}
 
@@
 \fi
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Code must be computer friendly}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+There are multiple ways of publishing code,
+that require increasingly more knowledge to do and result
+in increasingly more reproducible experiments:
+(1) publishing the code as text only, e.g. by printing
+it in a Supplementary Materials, (2) publishing
+the code on a hosted website, e.g on GitHub, with
+added features (3) publishing 
+the installed code within a virtual environment,
+for example, using a Singularity container.
+In this chapter, details of each of these ways will be described,
+including advantages, disadvantages, as well as how to
+make the code more FAIR data \cite{wilkinson2016fair}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Publishing raw code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Code is paradata, as it directly describes how the results are generated.
-The part of a scholarly article 
-that describes the experiment in a natural language is metadata, 
-as it describes indirectly what the code does (or should be doing).
-The code that does the work in a computational experiment should
-always be published together with a scholarly article,
-as, if the paper and code differ, it is the code that generated the results.
+This paper started with some suggestions to 
+make paradata useful for data re(use):
+paradata should be comprehensive, comprehensively documented
+and friendly to computers \cite{huvila2022improving}.
+When applying these general recomendations to code, 
+this list can be phrased more precisely:
+code should be comprehensive in supplying 
+automatically generated metadata (such as commit history and code coverage).
+The documentation should be as extensive as recommended by the 
+software development literature. To be computer friendly,
+code should be able to run on at least one independent entity,
+such as a GitHub Actions server or from a Singularity container.
 

### 2023-02-06 article.tex
It is '\LaTeX'

@@
 This paper has been produced following all recommendations in it:
-it's \latex code is hosted on GitHub 
+it's \LaTeX code is hosted on GitHub 
 at \url{https://github.com/richelbilderbeek/chapter_paradata}.

### 2023-02-06 article.tex
Improve

@@
 
-\paragraph{Code must be reviewed}
+\paragraph{Code must be reviewed to ensure the correct amount of documentation}
 
@@
 
-\paragraph{Code must be clean}
+\subsubsection{Code must be clean}
 
@@
 
-\paragraph{Code must be hosted}
+\subsubsection{Code must be hosted}
 
@@
 
-\paragraph{Code must be tested}
+\subsubsection{Code must be tested}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Publishing raw code}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{The problems with hosted code}
-
-Albeit hosted code is vastly superior over raw code
-in case of reproduciblity and correctness,
-there are some problems with hosted code as well.
-The most important one is that hosted code needs maintanance:
+\paragraph{Publishing a virtual environment}
+
+The most reproducible way of submitting the code of an experiment,
+is by providing the code with all the (software) dependencies 
+it needs installed in a virtual container.
+This is close to the golden standard suggested by 
+\cite{peng2011reproducible} 
+\iffalse
+(see also Figure \ref{fig:peng2011reproducible})
+\fi
+,
+however, that article was written before virtualization existed.
+Virtualization allows one to provide a standard environment,
+that can be used on any other computer.
+The best known examples of container software are Docker and Singularity.
+Due to access rights, Singularity is the only viable option to run
+an experiment on a 
+high performance computing (HPC) environment (e.g. a computer cluster).
+
+\paragraph{Containers allow for runnable code, for longer periods of time}
+
+Unlike English, code is fragile in standing the test of time.
+Containers can alleviate this, as these can be created at a point
+in time and then run later. Because the container is stand-alone,
+all software that the code depends on remain in their current/old state.
+As with hosted code, an experiment may not run with newer versions
+of libraries anymore, but a containerized experiment can run regardless, 
+if one refrains from updating (or only does so when there are there problems with fixing).
+
+Ideally, to re-do an experiment, one would need a one-liner to a container,
+for example, \verb|singularity run my_experiment.sif run|.
+
+\paragraph{Runnable code can precisely reproduce an experiment}
+
+Containers allow a computation experiment to be highly reproducible:
+given the same data, an experiment put into a container will give
+the same results on different platforms, at least in theory.
+In practice, differences may be observed when peripheral factors
+are different, such as the random numbers as generated by an operating
+system, or data that are downloaded from additional online sources.
+
+\paragraph{Runnable code allows others to re-use an experiment}
+
+Containers allow a computation experiment to be highly re-usable,
+as any scientist can work with it and tweak code.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Making code FAIR data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Suggestions for knowledge management}
+
+Here are multiple suggestions to make containers with runnable code FAIR.
+To make such containers findable, the most common 
+environments (Syslabs and DockerHub) to upload containers must be indexed.
+Uploading a container to those environments is, however, limited by
+a restricted amount of free storage, with the possible consequence that
+those containers are uploaded somewhere where they escape proper indexing.
+Having containers uploaded at the existing websites ensures they are
+accessible, interoperable and reusable.
+One could argue however, that the scientific community need its own,
+similar website to upload containers, as the metadata for these
+sites are limited to free-text comments and tags (DockerHub)
+or not metadata at all (Syslabs).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sensitive data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For a computational experiment to be reproducible,
+an independent person should be able to
+reproduce the same results.
+The reproduce the same results, one needs the same
+input data and the same code (and the same 
+dependencies and environment).
+When the input data are sensitive, however, and thus cannot be shared with
+an independent researcher, one thus cannot reproduce the
+experiment.
+
+The solution for this problem is simple:
+supply a simulated dataset with the code.
+In the case of genetic epidemiology this would mean
+simulated genotypes and associated phenotypes,
+as can be done with the \verb|plinkr| R package \cite{plinkr}.
+This solution, however, is known to all developer that follow
+good practices when writing code, as public (i.e. simulated) 
+datasets are needed for automatic testing of code.
+One extra benefit of simulated data is that these can be used
+as a benchmark, as slightly different analyses should give 
+similar conclusions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In a perfect world, all code has the characteristis of ideal paradata
+and is written from software development best practices.
+This section discusses the problems that arise by doing so.
+
+\paragraph{Need more training}
+
+\paragraph{Code needs maintanance}
+
+Code needs maintanance:
 the libraries that are used by an experiment
@@
 
+\paragraph{Authors can be contacted}
+
 There is another problem that authors of academic code face,
@@
 (see \cite{barnes2010publish} for a better way to deal with the problem).
+
+\paragraph{How to handle bugs?}
+
 However, upon the find of a new bug, the question is:
@@
 
-% \paragraph{Suggestions for knowledge management}
-
-Here are multiple suggestions to make hosted code FAIR.
-To make the hosted code findable, commonly a paper refers to its URL.
-At the homepage of hosted code (e.g. a GitHub repository), 
-there is no standardized metadata, to,
-for example, link to the paper.
-Using a popular code host ensures the code is easily accessible.
-There is some interoperabily supplied by a version control manager,
-such as in obtaining the history of the code.
-The reusability of code is reasonably easy to assess, 
-by using so-called build badges.
-However, there is no guarantee that a build badge that claimes success,
-actually is an honest signal (i.e. that code may not actually build),
-nor are there standards for what 'success' entails.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Publishing a virtual environment}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The most reproducible way of submitting the code of an experiment,
-is by providing the code with all the (software) dependencies 
-it needs installed in a virtual container.
-This is close to the golden standard suggested by 
-\cite{peng2011reproducible} 
-\iffalse
-(see also Figure \ref{fig:peng2011reproducible})
-\fi
-,
-however, that article was written before virtualization existed.
-Virtualization allows one to provide a standard environment,
-that can be used on any other computer.
-The best known examples of container software are Docker and Singularity.
-Due to access rights, Singularity is the only viable option to run
-an experiment on a 
-high performance computing (HPC) environment (e.g. a computer cluster).
-
-% \paragraph{Containers allow for runnable code, for longer periods of time}
-
-Unlike English, code is fragile in standing the test of time.
-Containers can alleviate this, as these can be created at a point
-in time and then run later. Because the container is stand-alone,
-all software that the code depends on remain in their current/old state.
-As with hosted code, an experiment may not run with newer versions
-of libraries anymore, but a containerized experiment can run regardless, 
-if one refrains from updating (or only does so when there are there problems with fixing).
-
-Ideally, to re-do an experiment, one would need a one-liner to a container,
-for example, \verb|singularity run my_experiment.sif run|.
-
-% \paragraph{Runnable code can precisely reproduce an experiment}
-
-Containers allow a computation experiment to be highly reproducible:
-given the same data, an experiment put into a container will give
-the same results on different platforms, at least in theory.
-In practice, differences may be observed when peripheral factors
-are different, such as the random numbers as generated by an operating
-system, or data that are downloaded from additional online sources.
-
-% \paragraph{Runnable code allows others to re-use an experiment}
-
-Containers allow a computation experiment to be highly re-usable,
-as any scientist can work with it and tweak code.
-
-% \paragraph{The problems with runnable code}
+\paragraph{The problems with runnable code}
 
@@
 Also, journals do not request hosted code, nor runnable code.
-
-% \paragraph{Suggestions for knowledge management}
-
-Here are multiple suggestions to make containers with runnable code FAIR.
-To make such containers findable, the most common 
-environments (Syslabs and DockerHub) to upload containers must be indexed.
-Uploading a container to those environments is, however, limited by
-a restricted amount of free storage, with the possible consequence that
-those containers are uploaded somewhere where they escape proper indexing.
-Having containers uploaded at the existing websites ensures they are
-accessible, interoperable and reusable.
-One could argue however, that the scientific community need its own,
-similar website to upload containers, as the metadata for these
-sites are limited to free-text comments and tags (DockerHub)
-or not metadata at all (Syslabs).
-
-% \paragraph{Runnable code is the pinnacle of reproducible research}
-
-When research truly needs to be reproducible, putting the code 
-of an experiment into a container is a good solution, as it 
-can stand the test of time better.
-Creating such a container, however, requires more skill
-that -as of today- is not rewarded,
-partly as containers follow FAIR data principles only poorly.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sensitive data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For a computational experiment to be reproducible,
-an independent person should be able to
-reproduce the same results.
-The reproduce the same results, one needs the same
-input data and the same code (and the same 
-dependencies and environment).
-When the input data are sensitive, however, and thus cannot be shared with
-an independent researcher, one thus cannot reproduce the
-experiment.
-
-The solution for this problem is simple:
-supply a simulated dataset with the code.
-In the case of genetic epidemiology this would mean
-simulated genotypes and associated phenotypes,
-as can be done with the \verb|plinkr| R package \cite{plinkr}.
-This solution, however, is known to all developer that follow
-good practices when writing code, as public (i.e. simulated) 
-datasets are needed for automatic testing of code.
-One extra benefit of simulated data is that these can be used
-as a benchmark, as slightly different analyses should give 
-similar conclusions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+% \paragraph{Runnable code is the pinnacle of reproducible research}
+
+When research truly needs to be reproducible, putting the code 
+of an experiment into a container is a good solution, as it 
+can stand the test of time better.
+Creating such a container, however, requires more skill
+that -as of today- is not rewarded,
+partly as containers follow FAIR data principles only poorly.
+
 % \paragraph{Final word}

### 2023-02-06 article.tex
EOD

@@
 
-\begin{itemize}
-  \item Code must be clean
-  \item Code must be hosted
-  \begin{itemize}
-    \item Hosted code generated important metadata
-    \item Code must be tested
-    \item ...
-  \end{itemize}
-  \item Code must be tested
-  \begin{itemize}
-    \item ...
-  \end{itemize}
-\end{itemize}
-
 \subsubsection{Code must be clean}
 
-Raw code gives an indication of the quality 
+Code gives an indication of the quality 
 of a computational experiment and several metrics are
 devised to get an idea of the quality of code.
+
+\paragraph{SLOC}
 
@@
 improves software quality \cite{fang2001}.
+
+\paragraph{Cyclomatic complexity}
 
@@
 \cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}
+
+\paragraph{Error handling}
 
@@
 
-% \paragraph{Code must be hosted to be moderately computer-friendly}
-Code is sometimes published alongside a paper, 
-in the form of a code hosting website.
-This way of publishing code gives additional options,
-both for the authors of that code, as well as its users.
-This practice has increased,
+There exist website that are dedicated to host code.
+Such so-called code hosting websites have many advantages,
+both for the authors and users of the code,
+over copy-pasting code in the supplementary materials in a paper.
+
+The practice of using a code hosting websote has increased,
 with the abstracts from articles in the academic journal
@@
 
-\paragraph{Example of hosted code}
+\paragraph{Example of code hosting website}
 
@@
 
-% \paragraph{Runnable code is the pinnacle of reproducible research}
+\paragraph{First step: publish code alongside papers}
+
+\paragraph{Final goal: runnable code}
 
@@
 partly as containers follow FAIR data principles only poorly.
+Runnable code is the pinnacle of reproducible research.
 

### 2023-02-07 article.tex
Improve

@@
 
-The practice of using a code hosting websote has increased,
+The practice of using a code hosting website has increased,
 with the abstracts from articles in the academic journal
@@
 It is a general recommendation to put version control
-on all human-produced data \cite{wilson2014best}.
+on all human-produced data \cite{wilson2014best},
+as well as openly working on the code from the start \cite{jimenez2017four}.
 
@@
 
-% \paragraph{A reader can contact the author}
-
-A feature that gets too few attention is that code hosts
-allow users to post bug reports and/or ask questions.
-This allows other researchers to assess the health
-of the code.
-A healthy project has few open bug reports
-and these bug reports should be closed within reasonable time.
-When bug reports start to accumulate, this
-signals either that the code contained many bugs or that
-the developers lost interest in maintaining that code.
-The amount of bug reports is useful metadata, 
-especially when deciding to use (and depend) on a piece of software.
+\paragraph{Adding a license}
+
+Code hosting websites such as GitHub help developers to follow
+best practices. One of these is to add a software license \cite{jimenez2017four},
+so that it is clear that the software can be reused.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Making code FAIR data}
+\section{Preserving code}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Code is harder to preserve than an English text
+and preserving code is rarely done \cite{barnes2010publish}.
+Although code is the primary actor in computational experiments,
+there is no incentive to submit code alongside a publication.
+Most academic journal do not require authors to submit their code,
+nor is the submitted code peer reviewed, 
+rOpenSci \cite{ram2018community} being the pleasant exception.
+
+Code used in academic research complies badly with FAIR principles,
+where unpublished code is the worst offender.
+However, any idealistic researcher does not have the
+tools to following FAIR principles in an exemplary way:
+there are no standards in how a scholarly article
+refers to academic code (i.e. in a machine-friendly way).
+Creating a (Singularity) container is the best a researcher
+can do to ensure his/her code is reusable,
+yet, there are no incentives to make code accessible, interoperable
+or re-usable, with standarized metadata attributes being absent.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 similar conclusions.
+
+For research to be reproducible, one ideally has access to
+both the data used and the code.
+In some fields, such as genetic epidemiology, the data are
+sensitive, hence cannot be released,
+yet there are methods being devised to run code on sensitive
+data with assured privacy \cite{zhang2016review,azencott2018machine}.
+Additionally, it follows software development's best practices 
+(especially the practice to test code automatically) 
+to release a simulated/public dataset, with
+the additional benefit that this dataset can be used for comparisons.
 
@@
 Also, journals do not request hosted code, nor runnable code.
-
-% \paragraph{Definition of paradata}
-
-% Reviewer 1: this section starts with a sentence that appears to undercut the whole theme of this chapter!
-Code may or may not be paradata, depending on how the definition
-is interpreted.
-Here we repeat the definition of 'paradata' and discuss 
-its constituents.
-This paper defined paradata as 'data about the collecting of the data'.
-This implies that (1) code must be seen as data.
-This paper argues, that code is data in the form of text spread
-over one or more files with useful measurable properties.
-(2) downloading raw data and doing calculations must be seen as collecting.
-This paper argues that downloading raw datas and doing calculations, 
-such as a T-test,
-does describe how the bits and pieces of an end result are collected.
-(3) the results of an experiment must be seen as data.
-This paper argues that an experimental results is data, 
-as it can be measured and can be used as the raw data of a next experiment.
 
@@
 
-Code has additional useful information, similar to confidence intervals,
-that allow a reader to gauge how much one trusts the results.
-The more of the best practices are followed, 
-the more trustworthy are the results.
-The most important best practices discussed in this paper are
-automatic testing, having a high code coverage and code
-of low (cyclomatic) complexity.
-In academia, to uncover the truth, code correctness is essential,
-similar to cell biologists working sterile to not contaminate their
-cell cultures.
-These best practices are vital for a computational biologist.
-
-For research to be reproducible, one ideally has access to
-both the data used and the code.
-In some fields, such as genetic epidemiology, the data are
-sensitive, hence cannot be released,
-yet there are methods being devised to run code on sensitive
-data with assured privacy \cite{zhang2016review,azencott2018machine}.
-Additionally, it follows software development's best practices 
-(especially the practice to test code automatically) 
-to release a simulated/public dataset, with
-the additional benefit that this dataset can be used for comparisons.
-
-Code is harder to preserve than an English text
-and preserving code is rarely done \cite{barnes2010publish}.
-Although code is the primary actor in computational experiments,
-there is no incentive to submit code alongside a publication.
-Most academic journal do not require authors to submit their code,
-nor is the submitted code peer reviewed, 
-rOpenSci \cite{ram2018community} being the pleasant exception.
-
-Code used in academic research complies badly with FAIR principles,
-where unpublished code is the worst offender.
-However, any idealistic researcher does not have the
-tools to following FAIR principles in an exemplary way:
-there are no standards in how a scholarly article
-refers to academic code (i.e. in a machine-friendly way).
-Creating a (Singularity) container is the best a researcher
-can do to ensure his/her code is reusable,
-yet, there are no incentives to make code accessible, interoperable
-or re-usable, with standarized metadata attributes being absent.
-
-\paragraph{First step: publish code alongside papers}
-
 \paragraph{Final goal: runnable code}
@@
 Runnable code is the pinnacle of reproducible research.
+
+\paragraph{First step: publish code alongside papers}
 

### 2023-02-07 article.tex
Improvements

@@
 
+Code as paradata has multiple uses:
+(1) to reproduce an experiment, by running the unchanged code on the same data as
+used in a published research,
+(2) to replicate an experiment, by running the unchanged code on different data as
+used in a published research,
+(3) to extend an experiment, by changing the code.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsubsection{Code must be clean}
-
-Code gives an indication of the quality 
-of a computational experiment and several metrics are
-devised to get an idea of the quality of code.
-
-\paragraph{SLOC}
-
-The simplest metric to assess the quality of code is 
-the amount of (single) lines of code (SLOC). 
-Code that has a lower SLOC count, i.e. it is more concise,
-is usually regarded as having a lower chance of containing bugs.
-Additionally, following a style guide 
+\paragraph{Coding style}
+
+Code is just 'another kind of data' and should be designed 
+as such, for example, by using tools to work on it \cite{wilson2022twelve}.
+A first example is to use a tool to enforce a coding style 
 (e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
-or PEP 8 \cite{van2001pep} for Python), 
-improves software quality \cite{fang2001}.
-
-\paragraph{Cyclomatic complexity}
-
-A more interesting metric is the cyclomatic complexity of the code.
+or PEP 8 \cite{van2001pep} for Python).
+Following a consistent coding style improves software quality \cite{fang2001}.
+A second example is to use a tool to enforce a low cyclomatic complexity.
 The cyclomatic complexity is approximately defined 

### 2023-02-07 article.tex
Improve

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Making code useful paradata}
+\section{Making code useful paradata}\label{sec:making-code-useful-paradata}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Code as paradata has multiple uses:
+Code as paradata has multiple uses: to reproduce, to replicate and to 
+extend \cite{benureau2018re}.
+
+To reproduce an experiment is to obtain the same publish result,
+by using the same code and data as used for that publication 
+\cite{peng2006reproducible}.
+
 (1) to reproduce an experiment, by running the unchanged code on the same data as
@@
 
-For code to be ideal paradata, it must be documented.
-However, code does explain itself to some extent,
-for example, the function name \verb|t.test| as used above.
-
-\begin{itemize}
-  \item Code must be documented to the right amount
-  \item Code must be reviewed
-\end{itemize}
-
-\paragraph{Code must be documented to the right amount}
-
-Code should be documented to the right amount:
-too little documentation is unhelpful, where too much is a needless burden
-on the programmer.
-The first recommended practice is to write the code in such a way 
-that it becomes self-explanatory \cite{wilson2014best}.
-What is left to document are the reasons behind the code, 
+For code to be ideal paradata, it must be extensively documented.
+This extend is determined by its intended use: 
+to reproduce, replicate or extend \ref{sec:making-code-useful-paradata}.
+The amount of documentation is, however, also determined
+by the recommended practices for writing code in general.
+
+\subsubsection{... for reproducing}
+
+For reproducing purposes, the minimal amount of documentation
+is determined by software development best practices.
+When an independent researcher reproduces an experiment, 
+it is possible that no code is read at all.
+Due to this, one could argue that no documentation is needed.
+However, all code should be documented 'adequately' \cite{peng2006reproducible},
+ideally writing code in such a way that it becomes self-explanatory \cite{wilson2014best}
+and for the remaining code to document the reasons behind it, 
 its design and its purpose \cite{wilson2014best}.
+
+\subsubsection{... for replication}
+
+\subsubsection{... for extending}
 
@@
 
-There exist website that are dedicated to host code.
-Such so-called code hosting websites have many advantages,
-both for the authors and users of the code,
-over copy-pasting code in the supplementary materials in a paper.
-
-The practice of using a code hosting website has increased,
-with the abstracts from articles in the academic journal
-'Bioinformatics' referring to such a hosted website
-has increased tenfold, going from below 5 percent in 2009 
-to around 50 percent in 2017 \cite{russell2018large}.
-
-\paragraph{Example of code hosting website}
-
-There are multiple websites for hosting code, with BitBucket, GitHub,
-GitLab and SourceForge being the most well-known.
-GitHub is currenly the most popular website to host code in general,
-with millions of users (a simple user search can confirm this).
-GitHub users can create dedicated websites (called 'repositories')
-to upload code. Code hosts, such as GitHub, 
-store that uploaded code and provide ways 
-for visiters to download the code and/or interact with it.
+Code should be distributed in standard ways \cite{peng2006reproducible}.
+The standard way to publish code is to use a code hosting website.
+There are multiple code hosting websites, 
+with GitHub being the most popular one \cite{cosentino2017systematic}.
+GitHub is a website where users can create dedicated pages (called 'repositories')
+to upload code and interact with it.
 The use of such code hosting websites
-accommodates collaboration \cite{perez2016ten}.
+has increased strongly \cite{russell2018large},
+accommodates collaboration \cite{perez2016ten}
 and improves transparency \cite{gorgolewski2016practical}.
+
+\iffalse
 An example from the author is the website \cite{bbbqarticleissue157},
-that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}
-\iffalse
-, as shown in Figure \ref{fig:bbbqarticleissue157}
-\fi
-.
+that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
+as shown in Figure \ref{fig:bbbqarticleissue157}.
+\fi
 
@@
 
-For a computational experiment to be reproducible,
-an independent person should be able to
-reproduce the same results.
-The reproduce the same results, one needs the same
-input data and the same code (and the same 
-dependencies and environment).
+Next to the code, it is the data used in an experiment 
+that must be made available for an experiment to be called 'reproducible' \cite{peng2006reproducible}.
 When the input data are sensitive, however, and thus cannot be shared with
@@
 The solution for this problem is simple:
-supply a simulated dataset with the code.
+supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
+together with the code.
 In the case of genetic epidemiology this would mean
@@
 with multiple implementations \cite{hinsen2019dealing},
-this paper recommends a modern solution that potentially has more longevity:
+this paper recommends a modern solution that potentially has more 
+longevity:
 to publish the code installed in a virtual environment
 (including the dependencies used at that time), as described below.
+Write code that will stand the test of time perfectly
+is deemed 'impossible' \cite{benureau2018re}.
 

### 2023-02-07 article.tex
Feedback
Progress

@@
 
-% \paragraph{Simple example}
+\paragraph{Simple example}
 
@@
 
-% \paragraph{The first definition paradata}
+\paragraph{The first use of paradata}
 
@@
 
-% \paragraph{Definition of paradata}
+\paragraph{Definition of paradata}
 
@@
 
-% Code is paradata
+\paragraph{Code is paradata}
 
@@
 
-% \paragraph{Modern forms of paradata}
+\paragraph{Modern forms of paradata}
 
@@
 make paradata useful for data re(use):
-paradata should be comprehensive, comprehensively documented
-and friendly to computers \cite{huvila2022improving}.
-These suggestions appear to work with code as well,
-where these suggestions show up as best practices (such
-as \cite{stodden2013best}), some of which
-described below.
+paradata should be comprehensive, documented in a useful 
+way (and shaped by the data) and be friendly to computers \cite{huvila2022improving}.
+These suggestions appear to work with code as well.
 The lack of computer code is just as much 'a drastic constraint'
 in the use of a (computational) scientific experiment.
-The clearest effect of its absence is, most obviously,
-that a computational experiment cannot be redone 
-easily, yet there are wider implications, described below, as well.
-
-% \paragraph{This paper is about computer code}
-
-This paper is about paradata in the context of a computational
-experiment in general and uses an example from genetic 
-epidemiology in particular.
-It is argued that code, being paradata, is as vital as the academic
-paper itself, illustrated by the two scientists at the start of 
-this chapter. After that, the implications and recommendations
-follow.
+The reasons why paradata in the form of code matters, 
+are the same reaons as why paradata matters,
+which are data reuse, research reproducibility, 
+contextual understanding of disciplinary differences,
+Open Science and insights into 
+(scientific) knowledge production \cite{huvila2022improving}.
+This chapter will focus mostly on the goals of making science
+reproducible and code reusable.
 
@@
 
+% d: Documentation as a Factor for Project Popularity
+%
+% There exists a clear relationship between the popularity and the documentation effort [62], [64], [65]. 
+% cosentino2017systematic
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and improves transparency \cite{gorgolewski2016practical}.
+See \cite{cosentino2017systematic} for an extensive overview of
+research conducted on GitHub.
 
@@
 Next to the code, it is the data used in an experiment 
-that must be made available for an experiment to be called 'reproducible' \cite{peng2006reproducible}.
+that must be made available for an experiment 
+to be called 'reproducible' \cite{peng2006reproducible}.
 When the input data are sensitive, however, and thus cannot be shared with

### 2023-02-07 article.tex
Improve

@@
 
-\paragraph{Modern forms of paradata}
+\paragraph{Making paradata useful}
 
@@
 way (and shaped by the data) and be friendly to computers \cite{huvila2022improving}.
-These suggestions appear to work with code as well.
-The lack of computer code is just as much 'a drastic constraint'
-in the use of a (computational) scientific experiment.
-The reasons why paradata in the form of code matters, 
-are the same reaons as why paradata matters,
-which are data reuse, research reproducibility, 
+These suggestions appear to work well for code too
+and will be discussed in detail below.
+
+\paragraph{The goal of good paradata}
+
+The reasons why useful paradata matters,
+are data reuse, research reproducibility, 
 contextual understanding of disciplinary differences,
@@
 (scientific) knowledge production \cite{huvila2022improving}.
-This chapter will focus mostly on the goals of making science
-reproducible and code reusable.
+This chapter discusses these reasons below.
 
@@
     Left: general relation between data, paradata and metadata.
-    Middle: the same relations specified for eBird.
     Right: the same relations specified for genetic epidemiology.
@@
 
-% \paragraph{Introduction}
+\paragraph{Introduction}
 
@@
 
-% \paragraph{Publication of code}
+\paragraph{How commonly code is published}\label{para:how-commonly-code-is-published}
 
@@
 to redo the analysis \cite{obels2020analysis}.
-
-% \paragraph{For some fields, the experiment is actually run by code}
+Unpublished code has been the cause of some saddening examples,
+such as an algorithm that detects breast cancer from images 
+better than a human expert \cite{haibe2020importance}.
+
+\paragraph{For some fields, the experiment is actually run by code}
 
@@
 
-% \paragraph{The experiments within genetic epidemology works are done by code}
+\paragraph{The experiments within genetic epidemology works are done by code}
 
@@
 
+\paragraph{Code is vital}
+
 The code for this experiment is vital paradata.
-Like all paradata, its omission would be a constraint in the
-use of the conclusion reached by it:
-we can use it to assess the trustworthiness of the finding,
-for example, by running the code again.
-Ideally, the code is comprehensive, comprehensively documented
-and be friendly to computers \cite{huvila2022improving}.
-Code being friendly to computer seems like an obvious 'yes',
-yet this is naive, as commonly it is hard to run code in practice. 
-% TODO: Link to correct paragraphs
-For code to be called computer-friendly, it must be run on an independent entity, 
-as is discussed in the paragraphs about hosting code and/or putting the code in a container.
-% TODO: Link to correct paragraphs
-Also note that there is a problem due to the inherent sensitivity of the genetic data,
-which is discussed in paragraph.
-Lastly, note that the researcher used data that was itself generated by code:
-complex bioinformatics pipelines are used to obtain reliable DNA sequences 
-from the biological sample.
-
-% \paragraph{Science should be reproducible}
-
-A scientific claim must be reproducible before it is accepted
-by the scientific community.
+Ideally, the code is useful paradata, 
+as discussed in section \label{sec:making-code-useful-paradata}.
+
+\paragraph{The goal of useful paradata is reproducible science}
+
+The reason behind paradata is, among others,
+to assure reproducible research \cite{huvila2022improving}.
 Reproducibility (i.e. to reproduce the same results) 
@@
 are the exception.
-
-That same study recommended, as a solution to counter this, 
-to make reproducible research a minimal requirement for 
-publication \cite{peng2011reproducible}.
-
-% \paragraph{Code should be published}
+To counter the lack of published code, 
+it has been suggested to make the reproduction of research 
+a minimal requirement for publication \cite{peng2011reproducible}.
+
+\paragraph{Sensitive data}
+
+A genetic epidemiology works with sensitive data as well:
+the genetic sequences of participants are private.
+For research to be reproducible one needs both the code and the data
+to reproduce the (hopefully) same results.
+This problem is discussed in section \ref{sec:sensitive-data}.
+
+\paragraph{Code holds the ground truth}
 
@@
 \fi
-
-The code of a computational experiment must be published
-for an experiment to be easily reproducible.
-It is however (too) common that the original data
-and/or code are unavailable \cite{peng2021reproducible}.
-One saddening example is described in \cite{haibe2020importance}, where
-an algorithm that detects breast cancer from images better than a human expert,
-was irreproducible.
-In the case of genetic epidemiology, 
-the original data usually cannot be published, 
-due to the guaranteed privacy of the people having shared their characteristics.
-The code of a computational experiment, however, \emph{can} be
-distributed without such problems.
-In academic journals, however, its is not code that is published, 
-but an English description of what it does instead.
-
-% \paragraph{Example of code being the ground truth}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sensitive data}
+\section{Sensitive data}\label{sec:sensitive-data}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-02-07 article.tex
Progress

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Code must be extensively documented}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For code to be ideal paradata, it must be extensively documented.
-This extend is determined by its intended use: 
+\subsection{Code must be usefully documented}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For code to be ideal paradata, it must be usefully documented.
+What is useful is determined by its intended use: 
 to reproduce, replicate or extend \ref{sec:making-code-useful-paradata}.
@@
 However, all code should be documented 'adequately' \cite{peng2006reproducible},
-ideally writing code in such a way that it becomes self-explanatory \cite{wilson2014best}
+ideally writing code in such a way that it becomes 
+self-explanatory \cite{wilson2014best}
 and for the remaining code to document the reasons behind it, 
@@
 
+There exists a clear relationship between the popularity 
+and the documentation effort \cite{cosentino2017systematic,hata2015characteristics}.
+
 \paragraph{Code must be reviewed to ensure the correct amount of documentation}
 
-As it is the code that is the most important actor in an the experiment,
-it should be treated as equally or more important than the scientific
-paper describing what is does.
-Regardless of this, most academic journals, 
-when concerning computational experiments
-only review the (usually) English description of the code
-and make the review of code optional.
+More than half of all scientists have their code 
+rarely or never reviewed \cite{vable2021code},
+although code reviews are known to
+accelerate learning of the developers,
+improve the quality of the code
+and resulting to an experiment that is likelier to be reproducible
+\cite{vable2021code}.
 One notable exception is rOpenSci \cite{ram2018community},
@@
 code reviews are recommended as well \cite{wilson2014best}
-
-% d: Documentation as a Factor for Project Popularity
-%
-% There exists a clear relationship between the popularity and the documentation effort [62], [64], [65]. 
-% cosentino2017systematic
 
@@
 
+Coding errors are extremely common and contribute to the reproducibility crisis in science \cite{vable2021code}
+
+%.
+%. How to make more published research true
+%. NIH plans to enhance reproducibility
+%. The replication crisis in epidemiology: snowball, snow job, or winter solstice?
+%. The harm done to reproducibility by the culture of null hypothesis significance testing
+
 \paragraph{Continuous integration helps assess code quality}
@@
 
-\paragraph{Need more training}
+\paragraph{Invest in training}
+
+One needs to be trained to know the best practices,
+before these can be applied. 
+Articles that suggest these best practices, however, claim that this initial
+investment pays off.
+Code reviews are a good way to accelerate the 
+learning of team members \cite{vable2021code}.
 

### 2023-02-08 article.tex
Rewrote abstract and last paragraph

@@
 
-Here we define paradata as data about the data collection process.
-Computer code is the paradata for a computatioal/academic experiment,
-allowing the experiment to be reproduced.
-However, code is published alongside an academic
-paper in a minority of all cases.
-As reproducibility is a cornerstone of science,
-research code should be published.
-Earlier general recommendations about making paradata (more) useful
-is tailored to code, using the field of genetic epidemiology
+Paradata is data about the data collection process,
+that allows use and reuse of data.
+Within the context of computational research,
+computer code is the paradata of an experiment,
+allowing the study to be reproduced.
+A recent study recommended how to make paradata (more) useful,
+for paradata in general.
+This study applies those recommendations to computer code,
+using the field of genetic epidemiology
 as an example.
@@
 and helpful would the code that accompanies a scientific paper
-be treated like a first class citizen. As doing so in an exemplary way
-is yet to be rewarded, hence it has to be the idealististic scientists
-to wage this battle. I feel the truth and science are worth fighting for
-and I hope this paper helps others to join.
+be treated like a first class citizen. 
+Doing so, however, is yet to be rewarded
+and still both of the two scientists at the start of this paper 
+can provide a good rationale for their behavior.
+This will change when reward incentives are put into place 
+that reward making paradata useful.
+For code specifically, in any computational field,
+the rewards are even higher, as reproducibility should again be 
+a cornerstone in science.
 

### 2023-02-08 article.tex
Add response rates

@@
 
-The code used in experiments is just that, 'data about the data collection 
+The code used in computational experiments is just that, 
+'data about the data collection 
 process', as it commonly downloads data, selects relevant subsets in those
@@
 
-The reasons why useful paradata matters,
-are data reuse, research reproducibility, 
-contextual understanding of disciplinary differences,
-Open Science and insights into 
-(scientific) knowledge production \cite{huvila2022improving}.
-This chapter discusses these reasons below.
+There are multiple reasons why useful paradata matter. 
+Most obvious is that having useful paradata gives an understanding how
+data is produced.
+This insight helps researchers from different fields to understand each other
+and collaborate.
+Additionally, useful open data is needed for Open Science to convincingly show
+its benefits.
+Finally, in computational fields, 
+it can help understand how scholarly knowledge is produced
+\cite{huvila2022improving}.
 
@@
 to illustrate that the code is paradata,
-that this type of paradata is relevant, as it helps 
-to assess the correctness of the results,
-where the correctness of the results is relevant for 
-science in general and healthcare in particular.
+and that this type of paradata is relevant.
 However, any field that uses computation in its experiments
@@
 such as an algorithm that detects breast cancer from images 
-better than a human expert \cite{haibe2020importance}.
+better than a human expert, 
+yet failed to ever be reproduced \cite{haibe2020importance}.
 
@@
 
-Genetic epidemiology is a field within biology that, among 
-others, measures the spread of heritable traits,
-as well as how these come to be.
+Genetic epidemiology is a field within biology that
+tries to determine the mechanism and spread of heritable traits.
 For example, we know that lactose intolerance in adults is
@@
 and is most commonly found in south-east asia and south africa \cite{storhaug2017country}.
-The trait is caused by the genetic make-up, 
-or genotype 
+The trait is caused by the genetic make-up, or 'genotype'
 (see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
 of a person.
-The trait, also called phenotype (again, see Table \ref{tab:definitions}), 
+The trait, also called 'phenotype' (again, see Table \ref{tab:definitions}), 
 in this example is lactose intolerance at adult age,
 yet any human property, such as weight or height can be studied.
-Genetic epidemiology studies the spread of genotypes and phenotypes,
-as well as the relation between those two.
 When an association between genotype and phenotype is found,
 these associations are subsequently used to 
-create gene panels, where the gene causing 
+create so-called 'gene panels', 
+where the gene causing 
 an association is measured specifically, to, among others,
 detect people at risk for the associated phenotype.
-
 The rest of this section describes a genetic epidemiology study 
@@
 and the phenotype (in this case, the concentration of IL6RA).
+To be more precise, it was code that read the data,
+subsetted the data, removed outliers, 
+performed the statistics and generated the plots.
+All these steps are described in the paper, 
+but the code itself is unpublished
+and the paper does not mention any way to obtain it.
+The reponse rate of correnspondence authors with a request for data is around 50\% 
+\cite{manca2018non, teunis2015corresponding} 
+(note that for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
 

### 2023-02-08 article.tex
Progress

@@
 \cite{manca2018non, teunis2015corresponding} 
-(note that for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
-
-\paragraph{Code is vital}
-
-The code for this experiment is vital paradata.
-Ideally, the code is useful paradata, 
-as discussed in section \label{sec:making-code-useful-paradata}.
+(note that for the field of emergency medicine, 
+these rates are 73\% \cite{o2003email}).
 
@@
 
-Again, paradata should be comprehensive, comprehensively documented
+Again, paradata should be comprehensive, documented in a useful way
 and friendly to computers \cite{huvila2022improving}.
@@
 or -more relevant to the research- sending in a bug report.
-One solution for the author is to ignore such emails
-and it can be argued that no energy should be wasted on published code
+One solution for the author is to ignore such emails (as is done in a third
+of all cases \cite{teunis2015corresponding}):
+it can be argued that no energy should be wasted on published code
 and work on something new instead 

### 2023-02-09 article.tex
Progress

@@
 
-% This section has paragraph headers. This helps me, the author,
-% to focus the contents of my paragraphs.
-% In the final version of this manuscript, these will be removed.
-
 \paragraph{Simple example}
@@
 write two equally good manuscripts
-that involve computation.
+that describe an experiment with computational steps.
 Both manuscripts are accepted by an equally prestigous journal after peer view. 
@@
 Is this difference relevant and worth the effort?
-This chapter discusses the effect of supplying the computer 
-code  
-and claims that code is paradata and that this paradata 
-is just as important as the paper itself.
+This chapter discusses why code, a type of paradata, 
+should be supplied
+and how it should be supplied,
+and claims that code is just as important as the paper itself.
 
@@
 make paradata useful for data re(use):
-paradata should be comprehensive, documented in a useful 
-way (and shaped by the data) and be friendly to computers \cite{huvila2022improving}.
+paradata should be comprehensive, documented in a useful way, 
+the documention and data should have co-evolved, 
+and the paradata should be good to computers \cite{huvila2022improving}.
 These suggestions appear to work well for code too
@@
 and that this type of paradata is relevant.
-However, any field that uses computation in its experiments
+However, any field that uses computation 
+on sensitive data in its experiments,
 could be used as an example.
@@
 a field intimately familiar with computer code,
-only 192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
+192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
 Another study analysed the reproducibility of registered reports
 in the field of psychology, 
-where still only 37 out of 62 studies supplied the code 
+where 37 out of 62 studies supplied the code 
 to redo the analysis \cite{obels2020analysis}.
@@
 Genetic epidemiology is a field within biology that
-tries to determine the mechanism and spread of heritable traits.
+tries to determine the spread of heritable traits 
+and their underlying biological mechanism.
 For example, we know that lactose intolerance in adults is
@@
 The first type of primary data, the genotypes, 
-consists out of single nucleotide polymorphisms (SNPs, pronounced as 'snips').
+consists out of single nucleotide polymorphisms (SNPs, pronounced 'snips').
 A SNP has a name and a location on the DNA, at which there is a certain nucleotide.
@@
 their presence can be used to assess the health of an individual.
-IL6RA is one such protein and, spoiler alert, its concentration
-is associated with the SNP mentioned earlier.
+IL6RA is one such protein and its concentration may (and will, see below)
+be associated with the SNP mentioned earlier.
 
@@
 is associated with a disease that develops later in life,
-yet is preventable by lifestyle changes.
+yet is preventable by lifestyle 
+changes (see \cite{pope2003will} for an example in Alzheimer's disease).
 Would this be the case, we can create a tailored 
@@
 However, there are associations published for many diseases,
-in which one (e.g. phenylketonuria) or many (e.g. \cite{bruce2009metabolic}) 
-SNPs contribute to being more likely to develope a disease in the future. 
+in which one SNP (e.g. phenylketonuria) or many SNPs (e.g. \cite{bruce2009metabolic}) 
+contribute to being more likely to develope a disease in the future. 
 Here, the phenotype (having a disease in the future) 
@@
 performed the statistics and generated the plots.
-All these steps are described in the paper, 
+All these steps are described in the paper (\cite{ahsan2017relative}), 
 but the code itself is unpublished
 and the paper does not mention any way to obtain it.
-The reponse rate of correnspondence authors with a request for data is around 50\% 
+One could send an email to the correspending,
+yet it is known that the reponse rate of correnspondence authors 
+with a request for data in related fields is around 50\% 
 \cite{manca2018non, teunis2015corresponding} 
-(note that for the field of emergency medicine, 
-these rates are 73\% \cite{o2003email}).
+(for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
 
@@
 
-A genetic epidemiology works with sensitive data as well:
-the genetic sequences of participants are private.
+A genetic epidemiologist works with sensitive data as well:
+the genetic sequences of participants are private (\cite{clayton2019law}).
 For research to be reproducible one needs both the code and the data
@@
 
-Again, paradata should be comprehensive, documented in a useful way
-and friendly to computers \cite{huvila2022improving}.
+Again, useful paradata should be comprehensive, 
+documented appropriately, the documentation should co-evolve with the data,
+and the data should be friendly to computers \cite{huvila2022improving}.
 In this section, these ideal properties of paradata are applied to code.
-The resulting recommendations are compared with
-existing best practices in writing code.
-
-Code as paradata has multiple uses: to reproduce, to replicate and to 
-extend \cite{benureau2018re}.
-
-To reproduce an experiment is to obtain the same publish result,
-by using the same code and data as used for that publication 
-\cite{peng2006reproducible}.
-
-(1) to reproduce an experiment, by running the unchanged code on the same data as
-used in a published research,
-(2) to replicate an experiment, by running the unchanged code on different data as
-used in a published research,
-(3) to extend an experiment, by changing the code.
+Code as paradata, however, has multiple uses as well,
+which are to reproduce, to replicate and to 
+extend a computational experiment \cite{benureau2018re}.
+In this section, the recommendations how to make paradata useful
+is applied to code and compared to existing best practices in software 
+development, for the intended use of the code.
 
@@
 
-For code to be ideal paradata, it must be usefully documented.
-What is useful is determined by its intended use: 
+For code to be ideal paradata, it must be usefully documented \cite{huvila2022improving}.
+What is 'useful' is determined by its intended use: 
 to reproduce, replicate or extend \ref{sec:making-code-useful-paradata}.
@@
 
-\subsubsection{... for reproducing}
-
-For reproducing purposes, the minimal amount of documentation
+For purposes of reproducing (same code, same data) or replicating (same code,
+different data) a study, the minimal amount of documentation
 is determined by software development best practices.
-When an independent researcher reproduces an experiment, 
-it is possible that no code is read at all.
+When an independent researcher reproduces or replicates an experiment, 
+it is possible (even ideal!) that no code is read at all.
 Due to this, one could argue that no documentation is needed.
@@
 
-\subsubsection{... for replication}
-
-\subsubsection{... for extending}
-
-There exists a clear relationship between the popularity 
-and the documentation effort \cite{cosentino2017systematic,hata2015characteristics}.
-
-\paragraph{Code must be reviewed to ensure the correct amount of documentation}
-
-More than half of all scientists have their code 
-rarely or never reviewed \cite{vable2021code},
-although code reviews are known to
-accelerate learning of the developers,
-improve the quality of the code
-and resulting to an experiment that is likelier to be reproducible
-\cite{vable2021code}.
-One notable exception is rOpenSci \cite{ram2018community},
-which is an organisation that does peer-review academic code 
-and facilitates the optional publication of a paper.
-Although both code and paper are peer-reviewed,
-most effort is spent on reviewing the code,
-as can be seen in, for example, 
-the \verb|mcbette| R package \cite{bilderbeek2020mcbette}.
-Within software development best-practices, 
-code reviews are recommended as well \cite{wilson2014best}
+For purposes of extending a study and its code, 
+documentation becomes even more important,
+as the code will be read and modified.
+The extent of investing time in documenting code is recommended
+to be proportional to the intended reuse \cite{pianosi2020successfully}
+and there exists a clear relationship between the reuse 
+of code and its documentation 
+effort \cite{cosentino2017systematic,hata2015characteristics}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Code and documentation must align}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For code to be ideal paradata, 
+its creation and documentation need to match, 
+'not least because they shape each other' \cite{huvila2022improving}.
+This quote resonates strongly with the idea of 
+literate programming \cite{knuth1984literate},
+in which documentation and code is developed hand-in-hand.
+rOpenSci, a community that, among others, reviews programming
+code (\cite{ram2013ropensci,ram2018community}), is 
+one example of where literate programming is enforced:
+code must have extensive runnable examples
+as part of the documentation \cite{ropensci_2021_6619350}.
+Developing code and documentation at the same time 
+leads to both better code and documentation \cite{reenskaug1989environment}.
+In general, when developing software, 
+it is recommended to 
+to write documentation while writing software,
+as well as to include many examples \cite{lee2018ten}.
 
@@
 
-\paragraph{Coding style}
+\subsubsection{... to reproduce an experiment}
+
+\paragraph{Code should be hosted}
+
+For code to be extensive enough to reproduce an existing experiment,
+it should be distributed in standard ways \cite{peng2006reproducible}.
+The standard way to publish code is to use a code hosting website.
+There are multiple code hosting websites, 
+with GitHub being the most popular one \cite{cosentino2017systematic}.
+GitHub is a website where users can create dedicated pages (called 'repositories')
+to upload code and interact with it.
+The use of such code hosting websites
+has increased strongly \cite{russell2018large},
+accommodates collaboration \cite{perez2016ten}
+and improves transparency \cite{gorgolewski2016practical}.
+See \cite{cosentino2017systematic} for an extensive overview of
+research conducted on GitHub.
+
+\iffalse
+An example from the author is the website \cite{bbbqarticleissue157},
+that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
+as shown in Figure \ref{fig:bbbqarticleissue157}.
+\fi
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{bbbq_article_issue_157.png}
+  \caption{
+    A typical GitHub repository \cite{bbbqarticleissue157}, 
+    hosting the code for a
+    computational experiment, 
+    as published with \cite{bilderbeek2022transmembrane}.
+    Note that this repository has 72 commits, where a commit is a change
+    to the code.
+  }
+  \label{fig:bbbqarticleissue157}
+\end{figure}
+\fi
+
+\subsubsection{... to replicate an experiment}
+
+\paragraph{Code must have extensive error handling}
+
+Code must be more extensive when it is (intended to be) 
+used on different data,
+as then 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
+Error handling is one of the mechanisms to do so.
+A simple example is when one (e.g. a BSc student) 
+needs to calculate the variance of a distribution.
+Commonly one needs to call a function to do so, 
+e.g. \verb|var| in R, that needs the values of the distribution
+as input.
+However, to calculate the variance of a distribution, at least two values
+are required (else a division by zero occurs in the calcation of the variance). 
+When the variance of (zero or) only one value is requested,
+an error message can helpfully indicate that the function 
+needs at least two values to work as expected.
+
+\subsubsection{... to extend an experiment}
+
+In software development, 
+changes are common and should be embraced \cite{beck2000extreme}.
+When a scientist intends to reuse code by extending its functionality,
+first the code will be assessed,
+after which the code may be actually extended.
+
+\subsubsubsection{Assessment of code}
+
+\subsubsubsection{Extending code}
 
@@
 as such, for example, by using tools to work on it \cite{wilson2022twelve}.
+
+\paragraph{Using a tool for coding style}
+
 A first example is to use a tool to enforce a coding style 
@@
 Following a consistent coding style improves software quality \cite{fang2001}.
+
+\paragraph{Using a tool for complexity}
+
 A second example is to use a tool to enforce a low cyclomatic complexity.
@@
 
-\paragraph{Error handling}
-
-A feature that deserves more attention
-is that 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
-Error handling is one of the mechanisms to do so.
-A simple example is when one (e.g. a BSc student) 
-needs to calculate the variance of a distribution.
-Commonly one needs to call a function to do so, 
-e.g. \verb|var| in R, that needs the values of the distribution
-as input.
-However, to calculate the variance of a distribution, at least two values
-are required (else a division by zero occurs in the calcation of the variance). 
-When the variance of (zero or) only one value is requested,
-an error message can helpfully indicate that the function 
-needs at least two values to work as expected.
-
-\subsubsection{Code must be hosted}
-
-Code should be distributed in standard ways \cite{peng2006reproducible}.
-The standard way to publish code is to use a code hosting website.
-There are multiple code hosting websites, 
-with GitHub being the most popular one \cite{cosentino2017systematic}.
-GitHub is a website where users can create dedicated pages (called 'repositories')
-to upload code and interact with it.
-The use of such code hosting websites
-has increased strongly \cite{russell2018large},
-accommodates collaboration \cite{perez2016ten}
-and improves transparency \cite{gorgolewski2016practical}.
-See \cite{cosentino2017systematic} for an extensive overview of
-research conducted on GitHub.
-
-\iffalse
-An example from the author is the website \cite{bbbqarticleissue157},
-that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
-as shown in Figure \ref{fig:bbbqarticleissue157}.
+\paragraph{Code must be tested}
+
+Coding errors are extremely common 
+and contribute to the reproducibility crisis in science \cite{vable2021code}.
+
+%.
+%. How to make more published research true
+%. NIH plans to enhance reproducibility
+%. The replication crisis in epidemiology: snowball, snow job, or winter solstice?
+%. The harm done to reproducibility by the culture of null hypothesis significance testing
+
+\paragraph{Continuous integration helps assess code quality}
+
+Some code hosts, among others, GitHub, 
+allow the researcher/user to start software scripts upon uploading a change.
+This is useful to determine the reproducibility of the code: 
+one such script should run the code (on trivial data) 
+to verify if the code works at all.
+If the functionality of the code is lost, the researcher
+can immediatly see which small change was the cause of this.
+A formalized version of this workflow is called continuous integration (CI).
+Using a CI service (such as GitHub Actions for GitHub), 
+is known to significantly 
+increase the number of bugs exposed \cite{vasilescu2015} and increases
+the speed at which new features are added \cite{vasilescu2015}.
+Using CI opens up the possibility to informally annotate features of
+a repository. 
+\iffalse
+For example, Figure \ref{fig:build_badge} shows a 
+build badge that indicates that a \LaTeX document (i.e. this paper)
+could be built.
 \fi
@@
   \centering
-  \includegraphics[width=\linewidth]{bbbq_article_issue_157.png}
+  \includegraphics[width=\linewidth]{build_badge.png}
   \caption{
-    A typical GitHub repository \cite{bbbqarticleissue157}, 
-    hosting the code for a
-    computational experiment, 
-    as published with \cite{bilderbeek2022transmembrane}.
-    Note that this repository has 72 commits, where a commit is a change
-    to the code.
-  }
-  \label{fig:bbbqarticleissue157}
+    A build badge for the GitHub repository of this article.
+    The description indicates which process it is that the badge
+    signals success (as in this case) or failure of. 
+  }
+  \label{fig:build_badge}
 \end{figure}
 \fi
+
+\paragraph{Code coverage helps assess code quality}
+
+Testing, in general, is an important mechanism to ensure
+the correctness of code (an interesting example is \cite{rahman2020exploratory}
+showing bugs in scientific software on the COVID-19 pandemic:
+it recommends more testing).
+The percentage of (lines of) code tested is called the code coverage.
+Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
+The non-profit organisation rOpenSci \cite{ram2018community},
+which peer-reviews code,
+has made it a prerequisite to have a code coverage of 100\%.
+Also here, CI opens up the possibility to informally annotate the
+code coverage. 
+For example, build badges
+\iffalse
+(e.g. as shown in Figure \ref{fig:badge_codecov})
+\fi
+are visual indications on a website that can, among others,
+show a certain R package's code has a 100\% code coverage.
+
+Hence, code coverage is important metadata, that can automatically
+be acquired when using code hosting website.
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[]{badge_codecov.png}
+  \caption{
+    A build badge for the code coverage of an R 
+    package (\url{https://github.com/ropensci/beautier}).
+    The umbrella signals that the code coverage is hosted by
+    CodeCov (\url{https://codecov.io/}). 
+    The text indicates which process it is that the badge
+    signals, which in this case, is that the code is fully covered
+    by tests.
+  }
+  \label{fig:badge_codecov}
+\end{figure}
+\fi
+
+\paragraph{Adding a license}
+
+Code hosting websites such as GitHub help developers to follow
+best practices. One of these is to add a software license \cite{jimenez2017four},
+so that it is clear that the software can be reused.
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{issue_ldpred2.png}
+  \caption{
+    An overview of the issues for the code for
+    LDpred2 \cite{prive2020ldpred2}. Of the 310 issues, 300 have been closed.
+    The first four issues have been replied to, as indicated by the
+    talk balloon. The fifth issues has a label with the text 'feature request'
+    to signal the issue type.
+  }
+  \label{fig:issue_ldpred2}
+\end{figure}
+\fi
+
+\paragraph{Test-driven development}
+
+TDD improves code quality \cite{alkaoud2018quality,janzen2006test}
+and writing comment and documentation mixes well \cite{shmerlin2015document}.
+
+\paragraph{Code must be reviewed to ensure the correct amount of documentation}
+
+More than half of all scientists have their code 
+rarely or never reviewed \cite{vable2021code},
+although code reviews are known to
+accelerate learning of the developers,
+improve the quality of the code
+and resulting to an experiment that is likelier to be reproducible
+\cite{vable2021code}.
+Within software development best-practices, 
+code reviews are recommended as well \cite{wilson2014best}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Code must be computer friendly}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+There are multiple ways of publishing code,
+that require increasingly more knowledge to do and result
+in increasingly more reproducible experiments:
+(1) publishing the code as text only, e.g. by printing
+it in a Supplementary Materials, (2) publishing
+the code on a hosted website, e.g on GitHub, with
+added features (3) publishing 
+the installed code within a virtual environment,
+for example, using a Singularity container.
+In this chapter, details of each of these ways will be described,
+including advantages, disadvantages, as well as how to
+make the code more FAIR data \cite{wilkinson2016fair}.
+
+\paragraph{Publishing a virtual environment}
+
+The most reproducible way of submitting the code of an experiment,
+is by providing the code with all the (software) dependencies 
+it needs installed in a virtual container.
+This is close to the golden standard suggested by 
+\cite{peng2011reproducible} 
+\iffalse
+(see also Figure \ref{fig:peng2011reproducible})
+\fi
+,
+however, that article was written before virtualization existed.
+Virtualization allows one to provide a standard environment,
+that can be used on any other computer.
+The best known examples of container software are Docker and Singularity.
+Due to access rights, Singularity is the only viable option to run
+an experiment on a 
+high performance computing (HPC) environment (e.g. a computer cluster).
+
+\paragraph{Containers allow for runnable code, for longer periods of time}
+
+Unlike English, code is fragile in standing the test of time.
+Containers can alleviate this, as these can be created at a point
+in time and then run later. Because the container is stand-alone,
+all software that the code depends on remain in their current/old state.
+As with hosted code, an experiment may not run with newer versions
+of libraries anymore, but a containerized experiment can run regardless, 
+if one refrains from updating (or only does so when there are there problems with fixing).
+
+Ideally, to re-do an experiment, one would need a one-liner to a container,
+for example, \verb|singularity run my_experiment.sif run|.
+
+\paragraph{Runnable code can precisely reproduce an experiment}
+
+Containers allow a computation experiment to be highly reproducible:
+given the same data, an experiment put into a container will give
+the same results on different platforms, at least in theory.
+In practice, differences may be observed when peripheral factors
+are different, such as the random numbers as generated by an operating
+system, or data that are downloaded from additional online sources.
+
+\paragraph{Runnable code allows others to re-use an experiment}
+
+Containers allow a computation experiment to be highly re-usable,
+as any scientist can work with it and tweak code.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Preserving code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Preserving code by uploading it to a code hosting webite
+has many advantages.
 
@@
 
-\subsubsection{Code must be tested}
-
-Coding errors are extremely common and contribute to the reproducibility crisis in science \cite{vable2021code}
-
-%.
-%. How to make more published research true
-%. NIH plans to enhance reproducibility
-%. The replication crisis in epidemiology: snowball, snow job, or winter solstice?
-%. The harm done to reproducibility by the culture of null hypothesis significance testing
-
-\paragraph{Continuous integration helps assess code quality}
-
-Some code hosts, among others, GitHub, 
-allow the researcher/user to start software scripts upon uploading a change.
-This is useful to determine the reproducibility of the code: 
-one such script should run the code (on trivial data) 
-to verify if the code works at all.
-If the functionality of the code is lost, the researcher
-can immediatly see which small change was the cause of this.
-A formalized version of this workflow is called continuous integration (CI).
-Using a CI service (such as GitHub Actions for GitHub), 
-is known to significantly 
-increase the number of bugs exposed \cite{vasilescu2015} and increases
-the speed at which new features are added \cite{vasilescu2015}.
-Using CI opens up the possibility to informally annotate features of
-a repository. 
-\iffalse
-For example, Figure \ref{fig:build_badge} shows a 
-build badge that indicates that a \LaTeX document (i.e. this paper)
-could be built.
-\fi
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\linewidth]{build_badge.png}
-  \caption{
-    A build badge for the GitHub repository of this article.
-    The description indicates which process it is that the badge
-    signals success (as in this case) or failure of. 
-  }
-  \label{fig:build_badge}
-\end{figure}
-\fi
-
-% \paragraph{Code coverage helps assess code quality}
-
-Testing, in general, is an important mechanism to ensure
-the correctness of code (an interesting example is \cite{rahman2020exploratory}
-showing bugs in scientific software on the COVID-19 pandemic:
-it recommends more testing).
-The percentage of (lines of) code tested is called the code coverage.
-Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
-The non-profit organisation rOpenSci \cite{ram2018community},
-which peer-reviews code,
-has made it a prerequisite to have a code coverage of 100\%.
-Also here, CI opens up the possibility to informally annotate the
-code coverage. 
-For example, build badges
-\iffalse
-(e.g. as shown in Figure \ref{fig:badge_codecov})
-\fi
-are visual indications on a website that can, among others,
-show a certain R package's code has a 100\% code coverage.
-
-Hence, code coverage is important metadata, that can automatically
-be acquired when using code hosting website.
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[]{badge_codecov.png}
-  \caption{
-    A build badge for the code coverage of an R 
-    package (\url{https://github.com/ropensci/beautier}).
-    The umbrella signals that the code coverage is hosted by
-    CodeCov (\url{https://codecov.io/}). 
-    The text indicates which process it is that the badge
-    signals, which in this case, is that the code is fully covered
-    by tests.
-  }
-  \label{fig:badge_codecov}
-\end{figure}
-\fi
-
-% \paragraph{Metrics help assess the honest division of labour}
+\paragraph{Metrics help assess the honest division of labour}
 
@@
 \fi
-
-\paragraph{Adding a license}
-
-Code hosting websites such as GitHub help developers to follow
-best practices. One of these is to add a software license \cite{jimenez2017four},
-so that it is clear that the software can be reused.
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\linewidth]{issue_ldpred2.png}
-  \caption{
-    An overview of the issues for the code for
-    LDpred2 \cite{prive2020ldpred2}. Of the 310 issues, 300 have been closed.
-    The first four issues have been replied to, as indicated by the
-    talk balloon. The fifth issues has a label with the text 'feature request'
-    to signal the issue type.
-  }
-  \label{fig:issue_ldpred2}
-\end{figure}
-\fi
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Code must be computer friendly}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-There are multiple ways of publishing code,
-that require increasingly more knowledge to do and result
-in increasingly more reproducible experiments:
-(1) publishing the code as text only, e.g. by printing
-it in a Supplementary Materials, (2) publishing
-the code on a hosted website, e.g on GitHub, with
-added features (3) publishing 
-the installed code within a virtual environment,
-for example, using a Singularity container.
-In this chapter, details of each of these ways will be described,
-including advantages, disadvantages, as well as how to
-make the code more FAIR data \cite{wilkinson2016fair}.
-
-\paragraph{Publishing a virtual environment}
-
-The most reproducible way of submitting the code of an experiment,
-is by providing the code with all the (software) dependencies 
-it needs installed in a virtual container.
-This is close to the golden standard suggested by 
-\cite{peng2011reproducible} 
-\iffalse
-(see also Figure \ref{fig:peng2011reproducible})
-\fi
-,
-however, that article was written before virtualization existed.
-Virtualization allows one to provide a standard environment,
-that can be used on any other computer.
-The best known examples of container software are Docker and Singularity.
-Due to access rights, Singularity is the only viable option to run
-an experiment on a 
-high performance computing (HPC) environment (e.g. a computer cluster).
-
-\paragraph{Containers allow for runnable code, for longer periods of time}
-
-Unlike English, code is fragile in standing the test of time.
-Containers can alleviate this, as these can be created at a point
-in time and then run later. Because the container is stand-alone,
-all software that the code depends on remain in their current/old state.
-As with hosted code, an experiment may not run with newer versions
-of libraries anymore, but a containerized experiment can run regardless, 
-if one refrains from updating (or only does so when there are there problems with fixing).
-
-Ideally, to re-do an experiment, one would need a one-liner to a container,
-for example, \verb|singularity run my_experiment.sif run|.
-
-\paragraph{Runnable code can precisely reproduce an experiment}
-
-Containers allow a computation experiment to be highly reproducible:
-given the same data, an experiment put into a container will give
-the same results on different platforms, at least in theory.
-In practice, differences may be observed when peripheral factors
-are different, such as the random numbers as generated by an operating
-system, or data that are downloaded from additional online sources.
-
-\paragraph{Runnable code allows others to re-use an experiment}
-
-Containers allow a computation experiment to be highly re-usable,
-as any scientist can work with it and tweak code.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Preserving code}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2023-02-09 article.tex
Improve

@@
 
-% \paragraph{The drawbacks of publishing code}
+\paragraph{The drawbacks of publishing code}
 
@@
 
-% \paragraph{The drawbacks of publishing version-controlled code}
+\paragraph{The drawbacks of publishing version-controlled code}
 
@@
 
-% \paragraph{The drawbacks of publishing a running version of code}
+\paragraph{The drawbacks of publishing a running version of code}
 
@@
 
-% \paragraph{Final word}
+The simplest and most impactful step to make code more useful paradata
+is, however, to publish it on a code hosting website 
+along a publication. From then on, the next steps can be taken 
+gradually as the skills of the author(s) progress.
+
+\paragraph{Final word}
 

### 2023-02-09 scripts/create.sh
Add index

@@ -9,6 +9,27 @@
 #
 #   ./scripts/create.sh
 #
+
+
+if [ ! -f "svmult.cls" ]; then
+  cp style/svmult.cls svmult.cls
+fi
+
+
+if [ ! -f "svmult.cls" ]; then
+  echo "'svmult.cls' does not exist."
+  exit 42
+fi
+
+if [ ! -f "svind.ist" ]; then
+  cp style/author/svind.ist svind.ist
+fi
+
+if [ ! -f "svind.ist" ]; then
+  echo "'svind.ist' does not exist."
+  exit 43
+fi
+
 pdflatex header.tex
 bibtex header >/dev/null
 pdflatex header.tex >/dev/null

### 2023-02-10 article.tex
Use abstract*
Remove subsubsubsection
Use begin and end for abstact
Use begin and end for abstact (no asterisk)
Check for 'svmult.cls'
Improved title

@@
 \setcounter{chapter}{7}
-\chapter{Research code is paradata and as vital as the article itself}
-
-\abstract{
-
+\chapter{How to make research code useful paradata}
+
+\abstract*{
 Paradata is data about the data collection process,
@@
 and hence allowing computational research to be truly reproducible.
-
 } % abstract
@@
 after which the code may be actually extended.
-
-\subsubsubsection{Assessment of code}
-
-\subsubsubsection{Extending code}
 

### 2023-02-10 article.tex
Add code availablity

@@
 
-\paragraph{Simple example}
+% \paragraph{Simple example}
 
@@
 that describe an experiment with computational steps.
-Both manuscripts are accepted by an equally prestigous journal after peer view. 
+Both manuscripts are accepted by an equally prestigious journal after peer review. 
 One researcher, however, does not supply the
@@
 
-\paragraph{The first use of paradata}
+% \paragraph{The first use of paradata}
 
@@
 
-\paragraph{Code is paradata}
+% \paragraph{Code is paradata}
 
@@
 'data about the data collection 
-process', as it commonly downloads data, selects relevant subsets in those
-downloaded data and performs statistical tests to generate results.
+process', as it commonly downloads data,
+selects relevant subsets in those data,
+performs statistical tests and generates figures.
+In each case, code answers the question: 
+'Where is it (i.e. the result) coming from?'.
 Code, hence, is paradata and this chapter explores and illustrates the
@@
 
-\paragraph{Making paradata useful}
+% \paragraph{Making paradata useful}
 
@@
 the documention and data should have co-evolved, 
-and the paradata should be good to computers \cite{huvila2022improving}.
-These suggestions appear to work well for code too
+and the paradata should be computer-friendly \cite{huvila2022improving}.
+At first glance, these suggestions appear to work well for code too
 and will be discussed in detail below.
 
-\paragraph{The goal of good paradata}
+% \paragraph{The goal of good paradata}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Genetic epidemiology}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\paragraph{Introduction}
-
-This paper uses genetic epidemiology as a specific example,
-to illustrate that the code is paradata,
-and that this type of paradata is relevant.
-However, any field that uses computation 
-on sensitive data in its experiments,
-could be used as an example.
-
-\paragraph{How commonly code is published}\label{para:how-commonly-code-is-published}
-
-Going back to the example scientists in the introduction,
-where one does and the other does not publish the code,
-it is not common to publish the code of an experiment or analysis 
+\section{Code availability}\label{sec:code-availability}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To determine how useful code is, it needs to be available.
+However, it is not common to publish the code of an experiment or analysis 
 \cite{stodden2011trust,read2015sizing} (with a pleasant exception 
@@
 to redo the analysis \cite{obels2020analysis}.
+For arcticles in Science magazine (years 2011-2012), 24 out of 204
+studies published the code \cite{stodden2018empirical}.
 Unpublished code has been the cause of some saddening examples,
@@
 
-\paragraph{For some fields, the experiment is actually run by code}
+% \paragraph{Obtaining code by email}
+
+When code of an academic paper is not published, 
+one can contact the correspending author and request it.
+However, the reponse rate of correnspondence authors 
+with a request for data in computational fields is around 50\% 
+\cite{manca2018non, stodden2018empirical, teunis2015corresponding} 
+(for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
+When getting an answer of the corresponding author, 
+half (65 out of 134) will actually share the code \cite{stodden2018empirical}.
+The responses of unwilling authors (see \cite{stodden2018empirical} for 
+some real examples) can come across as so caustic, 
+that one is excused from not contacting an a corresponding author.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Genetic epidemiology}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{Introduction}
+
+This paper uses genetic epidemiology as a specific example,
+to illustrate that the code is paradata,
+and that this type of paradata is relevant.
+However, any field that uses computation 
+and sensitive data in its experiments,
+could be used as an example.
+
+% \paragraph{For some fields, the experiment is actually run by code}
 
@@
 and the paper does not mention any way to obtain it.
-One could send an email to the correspending,
-yet it is known that the reponse rate of correnspondence authors 
-with a request for data in related fields is around 50\% 
-\cite{manca2018non, teunis2015corresponding} 
-(for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
-
-\paragraph{The goal of useful paradata is reproducible science}
+Sending an email is known to be unlikely to work 
+(see section \ref{sec:code-availability}).
+
+% \paragraph{The goal of useful paradata is reproducible science}
 
@@
 
-\paragraph{Sensitive data}
+% \paragraph{Sensitive data}
 
@@
 
-\paragraph{Code holds the ground truth}
+% \paragraph{Code holds the ground truth}
 
@@
 
-\paragraph{Code should be hosted}
+% \paragraph{Code should be hosted}
 
@@
 
-\paragraph{Code must have extensive error handling}
+% \paragraph{Code must have extensive error handling}
 
@@
 
-\paragraph{Using a tool for coding style}
+% \paragraph{Using a tool for coding style}
 
@@
 
-\paragraph{Using a tool for complexity}
+% \paragraph{Using a tool for complexity}
 
@@
 
-\paragraph{Code must be tested}
+% \paragraph{Code must be tested}
 
@@
 
-\paragraph{Adding a license}
+% \paragraph{Adding a license}
 
@@
 
-\paragraph{Test-driven development}
+% \paragraph{Test-driven development}
 
@@
 
-\paragraph{Code must be reviewed to ensure the correct amount of documentation}
+% \paragraph{Code must be reviewed to ensure the correct amount of documentation}
 
@@
 
-\paragraph{Publishing a virtual environment}
+% \paragraph{Publishing a virtual environment}
 
@@
 
-\paragraph{Hosted code keeps a history of changes}
+% \paragraph{Hosted code keeps a history of changes}
 
@@
 
-\paragraph{Invest in training}
+% \paragraph{Invest in training}
 
@@
 
-\paragraph{Code needs maintanance}
+% \paragraph{Code needs maintanance}
 
@@
 
-\paragraph{Authors can be contacted}
+% \paragraph{Authors can be contacted}
 
@@
 
-\paragraph{How to handle bugs?}
+% \paragraph{How to handle bugs?}
 
@@
 
-% \paragraph{Still need the same data to do a reproducible experiment}
-
-% \paragraph{Recommendations}
+% % \paragraph{Still need the same data to do a reproducible experiment}
+
+% % \paragraph{Recommendations}
 
@@
 
-\paragraph{Final goal: runnable code}
+% \paragraph{Final goal: runnable code}
 
@@
 
-\paragraph{First step: publish code alongside papers}
+% \paragraph{First step: publish code alongside papers}
 

### 2023-02-10 article.tex
Streamline

@@
 to redo the analysis \cite{obels2020analysis}.
-For arcticles in Science magazine (years 2011-2012), 24 out of 204
+For articles in Science magazine (years 2011-2012), 24 out of 204
 studies published the code \cite{stodden2018empirical}.
@@
 performed the statistics and generated the plots.
-All these steps are described in the paper (\cite{ahsan2017relative}), 
-but the code itself is unpublished
-and the paper does not mention any way to obtain it.
-Sending an email is known to be unlikely to work 
-(see section \ref{sec:code-availability}).
+For the rest of the discussion, we assume that the
+code is available to us (note that it is not
+published with the paper, see section \ref{sec:code-availability} 
+for a glum estimate of the chance of obtaining the code).
 
@@
 are fundamental characteristics of scientific studies \cite{patil2019visual}.
-For computational science, theoretically, it should be relatively easy to 
+For computational science, it seems to be relatively easy to 
 reproduce an experiment, as all it takes is a computer, electricity,
@@
 are the exception.
-To counter the lack of published code, 
-it has been suggested to make the reproduction of research 
-a minimal requirement for publication \cite{peng2011reproducible}.
+One suggested way forward is to make the reproduction of 
+research a minimal requirement for publication \cite{peng2011reproducible}.
 
@@
 it is possible (even ideal!) that no code is read at all.
-Due to this, one could argue that no documentation is needed.
-However, all code should be documented 'adequately' \cite{peng2006reproducible},
+However, the use of the code should be documented,
+as 57\% out of 56 Science magazine corresponding authors 
+that shared their code
+has documented their code well enough to run it 
+and understand what it ought to do \cite{stodden2018empirical}.
+When one is able to run the code, it could argued 
+that no further documentation is needed.
+However, all code in general should be documented 
+'adequately' \cite{peng2006reproducible},
 ideally writing code in such a way that it becomes 
@@
 so that it is clear that the software can be reused.
+About one in three of 56 Science magazine corresponding authors 
+that shared their code via a URL or email
+do not supply a software license \cite{stodden2018empirical}.
 
@@
 
+For those 22 (out of 180) Science magazine corresponding authors 
+that supplied code and that code was estimated to be 
+reproducible, one in three had the results reproduced with 
+minimal or no effort \cite{stodden2018empirical}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Hosted code commonly has a commit history. For example, GitHub
+Hosted code commonly has a commit history,
+where a commit is the description of the change, 
+as well as the changes in one or more file.
+For example, GitHub
 uses a version control system called \verb|git| (hence its name).
@@
 as well as openly working on the code from the start \cite{jimenez2017four}.
+However, half of 56 Science magazine corresponding authors 
+does not have such a version control mechanism \cite{stodden2018empirical}.
 
@@
 this allows reproducible research 
-(again, see \cite{haibe2020importance} for a tragic example).
+(see \cite{haibe2020importance} for a tragic example).
 For an author, publishing code alongside an experiment opens up

### 2023-02-10 article.tex
Progress

@@
 Again, useful paradata should be comprehensive, 
-documented appropriately, the documentation should co-evolve with the data,
+documented appropriately and in co-evolvution with the data,
 and the data should be friendly to computers \cite{huvila2022improving}.
@@
 
-For code to be ideal paradata, it must be usefully documented \cite{huvila2022improving}.
+For code to be ideal paradata, it must be usefully 
+documented \cite{huvila2022improving}.
 What is 'useful' is determined by its intended use: 
@@
 
-For purposes of reproducing (same code, same data) or replicating (same code,
-different data) a study, the minimal amount of documentation
-is determined by software development best practices.
-When an independent researcher reproduces or replicates an experiment, 
+For purposes of reproducing (same code, same data) code,
+it should at least be documented how the code is run.
+However, 57\% out of 56 Science magazine corresponding authors 
+that shared their code has documented their code well enough to run it 
+and to be understood what it ought to do \cite{stodden2018empirical}.
+
+However, when an independent researcher can redo an experiment, 
 it is possible (even ideal!) that no code is read at all.
-However, the use of the code should be documented,
-as 57\% out of 56 Science magazine corresponding authors 
-that shared their code
-has documented their code well enough to run it 
-and understand what it ought to do \cite{stodden2018empirical}.
-When one is able to run the code, it could argued 
-that no further documentation is needed.
+Within that context, it could argued that no further documentation is needed.
 However, all code in general should be documented 
@@
 For code to be ideal paradata, 
-its creation and documentation need to match, 
-'not least because they shape each other' \cite{huvila2022improving}.
-This quote resonates strongly with the idea of 
+its creation and documentation need to align, 
+\emph{not least because they shape each other} \cite{huvila2022improving}.
+This emphasized part of the quote resonates strongly with the idea of 
 literate programming \cite{knuth1984literate},
 in which documentation and code is developed hand-in-hand.
+It is the practice of writing a document mixed with code that 
+will be run.
+Contemporary examples of this idea are, among others, 
+vignettes \cite{wickham2015r} (for the R programming language \cite{r}) 
+and Jupyter notebooks \cite{wang2020assessing} 
+(for the Julia \cite{Julia-2017}, Python \cite{van1995python} and R \cite{r}
+programming languages).
 rOpenSci, a community that, among others, reviews programming

### 2023-02-10 article.tex
Copy file

@@
 For code to be extensive enough to reproduce an existing experiment,
-it should be distributed in standard ways \cite{peng2006reproducible}.
+it should be distributed in standard ways \cite{peng2006reproducible}
+(see section \ref{sec:code-availability}
+for the general availability of code).
 The standard way to publish code is to use a code hosting website.
@@
 changes are common and should be embraced \cite{beck2000extreme}.
-When a scientist intends to reuse code by extending its functionality,
-first the code will be assessed,
-after which the code may be actually extended.
-
-Code is just 'another kind of data' and should be designed 
-as such, for example, by using tools to work on it \cite{wilson2022twelve}.
-
-% \paragraph{Using a tool for coding style}
-
-A first example is to use a tool to enforce a coding style 
-(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
-or PEP 8 \cite{van2001pep} for Python).
-Following a consistent coding style improves software quality \cite{fang2001}.
-
-% \paragraph{Using a tool for complexity}
-
-A second example is to use a tool to enforce a low cyclomatic complexity.
-The cyclomatic complexity is approximately defined 
-as the number of independent paths that
-the code can be executed. 
-For example, code that has only one \verb|if| statement
-has a cyclomatic complexity of 2, as the condition within the \verb|if|
-statement can be true or false,
-resulting in the body of the \verb|if| statement being either
-exectuted or ignored.
-The cyclomatic complexity correlates with code complexity,
-where more complex code is likelier to contain or give rise to bugs 
-\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}
 
@@
 and contribute to the reproducibility crisis in science \cite{vable2021code}.
-
-%.
-%. How to make more published research true
-%. NIH plans to enhance reproducibility
-%. The replication crisis in epidemiology: snowball, snow job, or winter solstice?
-%. The harm done to reproducibility by the culture of null hypothesis significance testing
+Testing, in general, is an important mechanism to ensure
+the correctness of code (an interesting example is \cite{rahman2020exploratory}
+showing bugs in scientific software on the COVID-19 pandemic).
+
+% \paragraph{Code coverage helps assess code quality}
+
+The percentage of (lines of) code tested is called the code coverage.
+Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
+The non-profit organisation rOpenSci \cite{ram2018community},
+which peer-reviews code,
+has made it a prerequisite to have a code coverage of 100\%.
+Also here, CI opens up the possibility to informally annotate the
+code coverage. 
+For example, build badges
+\iffalse
+(e.g. as shown in Figure \ref{fig:badge_codecov})
+\fi
+are visual indications on a website that can, among others,
+show a certain R package's code has a 100\% code coverage.
+
+Hence, code coverage is important metadata, that can automatically
+be acquired when using code hosting website.
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[]{badge_codecov.png}
+  \caption{
+    A build badge for the code coverage of an R 
+    package (\url{https://github.com/ropensci/beautier}).
+    The umbrella signals that the code coverage is hosted by
+    CodeCov (\url{https://codecov.io/}). 
+    The text indicates which process it is that the badge
+    signals, which in this case, is that the code is fully covered
+    by tests.
+  }
+  \label{fig:badge_codecov}
+\end{figure}
+\fi
 
@@
 
-% \paragraph{Code coverage helps assess code quality}
-
-Testing, in general, is an important mechanism to ensure
-the correctness of code (an interesting example is \cite{rahman2020exploratory}
-showing bugs in scientific software on the COVID-19 pandemic:
-it recommends more testing).
-The percentage of (lines of) code tested is called the code coverage.
-Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
-The non-profit organisation rOpenSci \cite{ram2018community},
-which peer-reviews code,
-has made it a prerequisite to have a code coverage of 100\%.
-Also here, CI opens up the possibility to informally annotate the
-code coverage. 
-For example, build badges
-\iffalse
-(e.g. as shown in Figure \ref{fig:badge_codecov})
-\fi
-are visual indications on a website that can, among others,
-show a certain R package's code has a 100\% code coverage.
-
-Hence, code coverage is important metadata, that can automatically
-be acquired when using code hosting website.
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[]{badge_codecov.png}
-  \caption{
-    A build badge for the code coverage of an R 
-    package (\url{https://github.com/ropensci/beautier}).
-    The umbrella signals that the code coverage is hosted by
-    CodeCov (\url{https://codecov.io/}). 
-    The text indicates which process it is that the badge
-    signals, which in this case, is that the code is fully covered
-    by tests.
-  }
-  \label{fig:badge_codecov}
-\end{figure}
-\fi
-
 % \paragraph{Adding a license}
@@
 so that it is clear that the software can be reused.
-About one in three of 56 Science magazine corresponding authors 
-that shared their code via a URL or email
-do not supply a software license \cite{stodden2018empirical}.
+About two-thirds of 56 Science magazine corresponding authors 
+that shared their code supply a software license \cite{stodden2018empirical}.
 
@@
 
+% \paragraph{Ideal paradata}
+
+For paradata to be useful, it has to be computer-friendly, yet 
+'the best paradata does not necessarily look like 'data' at all for its
+human users' \cite{huvila2022improving}.
+There are features of code that humans find useful,
+without directly being able to measure these.
+In the end, code is just 'another kind of data' and should be designed 
+as such, for example, by using tools to work on it \cite{wilson2022twelve}.
+
+% \paragraph{Using a tool for coding style}
+
+A first example is to use a tool to enforce a coding style 
+(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
+or PEP 8 \cite{van2001pep} for Python),
+as following a consistent coding style improves software quality \cite{fang2001}.
+
+% \paragraph{Using a tool for complexity}
+
+A second example is to use a tool to enforce a low cyclomatic complexity.
+The cyclomatic complexity is approximately defined 
+as the number of independent paths that
+the code can be executed. 
+For example, code that has only one \verb|if| statement
+has a cyclomatic complexity of 2, as the condition within the \verb|if|
+statement can be true or false,
+resulting in the body of the \verb|if| statement being either
+executed or ignored.
+The cyclomatic complexity correlates with code complexity,
+where more complex code is likelier to contain or give rise to bugs 
+\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}.
+
+Such tools combine great with CI services.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 has many advantages.
+Code is harder to preserve than an English text
+and preserving code is rarely done \cite{barnes2010publish}.
+Although code is the primary actor in computational experiments,
+there is no incentive to submit code alongside a publication.
+Most academic journal do not require authors to submit their code,
+nor is the submitted code peer reviewed, 
+rOpenSci \cite{ram2018community} being the pleasant exception.
 
@@
 or not metadata at all (Syslabs).
-
-Code is harder to preserve than an English text
-and preserving code is rarely done \cite{barnes2010publish}.
-Although code is the primary actor in computational experiments,
-there is no incentive to submit code alongside a publication.
-Most academic journal do not require authors to submit their code,
-nor is the submitted code peer reviewed, 
-rOpenSci \cite{ram2018community} being the pleasant exception.
 

### 2023-02-10 scripts/create_normal.sh
Prove that refs can be normal

@@ -1,0 +1,17 @@
+#!/bin/bash
+#
+# Create the PDF 'article.pdf'
+#
+# It does so by converting the TeX to PDF, 
+# as well as stitching some other PDFs to it
+#
+# Usage:
+#
+#   ./scripts/create_normal.sh
+#
+
+pdflatex header_normal.tex
+bibtex header_normal >/dev/null
+pdflatex header_normal.tex >/dev/null
+pdflatex header_normal.tex
+mv header_normal.pdf article.pdf

### 2023-02-11 article.tex
Add data preservation

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Preserving code}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Preserving code in a reliable way is rare \cite{barnes2010publish}.
+In general, within software development, 
+changes at all levels are common
+and instead of ignoring that, it should be embraced \cite{beck2000extreme}.
+This section discusses the preservation of code for a short, medium
+and long term.
+
+\subsection{Code hosting}
+
+To preserve code for a short term, 
+a code hosting website is a good first step.
+A code hosting website is a website where 
+its users can create dedicated pages (called 'repositories')
+for a project, upload code and interact with that code.
+There are multiple code hosting websites, 
+with GitHub being the most popular one \cite{cosentino2017systematic}.
+The \LaTeX code of this chapter is hosted on GitHub 
+as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
+The use of code hosting websites
+has increased strongly \cite{russell2018large},
+accommodates collaboration \cite{perez2016ten}
+and improves transparency \cite{gorgolewski2016practical}.
+See \cite{cosentino2017systematic} for an extensive overview of
+research conducted on GitHub.
+
+\iffalse
+An example from the author is the website \cite{bbbqarticleissue157},
+that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
+as shown in Figure \ref{fig:bbbqarticleissue157}.
+\fi
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{bbbq_article_issue_157.png}
+  \caption{
+    A typical GitHub repository \cite{bbbqarticleissue157}, 
+    hosting the code for a
+    computational experiment, 
+    as published with \cite{bilderbeek2022transmembrane}.
+    Note that this repository has 72 commits, where a commit is a change
+    to the code.
+  }
+  \label{fig:bbbqarticleissue157}
+\end{figure}
+\fi
+
+However, if the code breaks tomorrow, it will go unnoticed.
+
+\subsection{Continuous integration}
+
+To preserve code for a longer timespan, 
+continuous integration allows one to 
+verify if code still works and, if not, to be notified on the need for an update.
+
+% \paragraph{Continuous integration helps assess code quality}
+
+Some code hosts, among others, GitHub, 
+allow the researcher/user to start software scripts upon uploading a change.
+This is useful to determine the reproducibility of the code: 
+one such script should run the code (on trivial data) 
+to verify if the code works at all.
+If the functionality of the code is lost, the researcher
+can immediatly see which small change was the cause of this.
+A formalized version of this workflow is called continuous integration (CI).
+Using a CI service (such as GitHub Actions for GitHub), 
+is known to significantly 
+increase the number of bugs exposed \cite{vasilescu2015} and increases
+the speed at which new features are added \cite{vasilescu2015}.
+Using CI opens up the possibility to informally annotate features of
+a repository. 
+\iffalse
+For example, Figure \ref{fig:build_badge} shows a 
+build badge that indicates that a \LaTeX document (i.e. this paper)
+could be built.
+\fi
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{build_badge.png}
+  \caption{
+    A build badge for the GitHub repository of this article.
+    The description indicates which process it is that the badge
+    signals success (as in this case) or failure of. 
+  }
+  \label{fig:build_badge}
+\end{figure}
+\fi
+
+However, for code to work longer, one can put the code and its dependencies in
+a container, that should be best in standing the test of time.
+
+\subsection{Containerization}
+
+To preserve code for the longest time, 
+both code and its dependencies are put into a so-called container.
+
+The most reproducible way of submitting the code of an experiment,
+is by providing the code with all the (software) dependencies 
+it needs installed in a virtual container.
+This is close to the golden standard suggested by 
+\cite{peng2011reproducible} 
+\iffalse
+(see also Figure \ref{fig:peng2011reproducible})
+\fi
+,
+however, that article was written before virtualization existed.
+Virtualization allows one to provide a standard environment,
+that can be used on any other computer.
+The best known examples of container software are Docker and Singularity.
+Due to access rights, Singularity is the only viable option to run
+an experiment on a 
+high performance computing (HPC) environment (e.g. a computer cluster).
+
+% \paragraph{Containers allow for runnable code, for longer periods of time}
+
+Unlike English, code is fragile in standing the test of time.
+Containers can alleviate this, as these can be created at a point
+in time and then run later. Because the container is stand-alone,
+all software that the code depends on remain in their current/old state.
+As with hosted code, an experiment may not run with newer versions
+of libraries anymore, but a containerized experiment can run regardless, 
+if one refrains from updating (or only does so when there are there problems with fixing).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Making code useful paradata}\label{sec:making-code-useful-paradata}
@@
 
-Again, useful paradata should be comprehensive, 
-documented appropriately and in co-evolvution with the data,
+To repeat features of useful paradata: paradata should be comprehensive, 
+documented appropriately, documented in co-evolution with the data,
 and the data should be friendly to computers \cite{huvila2022improving}.
 In this section, these ideal properties of paradata are applied to code.
-Code as paradata, however, has multiple uses as well,
-which are to reproduce, to replicate and to 
-extend a computational experiment \cite{benureau2018re}.
+However, code has multiple uses, which have different requirements
+for being usesful paradata.
+Code can be used to reproduce, to replicate and to 
+extend a computational experiment \cite{benureau2018re},
+each of these uses has different requirements for the code to be useful paradata.
 In this section, the recommendations how to make paradata useful
@@
 
+Code should be hosted.
+
 For code to be extensive enough to reproduce an existing experiment,
@@
 The standard way to publish code is to use a code hosting website.
-There are multiple code hosting websites, 
-with GitHub being the most popular one \cite{cosentino2017systematic}.
-GitHub is a website where users can create dedicated pages (called 'repositories')
-to upload code and interact with it.
-The use of such code hosting websites
-has increased strongly \cite{russell2018large},
-accommodates collaboration \cite{perez2016ten}
-and improves transparency \cite{gorgolewski2016practical}.
-See \cite{cosentino2017systematic} for an extensive overview of
-research conducted on GitHub.
-
-\iffalse
-An example from the author is the website \cite{bbbqarticleissue157},
-that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
-as shown in Figure \ref{fig:bbbqarticleissue157}.
-\fi
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\linewidth]{bbbq_article_issue_157.png}
-  \caption{
-    A typical GitHub repository \cite{bbbqarticleissue157}, 
-    hosting the code for a
-    computational experiment, 
-    as published with \cite{bilderbeek2022transmembrane}.
-    Note that this repository has 72 commits, where a commit is a change
-    to the code.
-  }
-  \label{fig:bbbqarticleissue157}
-\end{figure}
-\fi
 
@@
 
-In software development, 
-changes are common and should be embraced \cite{beck2000extreme}.
-
 % \paragraph{Code must be tested}
@@
 
-% \paragraph{Continuous integration helps assess code quality}
-
-Some code hosts, among others, GitHub, 
-allow the researcher/user to start software scripts upon uploading a change.
-This is useful to determine the reproducibility of the code: 
-one such script should run the code (on trivial data) 
-to verify if the code works at all.
-If the functionality of the code is lost, the researcher
-can immediatly see which small change was the cause of this.
-A formalized version of this workflow is called continuous integration (CI).
-Using a CI service (such as GitHub Actions for GitHub), 
-is known to significantly 
-increase the number of bugs exposed \cite{vasilescu2015} and increases
-the speed at which new features are added \cite{vasilescu2015}.
-Using CI opens up the possibility to informally annotate features of
-a repository. 
-\iffalse
-For example, Figure \ref{fig:build_badge} shows a 
-build badge that indicates that a \LaTeX document (i.e. this paper)
-could be built.
-\fi
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\linewidth]{build_badge.png}
-  \caption{
-    A build badge for the GitHub repository of this article.
-    The description indicates which process it is that the badge
-    signals success (as in this case) or failure of. 
-  }
-  \label{fig:build_badge}
-\end{figure}
-\fi
-
 % \paragraph{Adding a license}
@@
 it needs installed in a virtual container.
-This is close to the golden standard suggested by 
-\cite{peng2011reproducible} 
-\iffalse
-(see also Figure \ref{fig:peng2011reproducible})
-\fi
-,
-however, that article was written before virtualization existed.
-Virtualization allows one to provide a standard environment,
-that can be used on any other computer.
-The best known examples of container software are Docker and Singularity.
-Due to access rights, Singularity is the only viable option to run
-an experiment on a 
-high performance computing (HPC) environment (e.g. a computer cluster).
-
-% \paragraph{Containers allow for runnable code, for longer periods of time}
-
-Unlike English, code is fragile in standing the test of time.
-Containers can alleviate this, as these can be created at a point
-in time and then run later. Because the container is stand-alone,
-all software that the code depends on remain in their current/old state.
-As with hosted code, an experiment may not run with newer versions
-of libraries anymore, but a containerized experiment can run regardless, 
-if one refrains from updating (or only does so when there are there problems with fixing).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Preserving code by uploading it to a code hosting webite
-has many advantages.
-Code is harder to preserve than an English text
-and preserving code is rarely done \cite{barnes2010publish}.
-Although code is the primary actor in computational experiments,
-there is no incentive to submit code alongside a publication.
-Most academic journal do not require authors to submit their code,
-nor is the submitted code peer reviewed, 
-rOpenSci \cite{ram2018community} being the pleasant exception.
 

### 2023-02-11 article.tex
EOD

@@
 The chapter concludes by some rules how to better code to serve as paradata,
-and hence allowing computational research to be truly reproducible.
+and hence allowing computational research to be more reproducible.
 } % abstract
@@
     Code         & Body of text to be run by a computer                                         & The scripts run by a computational experiment                   \\
-    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
+    Data         & Individual facts, statistics, or items of information                        & The name of a SNP known to have a significant association       \\
     Genotype     & The DNA allele at a certain location                                         & AA, AC, CG, GT, ...                                             \\
@@
     Phenotype    & How an organism looks like in the broadest sense                             & The concentration of IL6RA in the blood                         \\
-    Primary data & Data that are worked with in a research            & The genotype of individuals   \\
+    Primary data & Data that is the foundation of an experiment                                 & The genotype of individuals                                     \\
     Metadata     & Data that provide information about other data                               & The article that describes an experiment                        \\
     Raw data     & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals         \\
+    SNP          & A named location within a genome                                             & \verb|rs12133641|                                               \\
     Trait        & A phenotype                                                                  & The concentration of IL6RA in the blood                         
@@
 Is this difference relevant and worth the effort?
+How common is it to share code, and, if shared, how can it be preserved?
 This chapter discusses why code, a type of paradata, 
-should be supplied
-and how it should be supplied,
-and claims that code is just as important as the paper itself.
+should be supplied and what features it needs to have
+for it to be useful.
 
@@
 and the paradata should be computer-friendly \cite{huvila2022improving}.
-At first glance, these suggestions appear to work well for code too
+At first glance, these suggestions appear to work well for code
 and will be discussed in detail below.
@@
 data is produced.
-This insight helps researchers from different fields to understand each other
+This knowledge helps researchers from different fields to understand each other
 and collaborate.
@@
 a field intimately familiar with computer code,
-192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
+42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
 Another study analysed the reproducibility of registered reports
 in the field of psychology, 
-where 37 out of 62 studies supplied the code 
+where 60\% of 62 studies supplied the code 
 to redo the analysis \cite{obels2020analysis}.
-For articles in Science magazine (years 2011-2012), 24 out of 204
+For articles in Science magazine (years 2011-2012), 12\% of 204
 studies published the code \cite{stodden2018empirical}.
@@
 When code of an academic paper is not published, 
-one can contact the correspending author and request it.
+one could contact the correspending author and request it.
 However, the reponse rate of correnspondence authors 
@@
 \cite{manca2018non, stodden2018empirical, teunis2015corresponding} 
-(for the field of emergency medicine, these rates are 73\% \cite{o2003email}).
+(the field of emergency medicine seems to be a pleasant outlier, 
+with 73\% of 118 emails being replied \cite{o2003email}).
 When getting an answer of the corresponding author, 
-half (65 out of 134) will actually share the code \cite{stodden2018empirical}.
+48\% out of 134 replied emails will actually result 
+in a sharing of the code \cite{stodden2018empirical}.
 The responses of unwilling authors (see \cite{stodden2018empirical} for 
@@
 This paper uses genetic epidemiology as a specific example,
-to illustrate that the code is paradata,
-and that this type of paradata is relevant.
+to illustrate in which way code is paradata,
+and why this type of paradata is relevant.
 However, any field that uses computation 
@@
 create so-called 'gene panels', 
-where the gene causing 
-an association is measured specifically, to, among others,
-detect people at risk for the associated phenotype.
+where the location of the gene causing 
+an association is measured specifically, 
+to detect people at risk for the associated phenotype.
 The rest of this section describes a genetic epidemiology study 
@@
 consists out of single nucleotide polymorphisms (SNPs, pronounced 'snips').
-A SNP has a name and a location on the DNA, at which there is a certain nucleotide.
-DNA (organized into chromosomes and present in every (nucleated) cell, 
+A SNP has a name and a location within the genome.
+At the SNP's location in the DNA, there will be two nucleotides.
+One of these nucleotides is inherited from the mother, 
+the other from the father.
+The DNA consists out of billions of nucleotides.
+There are four types of nucleotides,
+called adenosine, cytosine, guanine and thyrosine, 
+commonly abreviated as A, C, G and T respectively
 \iffalse
@@
 \fi
-consists out of billions of nucleotides.
-There are four types of nucleotides, 
-called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
-as A, C, G and T respectively.
+.
+
 One SNP example is \verb|rs12133641|, which is a SNP located at position 
-154,428,283, where 67 percent of the people within this study have an A,
+154,428,283 (it is at the 154 millionth position, which happens to be located
+on the first chromosome), 
+where 67 percent of the people within this study have an A,
 and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).
+From this it follow (assuming the nucleotides act independently)
+that 45\% of subjects have the genotype AA, 44\% have AG and 11\% have GG.
 
@@
 are concentrations of proteins in the blood. 
-DNA contains the code for building proteins
+The nucleotides of the DNA contain the code for building proteins
 \iffalse
@@
 \fi
-, as well as the rate
-at which a protein is created. Some proteins end up in the blood and
+,
+as well as the rate at which a protein is created (for 
+sake of simplicity it is assumed that such a rate is constant,
+in practice there are complex regulation mechanisms).
+Some proteins end up in the blood and
 their presence can be used to assess the health of an individual.
@@
 . The amount of variance that can be explained by an association (i.e.
-the R squared value) is rarely 100 percent, which means that a trait (in
+the \begin{math}R^2\end{math}) is rarely 100\%, which means that a trait (in
 this case, the concentration of IL6RA) cannot be perfectly explained
 from the genotype (in this case, SNP \verb|rs12133641|) alone. 
-In this example, 43 percent of the variance 
+In this example, as much as 43\% of the variance 
 can be attributed to an individuals' genotype
@@
 such as the effect
-of the environment (e.g. geographic location, time of day), 
-lifestyle (e.g. smoking yes/no) or having a disease (e.g. diabetes) 
+of the environment (e.g. geographic location, time of day the measurement 
+was done), lifestyle (e.g. smoking yes/no) or having a disease (e.g. diabetes) 
 are needed to explain the additional variation.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Code is just as important as the paper}
+Additionally, there is an interdependency of scholarly findings here:
+the SNP has received its name based on a computational experiment.
+The experiment that concluded the usefulness of that SNP, 
+as well as this experiment, are based on DNA sequences.
+DNA sequences, however, are (nowadays) not simply read, 
+yet are the result of a complex computational analysis,
+with its own dedicated field of research.
+Both studies assumed a correct inference of DNA.
+This means that if the DNA sequence analysis contained a software bug, 
+this study may be invalidated. 
+Additionally, the result of this study may be used in follow-up studies,
+that assume the result to be correctly calculated: one paper's conclusion
+is the next paper's assumption.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Why code is useful paradata}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-02-12 article.tex
Review

@@
 For the rest of the discussion, we assume that the
-code is available to us (note that it is not
-published with the paper, see section \ref{sec:code-availability} 
+code is available to us (if not, see section \ref{sec:code-availability} 
 for a glum estimate of the chance of obtaining the code).
 
+% \paragraph{Why useful code}
+
+There are multiple reasons why (useful) code matters,
+and these are the same reasons as why useful paradata matters:
+code gives an understanding how the raw results 
+and the subsequent scholarly knowledge is obtained from an experiment.
+Additionally, code helps researchers from different fields 
+to understand each other and collaborate.
+Additionally, code helps Open Science reach its goals of openness and
+transparency.
+The core of these reasons it to achieve reproducible science:
+that any person in any field can redo a computational experiment
+and see exactly what happened.
+
 % \paragraph{The goal of useful paradata is reproducible science}
 
-The reason behind paradata is, among others,
-to assure reproducible research \cite{huvila2022improving}.
-Reproducibility (i.e. to reproduce the same results) 
-and replicability (i.e. to reconclude a conclusion)
-are fundamental characteristics of scientific studies \cite{patil2019visual}.
-For computational science, it seems to be relatively easy to 
+For computational science, it may appear to be relatively easy to 
 reproduce an experiment, as all it takes is a computer, electricity,
 an optional internet connection, the code and the data.
-In practice, however, it appears that 
+In practice, however, 
+only 18\% of 180 studies are easily reproducible \cite{stodden2018empirical}.
+To some, it appears that 
 the academic culture to reproduce results 
-has been lost over time in general \cite{peng2011reproducible},
+has been lost over time \cite{peng2011reproducible},
 with labs that embrace reproducibility (for example, \cite{barba2016hard})
@@
 
-From a broader perspective, code should be published, as it holds the
-ground truth of an experiment; it does the actual work.
+Code holds the ground truth of an experiment; it does the actual work.
 The more complex the computation pipeline is, the easier it is
@@
 simple (and may legitimately fear that those simple errors are common) 
-\cite{baggerly2009deriving},
+\cite{baggerly2009deriving, vable2021code},
 and this example is not too far off from the
@@
 changes at all levels are common
-and instead of ignoring that, it should be embraced \cite{beck2000extreme}.
+and embracing those changes is recommended \cite{beck2000extreme}
+and useful code does so.
 This section discusses the preservation of code for a short, medium
@@
 accommodates collaboration \cite{perez2016ten}
-and improves transparency \cite{gorgolewski2016practical}.
+and improves transparency \cite{gorgolewski2016practical},
+due to its inherent computer-friendliness.
 See \cite{cosentino2017systematic} for an extensive overview of
@@
   \label{fig:bbbqarticleissue157}
+\end{figure}
+\fi
+
+% \paragraph{Hosted code keeps a history of changes}
+
+Hosted code commonly has a commit history,
+where a commit is the description of the change, 
+as well as the changes in one or more file.
+For example, GitHub
+uses a version control system called \verb|git| (hence its name).
+In practice, this means that when a change is made to the code,
+a new version is created. In the case that the change was harmful,
+one can go back to an earlier version and continue again from there.
+It is a general recommendation to put version control
+on all human-produced data \cite{wilson2014best},
+as well as openly working on the code from the start \cite{jimenez2017four}.
+However, half of 56 Science magazine corresponding authors 
+does not have such a version control mechanism \cite{stodden2018empirical}.
+
+% \paragraph{A commit history helps assess honest reporting}
+
+Hosted code improves transparency \cite{gorgolewski2016practical},
+as, among others, the history of changes (i.e. the git commit history), 
+helps to assess the honesty in the reporting of the code. 
+Ideally, the full version history of the code 
+has been generated, by hosting the code directly at its inception,
+so that the ontology of the experiment can be viewed.
+However, in some repositories, only the final version of the code has been
+posted, which means all historical information is lost.
+The code history has many important properties,
+as, for example, 
+it can be observed how much statistical tests have really/actually been done 
+and see if all are reported:
+failing to reporting a (usually non-significant) statistical tests is 
+one of the many degrees of freedom 
+in p-value hacking \cite{wicherts2016degrees}.
+
+% \paragraph{Metrics help assess the honest division of labour}
+
+Most code hosts, among others, GitHub and GitLab,
+automatically keep track of certain metrics.
+One such metric is the amount of code each contributor has made.
+In that way, authors involved in the writing of the software
+are honestly acknowledged. 
+\iffalse
+Figure \ref{fig:daisie_contributors}
+shows the number of commits the authors of DAISIE \cite{etienne2020daisie}
+have made to its code.
+\fi
+The distribution of contributed code per author 
+is useful metadata, that can automatically
+be acquired when using code hosting website.
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{daisie_contributors.png}
+  \caption{
+    An overview of the commits that multiple contributors have
+    made, in this case, to the code of DAISIE \cite{etienne2020daisie}
+  }
+  \label{fig:daisie_contributors}
 \end{figure}
@@
 However, for code to work longer, one can put the code and its dependencies in
-a container, that should be best in standing the test of time.
+a container, that should be best in standing the test of time,
+as the degradation of software is a known feature for nearly 
+four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
+here a more modern (and in this context superior)
+definition is used: software collapse \cite{hinsen2019dealing},
+in which software fails due to dependencies on other 
+software.
+Where that paper recommends using older software foundations 
+with multiple implementations \cite{hinsen2019dealing},
+this paper recommends a modern solution that potentially has more 
+longevity:
+to publish the code installed in a virtual environment
+(including the dependencies used at that time), as described below.
+
+% \paragraph{Ideal paradata}
+
+For paradata to be useful, it has to be computer-friendly, yet 
+'the best paradata does not necessarily look like 'data' at all for its
+human users' \cite{huvila2022improving}.
+There are features of code that humans find useful,
+without directly being able to measure these.
+In the end, code is just 'another kind of data' and should be designed 
+as such, for example, by using tools to work on it \cite{wilson2022twelve}.
+
+% \paragraph{Using a tool for coding style}
+
+A first example is to use a tool to enforce a coding style 
+(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
+or PEP 8 \cite{van2001pep} for Python),
+as following a consistent coding style improves software quality \cite{fang2001}.
+
+% \paragraph{Using a tool for complexity}
+
+A second example is to use a tool to enforce a low cyclomatic complexity.
+The cyclomatic complexity is approximately defined 
+as the number of independent paths that
+the code can be executed. 
+For example, code that has only one \verb|if| statement
+has a cyclomatic complexity of 2, as the condition within the \verb|if|
+statement can be true or false,
+resulting in the body of the \verb|if| statement being either
+executed or ignored.
+The cyclomatic complexity correlates with code complexity,
+where more complex code is likelier to contain or give rise to bugs 
+\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}.
+
+Such tools combine great with CI services.
 
@@
 an experiment on a 
-high performance computing (HPC) environment (e.g. a computer cluster).
+high performance computing (HPC) environment (e.g. a computer cluster),
+as are commonly needed in the field of genetic epidemiology experiments.
 
@@
 What is 'useful' is determined by its intended use: 
-to reproduce, replicate or extend \ref{sec:making-code-useful-paradata}.
+to reproduce, replicate or extend (see section \ref{sec:making-code-useful-paradata}).
 The amount of documentation is, however, also determined
@@
 it should at least be documented how the code is run.
-However, 57\% out of 56 Science magazine corresponding authors 
-that shared their code has documented their code well enough to run it 
+However, 57\% out of 56 Science papers with obtainable code (in 
+total there were 180 papers) have documented their code well enough to run it 
 and to be understood what it ought to do \cite{stodden2018empirical}.
@@
 in which documentation and code is developed hand-in-hand.
-It is the practice of writing a document mixed with code that 
+Literate programming is the practice of writing a document mixed with code that 
 will be run.
@@
 Error handling is one of the mechanisms to do so.
-A simple example is when one (e.g. a BSc student) 
-needs to calculate the variance of a distribution.
-Commonly one needs to call a function to do so, 
-e.g. \verb|var| in R, that needs the values of the distribution
-as input.
-However, to calculate the variance of a distribution, at least two values
-are required (else a division by zero occurs in the calcation of the variance). 
-When the variance of (zero or) only one value is requested,
-an error message can helpfully indicate that the function 
-needs at least two values to work as expected.
+One example is, again, when doing a t-test:
+to compare the means of two distributions, each distribution
+must have at least two values
+and, if not, the programming language should signal the error clearly.
+In genetic epidemiology, it is common to have incomplete or missing data,
+so analyses should take into account that there are not always enough
+observations to do a t-test.
+The analytical code should clearly signal this problem.
 
@@
 has made it a prerequisite to have a code coverage of 100\%.
-Also here, CI opens up the possibility to informally annotate the
+CI opens up the possibility to informally annotate the
 code coverage. 
@@
 show a certain R package's code has a 100\% code coverage.
-
-Hence, code coverage is important metadata, that can automatically
-be acquired when using code hosting website.
 
@@
 
-Code hosting websites such as GitHub help developers to follow
-best practices. One of these is to add a software license \cite{jimenez2017four},
+It is considered good practice
+to add a software license \cite{jimenez2017four},
 so that it is clear that the software can be reused.
@@
 
-More than half of all scientists have their code 
-rarely or never reviewed \cite{vable2021code},
+In an informal Twitter poll,
+more than half of all 315 responding scientists 
+have their code rarely or never reviewed \cite{vable2021code},
 although code reviews are known to
@@
 code reviews are recommended as well \cite{wilson2014best} 
+The observation that code is reviewed, makes it likelier to be correct.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-There are multiple ways of publishing code,
-that require increasingly more knowledge to do and result
-in increasingly more reproducible experiments:
-(1) publishing the code as text only, e.g. by printing
-it in a Supplementary Materials, (2) publishing
-the code on a hosted website, e.g on GitHub, with
-added features (3) publishing 
-the installed code within a virtual environment,
-for example, using a Singularity container.
-In this chapter, details of each of these ways will be described,
-including advantages, disadvantages, as well as how to
-make the code more FAIR data \cite{wilkinson2016fair}.
 
@@
 it needs installed in a virtual container.
-
-Ideally, to re-do an experiment, one would need a one-liner to a container,
-for example, \verb|singularity run my_experiment.sif run|.
 
@@
 are different, such as the random numbers as generated by an operating
-system, or data that are downloaded from additional online sources.
+system, or data that are downloaded from online (and hence, changing) sources.
 
@@
 
-% \paragraph{Ideal paradata}
-
-For paradata to be useful, it has to be computer-friendly, yet 
-'the best paradata does not necessarily look like 'data' at all for its
-human users' \cite{huvila2022improving}.
-There are features of code that humans find useful,
-without directly being able to measure these.
-In the end, code is just 'another kind of data' and should be designed 
-as such, for example, by using tools to work on it \cite{wilson2022twelve}.
-
-% \paragraph{Using a tool for coding style}
-
-A first example is to use a tool to enforce a coding style 
-(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
-or PEP 8 \cite{van2001pep} for Python),
-as following a consistent coding style improves software quality \cite{fang2001}.
-
-% \paragraph{Using a tool for complexity}
-
-A second example is to use a tool to enforce a low cyclomatic complexity.
-The cyclomatic complexity is approximately defined 
-as the number of independent paths that
-the code can be executed. 
-For example, code that has only one \verb|if| statement
-has a cyclomatic complexity of 2, as the condition within the \verb|if|
-statement can be true or false,
-resulting in the body of the \verb|if| statement being either
-executed or ignored.
-The cyclomatic complexity correlates with code complexity,
-where more complex code is likelier to contain or give rise to bugs 
-\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}.
-
-Such tools combine great with CI services.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Preserving code}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{Hosted code has important metadata}
-
-Hosted code has many additional features, such as
-a version-controlled history, continuous integration scripts
-and automatically generated metrics.
-A code hosting website produces and shares these metadata.
-These metadata can help to determine the
-reproducibility, correctness and quality of code
-of computational experiments.
-Below, those features and why these are important are discussed.
-
-% \paragraph{Hosted code keeps a history of changes}
-
-Hosted code commonly has a commit history,
-where a commit is the description of the change, 
-as well as the changes in one or more file.
-For example, GitHub
-uses a version control system called \verb|git| (hence its name).
-In practice, this means that when a change is made to the code,
-a new version is created. In the case that the change was harmful,
-one can go back to an earlier version and continue again from there.
-It is a general recommendation to put version control
-on all human-produced data \cite{wilson2014best},
-as well as openly working on the code from the start \cite{jimenez2017four}.
-However, half of 56 Science magazine corresponding authors 
-does not have such a version control mechanism \cite{stodden2018empirical}.
-
-% \paragraph{A commit history helps assess honest reporting}
-
-Hosted code improves transparency \cite{gorgolewski2016practical},
-as, among others, the history of changes (i.e. the git commit history), 
-helps to assess the honesty in the reporting of the code. 
-Ideally, the full version history of the code 
-has been generated, by hosting the code directly at its inception,
-so that the ontology of the experiment can be viewed.
-However, in some repositories, only the final version of the code has been
-posted, which means all historical information is lost.
-The code history has many important properties,
-as, for example, 
-it can be observed how much statistical tests have really/actually been done 
-and see if all are reported:
-failing to reporting a (usually non-significant) statistical tests is 
-one of the many degrees of freedom 
-in p-value hacking \cite{wicherts2016degrees}.
-
-% \paragraph{Metrics help assess the honest division of labour}
-
-Most code hosts, among others, GitHub and GitLab,
-automatically keep track of certain metrics.
-One such metric is the amount of code each contributor has made.
-In that way, authors involved in the writing of the software
-are honestly acknowledged. 
-\iffalse
-Figure \ref{fig:daisie_contributors}
-shows the number of commits the authors of DAISIE \cite{etienne2020daisie}
-have made to its code.
-\fi
-The distribution of contributed code per author 
-is useful metadata, that can automatically
-be acquired when using code hosting website.
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\linewidth]{daisie_contributors.png}
-  \caption{
-    An overview of the commits that multiple contributors have
-    made, in this case, to the code of DAISIE \cite{etienne2020daisie}
-  }
-  \label{fig:daisie_contributors}
-\end{figure}
-\fi
-
-% \paragraph{Suggestions for knowledge management}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Suggestions for knowledge management}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Making code FAIR.
+
+\subsection{Hosted code}
+
+\subsection{CI code}
+
+\subsection{Code in container}
 
@@
 
-Code used in academic research complies badly with FAIR principles,
-where unpublished code is the worst offender.
-However, any idealistic researcher does not have the
-tools to following FAIR principles in an exemplary way:
-there are no standards in how a scholarly article
-refers to academic code (i.e. in a machine-friendly way).
-Creating a (Singularity) container is the best a researcher
-can do to ensure his/her code is reusable,
-yet, there are no incentives to make code accessible, interoperable
-or re-usable, with standarized metadata attributes being absent.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For research to be reproducible, one ideally has access to
+both the data used and the code.
+In some fields, such as genetic epidemiology, the data are
+sensitive, hence cannot be released,
+yet there are methods being devised to run code on sensitive
+data with assured privacy \cite{zhang2016review,azencott2018machine}.
 
@@
 
-For research to be reproducible, one ideally has access to
-both the data used and the code.
-In some fields, such as genetic epidemiology, the data are
-sensitive, hence cannot be released,
-yet there are methods being devised to run code on sensitive
-data with assured privacy \cite{zhang2016review,azencott2018machine}.
-Additionally, it follows software development's best practices 
-(especially the practice to test code automatically) 
-to release a simulated/public dataset, with
-the additional benefit that this dataset can be used for comparisons.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 over time. 
-The degradation of software is a known feature for nearly 
-four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
-here a more modern (and in this context superior)
-definition is used: software collapse \cite{hinsen2019dealing},
-in which software fails due to dependencies on other 
-software.
-Where that paper recommends using older software foundations 
-with multiple implementations \cite{hinsen2019dealing},
-this paper recommends a modern solution that potentially has more 
-longevity:
-to publish the code installed in a virtual environment
-(including the dependencies used at that time), as described below.
 Write code that will stand the test of time perfectly
@@
 to increase transparency in the handling of bug reports.
+Should there be a retraction notice on code?
 
@@
 at the cost of a scientific career \cite{baggerly2009deriving}.
-
-% \paragraph{The drawbacks of publishing version-controlled code}
-
-Is it worth it to publish version-controlled code?
-For an author, 
-there is additional training involved, and also here,
-publishing code alongside an experiment opens up
-the possibily to receive questions regarding that code,
-as well as other interactions, including helpful contributions.
-
-% \paragraph{The drawbacks of publishing a running version of code}
-
-Is it worth it to archive running versions of code?
-For an author, 
-there is additional training involved.
-There is no FAIR infrastructure for Singularity containers.
-
-% % \paragraph{Still need the same data to do a reproducible experiment}
-
-% % \paragraph{Recommendations}
-
-Being able to re-do an experiment is a core principle of the scientific method.
-Publishing only a paper about a computational experiment is not enough,
-as the results are too likely to mismatch that English description.
-Journal should make it mandatory for authors
-to publish their code alongside a computational experiment.
-Authors should follow the FAIR principes for their code as well,
-as can be done using the infrastructure as supplied by, 
-for example, GitHub and GitLab, opening up new angles in
-metascience regarding computational experiments.
-Knowledge managers should create the infrastructure for the preservation
-of runnable experiments, to allow scientist to upload a Singularity
-container, so these are as well following the FAIR principles.
 
@@
 make paradata useful for data re(use):
-paradata should be comprehensive, comprehensively documented
-and friendly to computers \cite{huvila2022improving}.
+paradata should be extensive, comprehensively documented,
+with the creation of documentation and code going hand-in-hand,
+as well as friendly to computers \cite{huvila2022improving}.
+
+Before applying these features, the first step is to publish 
+the code.
+
 When applying these general recomendations to code, 
@@
 The documentation should be as extensive as recommended by the 
-software development literature. To be computer friendly,
+software development literature and have co-evolved with the
+code following the best practices in literate programming. 
+To be computer friendly,
 code should be able to run on at least one independent entity,
 such as a GitHub Actions server or from a Singularity container.
+
+The second step for code to be useful paradata is to add CI, 
+to stand the test of time better.
+The ultimate step is to publish code in a container.
 
@@
 partly as containers follow FAIR data principles only poorly.
-Runnable code is the pinnacle of reproducible research.
+However, such runnable code is the current pinnacle of reproducible research.
 
@@
 gradually as the skills of the author(s) progress.
+To quote \cite{barnes2010publish}: Publish your code, it is good enough.
 

### 2023-02-12 article.tex
Improve order
Review

@@
 The concept of paradata (although not named as such yet) 
-was introduced by \cite{couper1998measuring},
+was introduced by Couper and colleagues,
 who developed a computer-assisted interview program
@@
 measures the time used to answer each question 
-and even the time the monitor is turned off.
+and even the time the monitor is turned off \cite{couper1998measuring}.
 The goal in that context was to assess and improve survey quality.
@@
 \cite{huvila2022improving}.
+This paper discusses these general reasons applied to code.
 
@@
 to redo the analysis \cite{obels2020analysis}.
-For articles in Science magazine (years 2011-2012), 12\% of 204
-studies published the code \cite{stodden2018empirical}.
+For articles in Science magazine, 12\% of 204
+studies published the code \cite{stodden2018empirical} 
+(note that these were years 2011-2012).
 Unpublished code has been the cause of some saddening examples,
@@
 some real examples) can come across as so caustic, 
-that one is excused from not contacting an a corresponding author.
+that one may be excused from not contacting an a corresponding author.
 
@@
 caused by a decline in the production of lactose-degrading enzymes,
-and is most commonly found in south-east asia and south africa \cite{storhaug2017country}.
+and is most commonly found in south-east Asia and south Africa \cite{storhaug2017country}.
 The trait is caused by the genetic make-up, or 'genotype'
@@
 When an association between genotype and phenotype is found,
-these associations are subsequently used to 
+these associations, when relevant enough, are used to 
 create so-called 'gene panels', 
@@
 called adenosine, cytosine, guanine and thyrosine, 
-commonly abreviated as A, C, G and T respectively
+commonly abreviated as A, C, G and T 
 \iffalse
@@
 \fi
-.
+respectively.
 
 One SNP example is \verb|rs12133641|, which is a SNP located at position 
-154,428,283 (it is at the 154 millionth position, which happens to be located
-on the first chromosome), 
+154,428,283 (that is, at the 154 millionth nucleotide), 
 where 67 percent of the people within this study have an A,
 and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).
-From this it follow (assuming the nucleotides act independently)
+From this it follow (assuming the nucleotides are inherited independently)
 that 45\% of subjects have the genotype AA, 44\% have AG and 11\% have GG.
@@
 sake of simplicity it is assumed that such a rate is constant,
-in practice there are complex regulation mechanisms).
+yet, in practice there are complex regulation mechanisms).
 Some proteins end up in the blood and
@@
 correlations between genetic data and biological traits.
-For example, \cite{ahsan2017relative} shows that
+For example, Ahsan and colleagues show that
 SNP \verb|rs12133641| is highly correlated (p-value is $3.0^{-73}$
@@
 ,
-for $n$ = 961 individuals) with protein IL6RA.
+for $n$ = 961 individuals) with protein IL6RA \cite{ahsan2017relative}.
 The direction of the association is also concluded:
-the more guanines (as opposited to adenines) are present at that SNPs location,
+the more guanines are present at that SNPs location,
 the higher concentration of IL6RA can be found in a human's blood
@@
 with its own dedicated field of research.
-Both studies assumed a correct inference of DNA.
+Both studies assumed a correctly calculated DNA sequence.
 This means that if the DNA sequence analysis contained a software bug, 
@@
 with labs that embrace reproducibility (for example, \cite{barba2016hard})
-are the exception.
+being the exception.
 One suggested way forward is to make the reproduction of 
@@
 A genetic epidemiologist works with sensitive data as well:
-the genetic sequences of participants are private (\cite{clayton2019law}).
+the genetic sequences of participants are private \cite{clayton2019law}.
 For research to be reproducible one needs both the code and the data
@@
 important is the content: an unpaired one-tailed T test is performed.
-Taking a look at the (in this case, the programming language R) code, 
+Taking a look at the (in this case, the programming language R \cite{r}) code, 
 we find the following line:
@@
 Note that there is another side of this same coin:
-correctly reporting the T-test in a reproducable ways
+correctly reporting the t-test in a reproducable ways
 is known to be a problem, with only 80 out of a selected 179 papers
-fully reporting the parity of T-test (i.e. paired or unpaired) 
+fully reporting the parity of t-test (i.e. paired or unpaired) 
 \cite{weissgerber2018we}.
@@
 as it is the code that generates the results.
-When humans are fallible and code gets bigger, the likelihood of
+As humans are fallible and code gets bigger, the likelihood of
 a mismatch between the English paper and the code increases.
@@
 
-Preserving code in a reliable way is rare \cite{barnes2010publish}.
-In general, within software development, 
-changes at all levels are common
-and embracing those changes is recommended \cite{beck2000extreme}
-and useful code does so.
+Code is rarely preserved \cite{barnes2010publish}.
+If it is, keeping it running is even more rare.
 This section discusses the preservation of code for a short, medium
@@
 
-\subsection{Code hosting}
+\subsection{Code hosting}\label{subsec:code-hosting}
 
@@
 with GitHub being the most popular one \cite{cosentino2017systematic}.
-The \LaTeX code of this chapter is hosted on GitHub 
-as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
+For example, the \LaTeX code of this chapter is hosted on GitHub 
+as well, at \begin{sloppypar}\url{https://github.com/richelbilderbeek/chapter_paradata}\end{sloppypar}.
 The use of code hosting websites
@@
 
-Hosted code commonly has a commit history,
+Hosted code commonly keeps a history of file changes,
+a so-called commit history,
 where a commit is the description of the change, 
-as well as the changes in one or more file.
+as well as the changes in one or more files.
 For example, GitHub
@@
 as well as openly working on the code from the start \cite{jimenez2017four}.
-However, half of 56 Science magazine corresponding authors 
-does not have such a version control mechanism \cite{stodden2018empirical}.
-
-% \paragraph{A commit history helps assess honest reporting}
-
-Hosted code improves transparency \cite{gorgolewski2016practical},
-as, among others, the history of changes (i.e. the git commit history), 
-helps to assess the honesty in the reporting of the code. 
-Ideally, the full version history of the code 
-has been generated, by hosting the code directly at its inception,
-so that the ontology of the experiment can be viewed.
-However, in some repositories, only the final version of the code has been
-posted, which means all historical information is lost.
-The code history has many important properties,
-as, for example, 
-it can be observed how much statistical tests have really/actually been done 
-and see if all are reported:
-failing to reporting a (usually non-significant) statistical tests is 
-one of the many degrees of freedom 
-in p-value hacking \cite{wicherts2016degrees}.
+Half of 56 Science magazine corresponding authors 
+have such a version control mechanism \cite{stodden2018empirical}.
 
@@
 
-However, if the code breaks tomorrow, it will go unnoticed.
+A code host, however, will not actively preserve the uploaded code:
+if the code breaks tomorrow, it will go unnoticed.
 
@@
 To preserve code for a longer timespan, 
-continuous integration allows one to 
+continuous integration ('CI') allows one to 
 verify if code still works and, if not, to be notified on the need for an update.
+Things change: from operating systems, to libraries that code uses,
+to the code itself. 
+Embracing those changes is recommended \cite{beck2000extreme} 
+and useful code does so.
 
@@
 can immediatly see which small change was the cause of this.
-A formalized version of this workflow is called continuous integration (CI).
 Using a CI service (such as GitHub Actions for GitHub), 
@@
 \fi
+
+% \paragraph{Ideal paradata}
+
+For paradata to be useful, it has to be computer-friendly, yet 
+'the best paradata does not necessarily look like 'data' at all for its
+human users' \cite{huvila2022improving}.
+There are features of code that humans find useful,
+without directly being able to measure these.
+In the end, code is just 'another kind of data' and should be designed 
+as such, for example, by using tools to work on it \cite{wilson2022twelve}.
+
+% \paragraph{Using a tool for coding style}
+
+A first example is to use a tool to enforce a coding style 
+(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
+or PEP 8 \cite{van2001pep} for Python),
+as following a consistent coding style improves software quality \cite{fang2001}.
+
+% \paragraph{Using a tool for complexity}
+
+A second example is to use a tool to enforce a low cyclomatic complexity.
+The cyclomatic complexity is approximately defined 
+as the number of independent paths that
+the code can be executed. 
+For example, code that has only one \verb|if| statement
+has a cyclomatic complexity of 2, as the condition within the \verb|if|
+statement can be true or false,
+resulting in the body of the \verb|if| statement being either
+executed or ignored.
+The cyclomatic complexity correlates with code complexity,
+where more complex code is likelier to contain or give rise to bugs 
+\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}.
 
@@
 
-% \paragraph{Ideal paradata}
-
-For paradata to be useful, it has to be computer-friendly, yet 
-'the best paradata does not necessarily look like 'data' at all for its
-human users' \cite{huvila2022improving}.
-There are features of code that humans find useful,
-without directly being able to measure these.
-In the end, code is just 'another kind of data' and should be designed 
-as such, for example, by using tools to work on it \cite{wilson2022twelve}.
-
-% \paragraph{Using a tool for coding style}
-
-A first example is to use a tool to enforce a coding style 
-(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
-or PEP 8 \cite{van2001pep} for Python),
-as following a consistent coding style improves software quality \cite{fang2001}.
-
-% \paragraph{Using a tool for complexity}
-
-A second example is to use a tool to enforce a low cyclomatic complexity.
-The cyclomatic complexity is approximately defined 
-as the number of independent paths that
-the code can be executed. 
-For example, code that has only one \verb|if| statement
-has a cyclomatic complexity of 2, as the condition within the \verb|if|
-statement can be true or false,
-resulting in the body of the \verb|if| statement being either
-executed or ignored.
-The cyclomatic complexity correlates with code complexity,
-where more complex code is likelier to contain or give rise to bugs 
-\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}.
-
 Such tools combine great with CI services.
+
+Using CI does not prevent code from breaking. Instead,
+it allows one to be the first to know if code breaks.
 
@@
 \fi
-,
-however, that article was written before virtualization existed.
+(however, that article was written before virtualization existed).
 Virtualization allows one to provide a standard environment,
@@
 
-% \paragraph{Containers allow for runnable code, for longer periods of time}
-
-Unlike English, code is fragile in standing the test of time.
-Containers can alleviate this, as these can be created at a point
-in time and then run later. Because the container is stand-alone,
-all software that the code depends on remain in their current/old state.
-As with hosted code, an experiment may not run with newer versions
-of libraries anymore, but a containerized experiment can run regardless, 
-if one refrains from updating (or only does so when there are there problems with fixing).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 However, code has multiple uses, which have different requirements
-for being usesful paradata.
+for being useful paradata.
+
 Code can be used to reproduce, to replicate and to 
-extend a computational experiment \cite{benureau2018re},
+extend a computational experiment \cite{benureau2018re} and
 each of these uses has different requirements for the code to be useful paradata.
@@
 
-\subsubsection{... to reproduce an experiment}
-
-% \paragraph{Code should be hosted}
-
-Code should be hosted.
-
-For code to be extensive enough to reproduce an existing experiment,
-it should be distributed in standard ways \cite{peng2006reproducible}
+Code should be distributed in standard ways \cite{peng2006reproducible}
 (see section \ref{sec:code-availability}
 for the general availability of code).
-The standard way to publish code is to use a code hosting website.
-
-\subsubsection{... to replicate an experiment}
+The standard way to publish code is to use a code hosting website
+(see subsection \ref{subsec:code-hosting}).
 
@@
 Error handling is one of the mechanisms to do so.
-One example is, again, when doing a t-test:
+One example is when doing a t-test:
 to compare the means of two distributions, each distribution
@@
 The analytical code should clearly signal this problem.
-
-\subsubsection{... to extend an experiment}
 
@@
 show a certain R package's code has a 100\% code coverage.
+
+% \paragraph{Test-driven development}
+
+Having a 100\% code coverage is not uncommon,
+as there is a common software development practice that 
+strives for this, called test-driven development ('TDD'). 
+TDD is a software development strategy in which tests are written
+before new 'real' code is written.
+TDD improves code quality \cite{alkaoud2018quality,janzen2006test}
+and writing comment and documentation mixes well \cite{shmerlin2015document}.
 
@@
 
-% \paragraph{Test-driven development}
-
-TDD improves code quality \cite{alkaoud2018quality,janzen2006test}
-and writing comment and documentation mixes well \cite{shmerlin2015document}.
-
 % \paragraph{Code must be reviewed to ensure the correct amount of documentation}
@@
 are different, such as the random numbers as generated by an operating
-system, or data that are downloaded from online (and hence, changing) sources.
+system, or data that are downloaded from online (and hence, probably changing) sources.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Suggestions for knowledge management}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Making code FAIR.
-
-\subsection{Hosted code}
-
-\subsection{CI code}
-
-\subsection{Code in container}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sensitive data}\label{sec:sensitive-data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For research to be reproducible, one ideally has access to
+both the data used and the code.
+In some fields, such as genetic epidemiology, the data are
+sensitive, hence cannot be released,
+yet there are methods being devised to run code on sensitive
+data with assured privacy \cite{zhang2016review,azencott2018machine}.
+
+Next to the code, it is the data used in an experiment 
+that must be made available for an experiment 
+to be called 'reproducible' \cite{peng2006reproducible}.
+When the input data are sensitive, however, and thus cannot be shared with
+an independent researcher, one thus cannot reproduce the
+experiment.
+
+The solution for this problem is simple:
+supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
+together with the code.
+In the case of genetic epidemiology this would mean
+simulated genotypes and associated phenotypes,
+as can be done with the \verb|plinkr| R package \cite{plinkr}.
+This solution, however, is known to all developers that follow
+good practices when writing code, as public (i.e. simulated) 
+datasets are needed for automatic CI testing of code.
+One extra benefit of simulated data is that these can be used
+as a benchmark, as slightly different analyses should give 
+similar conclusions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In a perfect world, all code has the characteristis of ideal paradata
+and is written from software development best practices.
+This section discusses the problems that arise by doing so.
+
+% \paragraph{Invest in training}
+
+One needs to be trained to know the best practices,
+before these can be applied. 
+Articles that suggest these best practices, however, claim that this initial
+investment pays off.
+Code reviews are a good way to accelerate the 
+learning of team members \cite{vable2021code}.
+
+% \paragraph{Code needs maintanance}
+
+Code needs maintanance:
+the libraries that are used by an experiment
+will change over time, inevitably causing the experiment to fail
+over time. 
+Code that will stand the test of time perfectly 
+is deemed 'impossible' \cite{benureau2018re}.
+
+% \paragraph{Authors can be contacted}
+
+There is another problem that authors of academic code face,
+which is that they may be contacted with a user 
+asking for technical report 
+(against which a 'no' is just fine \cite{barnes2010publish})
+or -more relevant to the research- sending in a bug report.
+One solution for the author is to ignore such emails (as is done in a third
+of all cases \cite{teunis2015corresponding}):
+it can be argued that no energy should be wasted on published code
+and work on something new instead 
+(see \cite{barnes2010publish} for a better way to deal with the problem).
+
+% \paragraph{How to handle bugs?}
+
+However, upon the find of a new bug, the question is:
+does all the research that use that software still 
+result in the same conclusions?
+To prevent authors chosing to hide bug reports, 
+a code hosting website usually has a public place to report bugs,
+to increase transparency in the handling of bug reports.
+Should there be a retraction notice on code?
+
+% \paragraph{The problems with runnable code}
+
+There are some problems with providing a container with a runnable container.
+First, a container is several gigabytes and need to be stored somewhere
+online. 
+Ideally, for Singularity, this would be at the main container hosting
+site called Syslabs. Using Syslabs is free upon registration, 
+yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
+be entered, such as the URL of the article, nor the URL of hosted code.
+
+% \paragraph{Code in container}
 
@@
 or not metadata at all (Syslabs).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sensitive data}\label{sec:sensitive-data}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For research to be reproducible, one ideally has access to
-both the data used and the code.
-In some fields, such as genetic epidemiology, the data are
-sensitive, hence cannot be released,
-yet there are methods being devised to run code on sensitive
-data with assured privacy \cite{zhang2016review,azencott2018machine}.
-
-Next to the code, it is the data used in an experiment 
-that must be made available for an experiment 
-to be called 'reproducible' \cite{peng2006reproducible}.
-When the input data are sensitive, however, and thus cannot be shared with
-an independent researcher, one thus cannot reproduce the
-experiment.
-
-The solution for this problem is simple:
-supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
-together with the code.
-In the case of genetic epidemiology this would mean
-simulated genotypes and associated phenotypes,
-as can be done with the \verb|plinkr| R package \cite{plinkr}.
-This solution, however, is known to all developer that follow
-good practices when writing code, as public (i.e. simulated) 
-datasets are needed for automatic testing of code.
-One extra benefit of simulated data is that these can be used
-as a benchmark, as slightly different analyses should give 
-similar conclusions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In a perfect world, all code has the characteristis of ideal paradata
-and is written from software development best practices.
-This section discusses the problems that arise by doing so.
-
-% \paragraph{Invest in training}
-
-One needs to be trained to know the best practices,
-before these can be applied. 
-Articles that suggest these best practices, however, claim that this initial
-investment pays off.
-Code reviews are a good way to accelerate the 
-learning of team members \cite{vable2021code}.
-
-% \paragraph{Code needs maintanance}
-
-Code needs maintanance:
-the libraries that are used by an experiment
-will change over time, inevitably causing the experiment to fail
-over time. 
-Write code that will stand the test of time perfectly
-is deemed 'impossible' \cite{benureau2018re}.
-
-% \paragraph{Authors can be contacted}
-
-There is another problem that authors of academic code face,
-which is that they may be contacted with a user 
-asking for technical report 
-(against which a 'no' is just fine \cite{barnes2010publish})
-or -more relevant to the research- sending in a bug report.
-One solution for the author is to ignore such emails (as is done in a third
-of all cases \cite{teunis2015corresponding}):
-it can be argued that no energy should be wasted on published code
-and work on something new instead 
-(see \cite{barnes2010publish} for a better way to deal with the problem).
-
-% \paragraph{How to handle bugs?}
-
-However, upon the find of a new bug, the question is:
-does all the research that use that software still 
-result in the same conclusions?
-To prevent authors chosing to hide bug reports, 
-a code hosting website usually has a public place to report bugs,
-to increase transparency in the handling of bug reports.
-Should there be a retraction notice on code?
-
-% \paragraph{The problems with runnable code}
-
-There are some problems with providing a container with a runnable container.
-First, a container is several gigabytes and need to be stored somewhere
-online. 
-Ideally, for Singularity, this would be at the main container hosting
-site called Syslabs. Using Syslabs is free upon registration, 
-yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
-be entered, such as the URL of the article, nor the URL of hosted code.
-Also, journals do not request hosted code, nor runnable code.
 

### 2023-02-13 article.tex
Fix conclusion

@@
 this list can be phrased more precisely:
-code should be comprehensive in supplying 
-automatically generated metadata (such as commit history and code coverage).
-The documentation should be as extensive as recommended by the 
-software development literature and have co-evolved with the
+
+\begin{enumerate}
+  \item Code should be comprehensive 
+    in supplying automatically generated metadata (such as commit history and code coverage).
+  \item The documentation should be as extensive as recommended by the 
+    software development literature.
+  \item The documentation should have co-evolved with the
     code following the best practices in literate programming. 
-To be computer friendly,
-code should be able to run on at least one independent entity,
-such as a GitHub Actions server or from a Singularity container.
-
-The second step for code to be useful paradata is to add CI, 
-to stand the test of time better.
-The ultimate step is to publish code in a container.
+  \item Code should be made machine-readable by, at least,
+    being uploaded to a code hosting website.
+    Ideally, the code is checked by CI and is put in a container.
+\end{enumerate}
+
+For the preservation of code, these are the recommendations:
+
+\begin{enumerate}
+  \item Uploading code to a code hosting website is better than
+    not publishing code at all.
+  \item Adding CI to code allows one to detect the day when that code 
+    does not run anymore.
+  \item Putting the code in a container is the best way to preserve code.
+\end{enumerate}
 
@@
 When research truly needs to be reproducible, putting the code 
-of an experiment into a container is a good solution, as it 
-can stand the test of time better.
+of an experiment into a container is today's best solution,
+as containers are the best solution to keep code running for the longest 
+amount of time.
 Creating such a container, however, requires more skill
 that -as of today- is not rewarded,
-partly as containers follow FAIR data principles only poorly.
-However, such runnable code is the current pinnacle of reproducible research.
+although an experiment put into a container 
+can be considered the pinnacle of reproducible research.
 
@@
 gradually as the skills of the author(s) progress.
-To quote \cite{barnes2010publish}: Publish your code, it is good enough.
+To quote: Publish your code, it is good enough \cite{barnes2010publish}.
 

### 2023-02-13 scripts/create.sh
Can create both PDFs

@@ -1,6 +1,7 @@
 #!/bin/bash
 #
-# Create the PDF 'article.pdf'
+# Create the PDF 'article.pdf' using the Springer class.
+# :warning: If there is a PDF present, it will be deleted!
 #
 # It does so by converting the TeX to PDF, 
 # as well as stitching some other PDFs to it
@@ -29,6 +30,8 @@ if [ ! -f "svind.ist" ]; then
   echo "'svind.ist' does not exist."
   exit 43
 fi
+
+rm -f article.pdf
 
 pdflatex header.tex
 bibtex header >/dev/null

### 2023-02-13 scripts/create_both.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+#
+# Create the PDFs 'article_normal.pdf' and 'article_springer.pdf'
+#
+# It does so by converting the TeX to PDF, 
+# as well as stitching some other PDFs to it
+#
+# Usage:
+#
+#   ./scripts/create_both.sh
+#
+
+rm article.pdf
+./scripts/create.sh
+mv article.pdf article_springer.pdf
+./scripts/create_normal.sh
+mv article.pdf article_normal.pdf

### 2023-02-13 scripts/create_normal.sh

@@ -1,6 +1,7 @@
 #!/bin/bash
 #
 # Create the PDF 'article.pdf'
+# :warning: If there is a PDF present, it will be deleted!
 #
 # It does so by converting the TeX to PDF, 
 # as well as stitching some other PDFs to it
@@ -9,6 +10,7 @@
 #
 #   ./scripts/create_normal.sh
 #
+rm -f article.pdf
 
 pdflatex header_normal.tex
 bibtex header_normal >/dev/null

### 2023-02-13 scripts/create_both.sh
Use 'header.tex' for both versions

@@ -1,17 +1,0 @@
-#!/bin/bash
-#
-# Create the PDFs 'article_normal.pdf' and 'article_springer.pdf'
-#
-# It does so by converting the TeX to PDF, 
-# as well as stitching some other PDFs to it
-#
-# Usage:
-#
-#   ./scripts/create_both.sh
-#
-
-rm article.pdf
-./scripts/create.sh
-mv article.pdf article_springer.pdf
-./scripts/create_normal.sh
-mv article.pdf article_normal.pdf

### 2023-02-13 scripts/create_normal.sh

@@ -1,19 +1,0 @@
-#!/bin/bash
-#
-# Create the PDF 'article.pdf'
-# :warning: If there is a PDF present, it will be deleted!
-#
-# It does so by converting the TeX to PDF, 
-# as well as stitching some other PDFs to it
-#
-# Usage:
-#
-#   ./scripts/create_normal.sh
-#
-rm -f article.pdf
-
-pdflatex header_normal.tex
-bibtex header_normal >/dev/null
-pdflatex header_normal.tex >/dev/null
-pdflatex header_normal.tex
-mv header_normal.pdf article.pdf

### 2023-02-13 article.tex
Use '\abstract' (instead of '\abstract*')
Make cite look correct, fix #11
Make figure 1 bigger, fix #9

@@
-\setcounter{chapter}{7}
 \chapter{How to make research code useful paradata}
@@
   \caption{
-    Left: general relation between data, paradata and metadata.
-    Right: the same relations specified for genetic epidemiology.
+    \textbf{Left}: general relation between data, paradata and metadata.
+    \textbf{Right}: the same relations specified for genetic epidemiology.
+    Input data: genetic data, phenotypic data and associations
+    found in earlier studies.
+    Results: associations between genetic data and phenotypic data.
+    Code: the computer code used in a computational experiment
+    Paper: the scholarly paper describing the (code of the) experiment
   }

### 2023-02-13 scripts/make_full_article.sh
Use script

@@ -1,0 +1,10 @@
+#!/bin/bash
+#
+# Usage:
+#
+#   ./scripts/make_full_article.sh
+#
+sed -i 's/^.iffalse$/% \iffalse/g' article.tex
+sed -i 's/^.fi$/% \fi/g' article.tex
+sed -i 's/^.iffalse$/% \iffalse/g' header.tex
+sed -i 's/^.fi$/% \fi/g' header.tex

### 2023-02-13 scripts/make_full_article.sh
Use script in GHA to make full article

@@ -4,7 +4,7 @@
 #
 #   ./scripts/make_full_article.sh
 #
-sed -i 's/^.iffalse$/% \iffalse/g' article.tex
-sed -i 's/^.fi$/% \fi/g' article.tex
-sed -i 's/^.iffalse$/% \iffalse/g' header.tex
-sed -i 's/^.fi$/% \fi/g' header.tex
+sed -i 's/^\\iffalse$/% \\iffalse/g' article.tex
+sed -i 's/^\\fi$/% \\fi/g' article.tex
+sed -i 's/^\\iffalse$/% \\iffalse/g' header.tex
+sed -i 's/^\\fi$/% \\fi/g' header.tex

### 2023-02-13 scripts/make_springer_article.sh
Can undo make full article

@@ -1,0 +1,10 @@
+#!/bin/bash
+#
+# Usage:
+#
+#   ./scripts/make_springer_article.sh
+#
+sed -i 's/^% \\iffalse$/\\iffalse/g' article.tex
+sed -i 's/^% \\fi$/\\fi/g' article.tex
+sed -i 's/^% \\iffalse$/\\iffalse/g' header.tex
+sed -i 's/^% \\fi$/\\fi/g' header.tex

### 2023-02-13 scripts/do_it.sh
Document scripts

@@ -1,6 +1,0 @@
-#!/bin/bash
-rm article.pdf        &&
-  ./scripts/edit.sh   &&
-  ./scripts/create.sh &&
-  ./scripts/view.sh 
-

### 2023-02-13 article.tex
Remove 'rowcolors'
Put example text in italics

@@
 
-\begin{verbatim}
+\textit{
 We compared the in-blood concentrations of IL6RA 
@@
 to be less than the average concentration in those with GG.
-\end{verbatim}
+}
 
@@
 \iffalse
-\rowcolors{2}{gray!10}{gray!40}
 \begin{table}[h]

### 2023-02-13 article.tex
URLs are pretty

@@
 For example, the \LaTeX code of this chapter is hosted on GitHub 
-as well, at \begin{sloppypar}\url{https://github.com/richelbilderbeek/chapter_paradata}\end{sloppypar}.
+as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
 The use of code hosting websites
@@
 This article and its metadata can be found at 
-\begin{sloppypar}\url{https://github.com/richelbilderbeek/chapter_paradata}\end{sloppypar}.
+\url{https://github.com/richelbilderbeek/chapter_paradata}.
 

### 2023-02-13 article.tex
Online paper has more figures

@@
 
-This paper has been produced following all recommendations in it:
+This book chapter has been produced following all recommendations in it:
 it's \LaTeX code is hosted on GitHub 
@@
 the effect of reviewers' feedback on a first manuscript.
+The website also supplies an alternative version of the same chapter
+with more figures.
 

### 2023-02-13 article.tex
Remove comments
Move definitions to supmat

@@
 genetic epidemiology
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Definitions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[h]
-  \begin{tabular}{lp{5cm}p{5cm}}
-    Term         & Definition                                                                   & Example                                                         \\
-    \hline
-    Code         & Body of text to be run by a computer                                         & The scripts run by a computational experiment                   \\
-    Data         & Individual facts, statistics, or items of information                        & The name of a SNP known to have a significant association       \\
-    Genotype     & The DNA allele at a certain location                                         & AA, AC, CG, GT, ...                                             \\
-    Paradata     & Data that describe the process of generating data                            & The code to conclude that a SNP has a significant association   \\
-    Phenotype    & How an organism looks like in the broadest sense                             & The concentration of IL6RA in the blood                         \\
-    Primary data & Data that is the foundation of an experiment                                 & The genotype of individuals                                     \\
-    Metadata     & Data that provide information about other data                               & The article that describes an experiment                        \\
-    Raw data     & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals         \\
-    SNP          & A named location within a genome                                             & \verb|rs12133641|                                               \\
-    Trait        & A phenotype                                                                  & The concentration of IL6RA in the blood                         
-  \end{tabular}
-  \caption{Terms used in this paper, and their definitions}
-  \label{tab:definitions}
-\end{table}
 
@@
 The trait is caused by the genetic make-up, or 'genotype'
-(see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
-of a person.
-The trait, also called 'phenotype' (again, see Table \ref{tab:definitions}), 
+(see Table \ref{tab:definitions} in the Supplementary Materials
+for the definitions of terms used in this paper), of a person.
+The trait, also called 'phenotype', 
 in this example is lactose intolerance at adult age,
@@
 
-% Duran et al., grounds for trust
-% [A] process is epistemically opaque relative to a cognitive agent X at time t
-% just in case X does not know at t all of the epistemically relevant elements of
-% the process (Humphreys 2009, 618)
-% 
-% On Computational reliability
-% 
-% (CR) if S’s believing p at t results from m, then S’s belief in p at t is justified.
-% where S is a cognitive agent, p is any truth-valued proposition related to the
-% results of a computer simulation, t is any given time, and m is a reliable com‑
-% puter simulation.
-%
-%
-% 4 Sources for Computational Reliabilism
-%
-% 1. Verification and validation methods
-% 2. Robustness analysis for computer simulations
-% 3. A history of (un)successful implementations
-% 4. Expert knowledge
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Definitions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[h]
+  \begin{tabular}{|l|p{5cm}|p{5cm}|}
+    \hline
+    Term         & Definition                                                                   & Example                                                         \\
+    \hline
+    Code         & Body of text to be run by a computer                                         & The scripts run by a computational experiment                   \\
+    Data         & Individual facts, statistics, or items of information                        & The name of a SNP known to have a significant association       \\
+    Genotype     & The DNA allele at a certain location                                         & AA, AC, CG, GT, ...                                             \\
+    Paradata     & Data that describe the process of generating data                            & The code to conclude that a SNP has a significant association   \\
+    Phenotype    & How an organism looks like in the broadest sense                             & The concentration of IL6RA in the blood                         \\
+    Primary data & Data that is the foundation of an experiment                                 & The genotype of individuals                                     \\
+    Metadata     & Data that provide information about other data                               & The article that describes an experiment                        \\
+    Raw data     & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals         \\
+    SNP          & A named location within a genome                                             & \verb|rs12133641|                                               \\
+    Trait        & A phenotype                                                                  & The concentration of IL6RA in the blood                         
+    \hline
+  \end{tabular}
+  \caption{Terms used in this paper, and their definitions}
+  \label{tab:definitions}
+\end{table}

### 2023-02-13 article.tex
Remove recommendations to supmat
Fix table

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This paper started with some suggestions to 
+make paradata useful for data re(use):
+paradata should be extensive, comprehensively documented,
+with the creation of documentation and code going hand-in-hand,
+as well as friendly to computers \cite{huvila2022improving}.
+
+Before applying these features, the first step is to publish 
+the code.
+
+When applying these general recomendations to code, 
+this list can be phrased more precisely:
+
+\begin{enumerate}
+  \item Code should be comprehensive 
+    in supplying automatically generated metadata (such as commit history and code coverage).
+  \item The documentation should be as extensive as recommended by the 
+    software development literature.
+  \item The documentation should have co-evolved with the
+    code following the best practices in literate programming. 
+  \item Code should be made machine-readable by, at least,
+    being uploaded to a code hosting website.
+    Ideally, the code is checked by CI and is put in a container.
+\end{enumerate}
+
+For the preservation of code, these are the recommendations:
+
+\begin{enumerate}
+  \item Uploading code to a code hosting website is better than
+    not publishing code at all.
+  \item Adding CI to code allows one to detect the day when that code 
+    does not run anymore.
+  \item Putting the code in a container is the best way to preserve code.
+\end{enumerate}
+
+% \paragraph{Final goal: runnable code}
+
+When research truly needs to be reproducible, putting the code 
+of an experiment into a container is today's best solution,
+as containers are the best solution to keep code running for the longest 
+amount of time.
+Creating such a container, however, requires more skill
+that -as of today- is not rewarded,
+although an experiment put into a container 
+can be considered the pinnacle of reproducible research.
+
+% \paragraph{First step: publish code alongside papers}
+
+The simplest and most impactful step to make code more useful paradata
+is, however, to publish it on a code hosting website 
+along a publication. From then on, the next steps can be taken 
+gradually as the skills of the author(s) progress.
+To quote: Publish your code, it is good enough \cite{barnes2010publish}.
+
+% \paragraph{Final word}
+
+The world of science would be a more open, humble, trustworthy, truthful
+and helpful would the code that accompanies a scientific paper
+be treated like a first class citizen. 
+Doing so, however, is yet to be rewarded
+and still both of the two scientists at the start of this paper 
+can provide a good rationale for their behavior.
+This will change when reward incentives are put into place 
+that reward making paradata useful.
+For code specifically, in any computational field,
+the rewards are even higher, as reproducibility should again be 
+a cornerstone in science.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Data Accessibility}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This article and its metadata can be found at 
+\url{https://github.com/richelbilderbeek/chapter_paradata}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Bibliography
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Vancouver style
+\bibliographystyle{unsrtnat}
+\bibliography{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\appendix
+\section{Supplementary materials}
+
+% Figures start from one and are prepended with an S
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+% Tables start from one and are prepended with an S
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{table}{0}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{This paper}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This book chapter has been produced following all recommendations in it:
+it's \LaTeX code is hosted on GitHub 
+at \url{https://github.com/richelbilderbeek/chapter_paradata}.
+Due this, one can, among others, see the complete version history of this paper.
+This alone could allow research to be done on, for example,
+the effect of reviewers' feedback on a first manuscript.
+The website also supplies an alternative version of the same chapter
+with more figures.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Funding}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+This chapter has not been supported by funding.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Definitions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[h]
+  \begin{tabular}{|l|p{5cm}|p{5cm}|}
+    \hline
+    Term         & Definition                                                                   & Example                                                         \\
+    \hline
+    Code         & Body of text to be run by a computer                                         & The scripts run by a computational experiment                   \\
+    Data         & Individual facts, statistics, or items of information                        & The name of a SNP known to have a significant association       \\
+    Genotype     & The DNA allele at a certain location                                         & AA, AC, CG, GT, ...                                             \\
+    Paradata     & Data that describe the process of generating data                            & The code to conclude that a SNP has a significant association   \\
+    Phenotype    & How an organism looks like in the broadest sense                             & The concentration of IL6RA in the blood                         \\
+    Primary data & Data that is the foundation of an experiment                                 & The genotype of individuals                                     \\
+    Metadata     & Data that provide information about other data                               & The article that describes an experiment                        \\
+    Raw data     & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals         \\
+    SNP          & A named location within a genome                                             & \verb|rs12133641|                                               \\
+    Trait        & A phenotype                                                                  & The concentration of IL6RA in the blood                         \\
+    \hline
+  \end{tabular}
+  \caption{Terms used in this paper, and their definitions}
+  \label{tab:definitions}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Recommendations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \iffalse
@@
 \fi
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This paper started with some suggestions to 
-make paradata useful for data re(use):
-paradata should be extensive, comprehensively documented,
-with the creation of documentation and code going hand-in-hand,
-as well as friendly to computers \cite{huvila2022improving}.
-
-Before applying these features, the first step is to publish 
-the code.
-
-When applying these general recomendations to code, 
-this list can be phrased more precisely:
-
-\begin{enumerate}
-  \item Code should be comprehensive 
-    in supplying automatically generated metadata (such as commit history and code coverage).
-  \item The documentation should be as extensive as recommended by the 
-    software development literature.
-  \item The documentation should have co-evolved with the
-    code following the best practices in literate programming. 
-  \item Code should be made machine-readable by, at least,
-    being uploaded to a code hosting website.
-    Ideally, the code is checked by CI and is put in a container.
-\end{enumerate}
-
-For the preservation of code, these are the recommendations:
-
-\begin{enumerate}
-  \item Uploading code to a code hosting website is better than
-    not publishing code at all.
-  \item Adding CI to code allows one to detect the day when that code 
-    does not run anymore.
-  \item Putting the code in a container is the best way to preserve code.
-\end{enumerate}
-
-% \paragraph{Final goal: runnable code}
-
-When research truly needs to be reproducible, putting the code 
-of an experiment into a container is today's best solution,
-as containers are the best solution to keep code running for the longest 
-amount of time.
-Creating such a container, however, requires more skill
-that -as of today- is not rewarded,
-although an experiment put into a container 
-can be considered the pinnacle of reproducible research.
-
-% \paragraph{First step: publish code alongside papers}
-
-The simplest and most impactful step to make code more useful paradata
-is, however, to publish it on a code hosting website 
-along a publication. From then on, the next steps can be taken 
-gradually as the skills of the author(s) progress.
-To quote: Publish your code, it is good enough \cite{barnes2010publish}.
-
-% \paragraph{Final word}
-
-The world of science would be a more open, humble, trustworthy, truthful
-and helpful would the code that accompanies a scientific paper
-be treated like a first class citizen. 
-Doing so, however, is yet to be rewarded
-and still both of the two scientists at the start of this paper 
-can provide a good rationale for their behavior.
-This will change when reward incentives are put into place 
-that reward making paradata useful.
-For code specifically, in any computational field,
-the rewards are even higher, as reproducibility should again be 
-a cornerstone in science.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Accessibility}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This article and its metadata can be found at 
-\url{https://github.com/richelbilderbeek/chapter_paradata}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Vancouver style
-\bibliographystyle{unsrtnat}
-\bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-\section{Supplementary materials}
-
-% Figures start from one and are prepended with an S
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
-
-% Tables start from one and are prepended with an S
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{table}{0}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{This paper}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This book chapter has been produced following all recommendations in it:
-it's \LaTeX code is hosted on GitHub 
-at \url{https://github.com/richelbilderbeek/chapter_paradata}.
-Due this, one can, among others, see the complete version history of this paper.
-This alone could allow research to be done on, for example,
-the effect of reviewers' feedback on a first manuscript.
-The website also supplies an alternative version of the same chapter
-with more figures.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Funding}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-This chapter has not been supported by funding.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Definitions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[h]
-  \begin{tabular}{|l|p{5cm}|p{5cm}|}
-    \hline
-    Term         & Definition                                                                   & Example                                                         \\
-    \hline
-    Code         & Body of text to be run by a computer                                         & The scripts run by a computational experiment                   \\
-    Data         & Individual facts, statistics, or items of information                        & The name of a SNP known to have a significant association       \\
-    Genotype     & The DNA allele at a certain location                                         & AA, AC, CG, GT, ...                                             \\
-    Paradata     & Data that describe the process of generating data                            & The code to conclude that a SNP has a significant association   \\
-    Phenotype    & How an organism looks like in the broadest sense                             & The concentration of IL6RA in the blood                         \\
-    Primary data & Data that is the foundation of an experiment                                 & The genotype of individuals                                     \\
-    Metadata     & Data that provide information about other data                               & The article that describes an experiment                        \\
-    Raw data     & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals         \\
-    SNP          & A named location within a genome                                             & \verb|rs12133641|                                               \\
-    Trait        & A phenotype                                                                  & The concentration of IL6RA in the blood                         
-    \hline
-  \end{tabular}
-  \caption{Terms used in this paper, and their definitions}
-  \label{tab:definitions}
-\end{table}

### 2023-02-13 article.tex
No supmat in book chapter

@@
 The trait is caused by the genetic make-up, or 'genotype'
+\iffalse
 (see Table \ref{tab:definitions} in the Supplementary Materials
-for the definitions of terms used in this paper), of a person.
+for the definitions of terms used in this paper)
+\fi
+, of a person.
 The trait, also called 'phenotype', 
@@
 
+\iffalse
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\iffalse
 \begin{table}[h]
@@
 \end{table}
-\fi
+
+\fi

### 2023-02-13 article.tex
Do spell check

@@
 % Alex and Blake?
-Two different researchers in genetic epideomology (more on that field later), 
+Two different researchers in genetic epidemiology (more on that field later), 
 write two equally good manuscripts
@@
 paradata should be comprehensive, documented in a useful way, 
-the documention and data should have co-evolved, 
+the documentation and data should have co-evolved, 
 and the paradata should be computer-friendly \cite{huvila2022improving}.
@@
 42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
-Another study analysed the reproducibility of registered reports
+Another study analyzed the reproducibility of registered reports
 in the field of psychology, 
@@
 When code of an academic paper is not published, 
-one could contact the correspending author and request it.
-However, the reponse rate of correnspondence authors 
+one could contact the corresponding author and request it.
+However, the response rate of corresponding authors 
 with a request for data in computational fields is around 50\% 
@@
     Each of these chromosomes (46 in humans) consist out of DNA. 
-    DNA itself conists out of 4 nucleotides, 
+    DNA itself consists out of 4 nucleotides, 
     as depicted by the horizontal sticks 
@@
 called adenosine, cytosine, guanine and thyrosine, 
-commonly abreviated as A, C, G and T 
+commonly abbreviated as A, C, G and T 
 \iffalse
@@
   \caption{
-    An example result of a genetic epideological research.
+    An example result of a genetic epidemiological research.
     It shows that the SNP named rs12133641 (located at position 154,428,283
@@
 in which one SNP (e.g. phenylketonuria) or many SNPs (e.g. \cite{bruce2009metabolic}) 
-contribute to being more likely to develope a disease in the future. 
+contribute to being more likely to develop a disease in the future. 
 Here, the phenotype (having a disease in the future) 
@@
 Note that there is another side of this same coin:
-correctly reporting the t-test in a reproducable ways
+correctly reporting the t-test in a reproducible ways
 is known to be a problem, with only 80 out of a selected 179 papers
@@
 If the functionality of the code is lost, the researcher
-can immediatly see which small change was the cause of this.
+can immediately see which small change was the cause of this.
 Using a CI service (such as GitHub Actions for GitHub), 
@@
 Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
-The non-profit organisation rOpenSci \cite{ram2018community},
+The non-profit organization rOpenSci \cite{ram2018community},
 which peer-reviews code,
@@
 
-In a perfect world, all code has the characteristis of ideal paradata
+In a perfect world, all code has the characteristics of ideal paradata
 and is written from software development best practices.
@@
 
-Code needs maintanance:
+Code needs maintenance:
 the libraries that are used by an experiment
@@
 result in the same conclusions?
-To prevent authors chosing to hide bug reports, 
+To prevent authors choosing to hide bug reports, 
 a code hosting website usually has a public place to report bugs,
@@
 For an author, publishing code alongside an experiment opens up
-the possibily to receive questions regarding that code.
+the possibility to receive questions regarding that code.
 Note, however, that not publishing code may put 
@@
 
-When applying these general recomendations to code, 
+When applying these general recommendations to code, 
 this list can be phrased more precisely:
@@
     \hline
-    Hosted code   & Use a standarized badge for build status \\
-                  & Use a standarized badge for code coverage \\
+    Hosted code   & Use a standardized badge for build status \\
+                  & Use a standardized badge for code coverage \\
                   & Standardize to link between paper and code host \\
@@
                   & Website to host containers must allow for annotation \\
-                  & Use a standarized badge for the container to be functioning  \\
-                  & Standarize to link between container, paper and code repository \\
+                  & Use a standardized badge for the container to be functioning  \\
+                  & Standardize to link between container, paper and code repository \\
     \hline

### 2023-02-13 scripts/do_spell_check.sh
Done spell check, fix #12

@@ -1,0 +1,11 @@
+#!/bin/bash
+#
+# Do a spell check
+#
+# Usage:
+#
+#   ./scripts/do_spell_check.sh
+#
+# Solution adapted from https://tex.stackexchange.com/a/19
+
+aspell -c -t article.tex -d en

### 2023-02-13 scripts/do_spell_check.sh
Be more explicit

@@ -8,4 +8,4 @@
 #
 # Solution adapted from https://tex.stackexchange.com/a/19
 
-aspell -c -t article.tex -d en
+aspell -c --mode=tex article.tex --lang=en

### 2023-02-13 article.tex
Use Great Britain English

@@
 42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
-Another study analyzed the reproducibility of registered reports
+Another study analysed the reproducibility of registered reports
 in the field of psychology, 
@@
     as depicted by the horizontal sticks 
-    with the colors red, yellow, green and blue.
+    with the colours red, yellow, green and blue.
     From \cite{sponk2012}
@@
     line (the columns) for different proteins (the rows).
-    The colors above the cells, with the dendrogram, shows
+    The colours above the cells, with the dendrogram, shows
     red for cells that do not respond (red) and do respond (blue) to an 
@@
 
-\subsection{Containerization}
+\subsection{Containerisation}
 
@@
 \fi
-(however, that article was written before virtualization existed).
-Virtualization allows one to provide a standard environment,
+(however, that article was written before virtualisation existed).
+Virtualisation allows one to provide a standard environment,
 that can be used on any other computer.
@@
 \emph{not least because they shape each other} \cite{huvila2022improving}.
-This emphasized part of the quote resonates strongly with the idea of 
+This emphasised part of the quote resonates strongly with the idea of 
 literate programming \cite{knuth1984literate},
@@
 Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
-The non-profit organization rOpenSci \cite{ram2018community},
+The non-profit organisation rOpenSci \cite{ram2018community},
 which peer-reviews code,
@@
 and still both of the two scientists at the start of this paper 
-can provide a good rationale for their behavior.
+can provide a good rationale for their behaviour.
 This will change when reward incentives are put into place 
@@
     \hline
-    Hosted code   & Use a standardized badge for build status \\
-                  & Use a standardized badge for code coverage \\
-                  & Standardize to link between paper and code host \\
+    Hosted code   & Use a standardised badge for build status \\
+                  & Use a standardised badge for code coverage \\
+                  & Standardise to link between paper and code host \\
     \hline
@@
                   & Website to host containers must allow for annotation \\
-                  & Use a standardized badge for the container to be functioning  \\
-                  & Standardize to link between container, paper and code repository \\
+                  & Use a standardised badge for the container to be functioning  \\
+                  & Standardise to link between container, paper and code repository \\
     \hline

### 2023-02-13 scripts/do_spell_check.sh

@@ -8,4 +8,4 @@
 #
 # Solution adapted from https://tex.stackexchange.com/a/19
 
-aspell -c --mode=tex article.tex --lang=en
+aspell -c --mode=tex article.tex --lang=en_GB

### 2023-02-13 article.tex
No chapter in chapter, show abstract
Remove illustration of mismatch between code and text

@@
-\chapter{How to make research code useful paradata}
-
-\abstract*{
+\abstract{
 Paradata is data about the data collection process,
@@
 
+\iffalse
 To illustrate how easy it is to get a mismatch between a paper
@@
 the code (if published) would clarify this.
+\fi
 

### 2023-02-13 article.tex
Use '\cite' with or without braces

@@
 measures the time used to answer each question 
-and even the time the monitor is turned off \cite{couper1998measuring}.
+and even the time the monitor is turned off (\cite{couper1998measuring}).
 The goal in that context was to assess and improve survey quality.
@@
 This paper defines paradata as 'data about the data collection 
-process' \cite{choumert2019using}.
+process' (\cite{choumert2019using}).
 This definition is chosen for its simplicity, 
@@
 the documentation and data should have co-evolved, 
-and the paradata should be computer-friendly \cite{huvila2022improving}.
+and the paradata should be computer-friendly (\cite{huvila2022improving}).
 At first glance, these suggestions appear to work well for code
@@
 Finally, in computational fields, 
-it can help understand how scholarly knowledge is produced
-\cite{huvila2022improving}.
+it can help understand how scholarly knowledge is produced (\cite{huvila2022improving}).
 This paper discusses these general reasons applied to code.
@@
 However, it is not common to publish the code of an experiment or analysis 
-\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
+(\cite{stodden2011trust,read2015sizing}) (with a pleasant exception 
 being \cite{conesa2019making}).
@@
 a field intimately familiar with computer code,
-42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
+42\% of 454 SIGGRAPH papers supply computer code (\cite{bonneel2020code}).
 Another study analysed the reproducibility of registered reports
@@
 where 60\% of 62 studies supplied the code 
-to redo the analysis \cite{obels2020analysis}.
+to redo the analysis (\cite{obels2020analysis}).
 For articles in Science magazine, 12\% of 204
-studies published the code \cite{stodden2018empirical} 
+studies published the code (\cite{stodden2018empirical}) 
 (note that these were years 2011-2012).
@@
 better than a human expert, 
-yet failed to ever be reproduced \cite{haibe2020importance}.
+yet failed to ever be reproduced (\cite{haibe2020importance}).
 
@@
 with a request for data in computational fields is around 50\% 
-\cite{manca2018non, stodden2018empirical, teunis2015corresponding} 
+(\cite{manca2018non, stodden2018empirical, teunis2015corresponding}) 
 (the field of emergency medicine seems to be a pleasant outlier, 
-with 73\% of 118 emails being replied \cite{o2003email}).
+with 73\% of 118 emails being replied (\cite{o2003email})).
 When getting an answer of the corresponding author, 
 48\% out of 134 replied emails will actually result 
-in a sharing of the code \cite{stodden2018empirical}.
+in a sharing of the code (\cite{stodden2018empirical}).
 The responses of unwilling authors (see \cite{stodden2018empirical} for 
@@
 caused by a decline in the production of lactose-degrading enzymes,
-and is most commonly found in south-east Asia and south Africa \cite{storhaug2017country}.
+and is most commonly found in south-east Asia and south 
+Africa (\cite{storhaug2017country}).
 The trait is caused by the genetic make-up, or 'genotype'
@@
 from a population study called the Northern Swedish Population
-Health Study (NSPHS) that started in 2010 \cite{igl2010northern}. 
+Health Study (NSPHS) that started in 2010 (\cite{igl2010northern}). 
 The approximately 1000 participants were initially mostly surveyed
-about lifestyle \cite{igl2010northern} and follow-up studies
+about lifestyle (\cite{igl2010northern}) and follow-up studies
 provided the type of data relevant for this paper, 
-which are (1) the genotypes \cite{johansson2013identification},
+which are (1) the genotypes (\cite{johansson2013identification}),
 (2) the phenotypes, in this case, concentrations of certain proteins in the 
-blood \cite{enroth2014strong,enroth2015effect}.
+blood (\cite{enroth2014strong,enroth2015effect}).
 
@@
 ,
-for $n$ = 961 individuals) with protein IL6RA \cite{ahsan2017relative}.
+for $n$ = 961 individuals) with protein IL6RA (\cite{ahsan2017relative}).
 The direction of the association is also concluded:
@@
 One suggested way forward is to make the reproduction of 
-research a minimal requirement for publication \cite{peng2011reproducible}.
+research a minimal requirement for publication (\cite{peng2011reproducible}).
 
@@
 A genetic epidemiologist works with sensitive data as well:
-the genetic sequences of participants are private \cite{clayton2019law}.
+the genetic sequences of participants are private (\cite{clayton2019law}).
 For research to be reproducible one needs both the code and the data
@@
 important is the content: an unpaired one-tailed T test is performed.
-Taking a look at the (in this case, the programming language R \cite{r}) code, 
+Taking a look at the (in this case, the programming language R (\cite{r})) code, 
 we find the following line:
@@
 fully reporting the parity of t-test (i.e. paired or unpaired) 
-\cite{weissgerber2018we}.
+(\cite{weissgerber2018we}).
 In those unresolved 99 papers, 
@@
 simple (and may legitimately fear that those simple errors are common) 
-\cite{baggerly2009deriving, vable2021code},
+(\cite{baggerly2009deriving, vable2021code}),
 and this example is not too far off from the
@@
 
-Code is rarely preserved \cite{barnes2010publish}.
+Code is rarely preserved (\cite{barnes2010publish}).
 If it is, keeping it running is even more rare.
@@
 There are multiple code hosting websites, 
-with GitHub being the most popular one \cite{cosentino2017systematic}.
-For example, the \LaTeX code of this chapter is hosted on GitHub 
+with GitHub being the most popular one (\cite{cosentino2017systematic}).
+For example, the \LaTeX \hspace{5mm} code of this chapter is hosted on GitHub 
 as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
 The use of code hosting websites
-has increased strongly \cite{russell2018large},
-accommodates collaboration \cite{perez2016ten}
-and improves transparency \cite{gorgolewski2016practical},
+has increased strongly (\cite{russell2018large}),
+accommodates collaboration (\cite{perez2016ten})
+and improves transparency (\cite{gorgolewski2016practical}),
 due to its inherent computer-friendliness.
@@
 \iffalse
-An example from the author is the website \cite{bbbqarticleissue157},
+An example from the author is the website (\cite{bbbqarticleissue157}),
 that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
@@
 It is a general recommendation to put version control
-on all human-produced data \cite{wilson2014best},
-as well as openly working on the code from the start \cite{jimenez2017four}.
+on all human-produced data (\cite{wilson2014best}),
+as well as openly working on the code from the start (\cite{jimenez2017four}).
 Half of 56 Science magazine corresponding authors 
-have such a version control mechanism \cite{stodden2018empirical}.
+have such a version control mechanism (\cite{stodden2018empirical}).
 
@@
 Figure \ref{fig:daisie_contributors}
-shows the number of commits the authors of DAISIE \cite{etienne2020daisie}
+shows the number of commits the authors of DAISIE (\cite{etienne2020daisie})
 have made to its code.
@@
 to the code itself. 
-Embracing those changes is recommended \cite{beck2000extreme} 
+Embracing those changes is recommended (\cite{beck2000extreme})
 and useful code does so.
@@
 is known to significantly 
-increase the number of bugs exposed \cite{vasilescu2015} and increases
-the speed at which new features are added \cite{vasilescu2015}.
+increase the number of bugs exposed (\cite{vasilescu2015}) and increases
+the speed at which new features are added (\cite{vasilescu2015}).
 Using CI opens up the possibility to informally annotate features of
@@
 In the end, code is just 'another kind of data' and should be designed 
-as such, for example, by using tools to work on it \cite{wilson2022twelve}.
+as such, for example, by using tools to work on it (\cite{wilson2022twelve}).
 
@@
 A first example is to use a tool to enforce a coding style 
-(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
-or PEP 8 \cite{van2001pep} for Python),
-as following a consistent coding style improves software quality \cite{fang2001}.
+(e.g. the Tidyverse style guide (\cite{wickham2019advanced}) for R,
+or PEP 8 (\cite{van2001pep}) for Python),
+as following a consistent coding style improves software quality (\cite{fang2001}).
 
@@
 the code can be executed. 
+\iffalse
 For example, code that has only one \verb|if| statement
@@
 executed or ignored.
+\fi
 The cyclomatic complexity correlates with code complexity,
 where more complex code is likelier to contain or give rise to bugs 
-\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}.
+(\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}).
 
@@
 here a more modern (and in this context superior)
-definition is used: software collapse \cite{hinsen2019dealing},
+definition is used: software collapse (\cite{hinsen2019dealing}),
 in which software fails due to dependencies on other 
@@
 Where that paper recommends using older software foundations 
-with multiple implementations \cite{hinsen2019dealing},
+with multiple implementations (\cite{hinsen2019dealing}),
 this paper recommends a modern solution that potentially has more 
@@
 documented appropriately, documented in co-evolution with the data,
-and the data should be friendly to computers \cite{huvila2022improving}.
+and the data should be friendly to computers (\cite{huvila2022improving}).
 In this section, these ideal properties of paradata are applied to code.
@@
 Code can be used to reproduce, to replicate and to 
-extend a computational experiment \cite{benureau2018re} and
+extend a computational experiment (\cite{benureau2018re}) and
 each of these uses has different requirements for the code to be useful paradata.
@@
 For code to be ideal paradata, it must be usefully 
-documented \cite{huvila2022improving}.
+documented (\cite{huvila2022improving}).
 What is 'useful' is determined by its intended use: 
@@
 total there were 180 papers) have documented their code well enough to run it 
-and to be understood what it ought to do \cite{stodden2018empirical}.
+and to be understood what it ought to do (\cite{stodden2018empirical}).
 
@@
 However, all code in general should be documented 
-'adequately' \cite{peng2006reproducible},
+'adequately' (\cite{peng2006reproducible}),
 ideally writing code in such a way that it becomes 
-self-explanatory \cite{wilson2014best}
+self-explanatory (\cite{wilson2014best})
 and for the remaining code to document the reasons behind it, 
-its design and its purpose \cite{wilson2014best}.
+its design and its purpose (\cite{wilson2014best}).
 
@@
 The extent of investing time in documenting code is recommended
-to be proportional to the intended reuse \cite{pianosi2020successfully}
+to be proportional to the intended reuse (\cite{pianosi2020successfully})
 and there exists a clear relationship between the reuse 
 of code and its documentation 
-effort \cite{cosentino2017systematic,hata2015characteristics}.
+effort (\cite{cosentino2017systematic,hata2015characteristics}).
 
@@
 its creation and documentation need to align, 
-\emph{not least because they shape each other} \cite{huvila2022improving}.
+\emph{not least because they shape each other} (\cite{huvila2022improving}).
 This emphasised part of the quote resonates strongly with the idea of 
-literate programming \cite{knuth1984literate},
-in which documentation and code is developed hand-in-hand.
+literate programming (\cite{knuth1984literate}),
+in which documentation and code are developed hand-in-hand.
 Literate programming is the practice of writing a document mixed with code that 
@@
 Contemporary examples of this idea are, among others, 
-vignettes \cite{wickham2015r} (for the R programming language \cite{r}) 
-and Jupyter notebooks \cite{wang2020assessing} 
-(for the Julia \cite{Julia-2017}, Python \cite{van1995python} and R \cite{r}
-programming languages).
+vignettes (\cite{wickham2015r}) for the R programming language (\cite{r}) 
+and Jupyter notebooks (\cite{wang2020assessing})
+for the Julia (\cite{Julia-2017}), Python (\cite{van1995python}) and R (\cite{r})
+programming languages.
 rOpenSci, a community that, among others, reviews programming
@@
 code must have extensive runnable examples
-as part of the documentation \cite{ropensci_2021_6619350}.
+as part of the documentation (\cite{ropensci_2021_6619350}).
 Developing code and documentation at the same time 
-leads to both better code and documentation \cite{reenskaug1989environment}.
+leads to both better code and documentation (\cite{reenskaug1989environment}).
 In general, when developing software, 
@@
 to write documentation while writing software,
-as well as to include many examples \cite{lee2018ten}.
+as well as to include many examples (\cite{lee2018ten}).
 
@@
 
-Code should be distributed in standard ways \cite{peng2006reproducible}
+Code should be distributed in standard ways (\cite{peng2006reproducible})
 (see section \ref{sec:code-availability}
@@
 used on different data,
-as then 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
+as then 'code must act as a teacher for future developers' (\cite{sadowski2018modern}).
 Error handling is one of the mechanisms to do so.
@@
 Coding errors are extremely common 
-and contribute to the reproducibility crisis in science \cite{vable2021code}.
+and contribute to the reproducibility crisis in science (\cite{vable2021code}).
 Testing, in general, is an important mechanism to ensure
@@
 The percentage of (lines of) code tested is called the code coverage.
-Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
-The non-profit organisation rOpenSci \cite{ram2018community},
+Code coverage correlates with code quality (\cite{horgan1994,del1995correlation}). 
+The non-profit organisation rOpenSci (\cite{ram2018community}),
 which peer-reviews code,
@@
 before new 'real' code is written.
-TDD improves code quality \cite{alkaoud2018quality,janzen2006test}
-and writing comment and documentation mixes well \cite{shmerlin2015document}.
+TDD improves code quality (\cite{alkaoud2018quality,janzen2006test})
+and writing comment and documentation as part of the process
+is easy (\cite{shmerlin2015document}).
 
@@
 It is considered good practice
-to add a software license \cite{jimenez2017four},
+to add a software license (\cite{jimenez2017four}),
 so that it is clear that the software can be reused.
 About two-thirds of 56 Science magazine corresponding authors 
-that shared their code supply a software license \cite{stodden2018empirical}.
+that shared their code supply a software license (\cite{stodden2018empirical}).
 
@@
 
-In an informal Twitter poll,
-more than half of all 315 responding scientists 
-have their code rarely or never reviewed \cite{vable2021code},
+More than half of 315 scientists 
+have their code rarely or never reviewed (\cite{vable2021code}),
 although code reviews are known to
@@
 and resulting to an experiment that is likelier to be reproducible
-\cite{vable2021code}.
+(\cite{vable2021code}).
 Within software development best-practices, 
-code reviews are recommended as well \cite{wilson2014best} 
+code reviews are recommended as well (\cite{wilson2014best}) 
 The observation that code is reviewed, makes it likelier to be correct.
@@
 reproducible, one in three had the results reproduced with 
-minimal or no effort \cite{stodden2018empirical}.
+minimal or no effort (\cite{stodden2018empirical}).
 
@@
 that must be made available for an experiment 
-to be called 'reproducible' \cite{peng2006reproducible}.
+to be called 'reproducible' (\cite{peng2006reproducible}).
 When the input data are sensitive, however, and thus cannot be shared with
@@
 The solution for this problem is simple:
-supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
+supply a simulated (also called 'analytical' (\cite{peng2006reproducible})) dataset
 together with the code.
@@
 simulated genotypes and associated phenotypes,
-as can be done with the \verb|plinkr| R package \cite{plinkr}.
+as can be done with the \verb|plinkr| R package (\cite{plinkr}).
 This solution, however, is known to all developers that follow
@@
 Code reviews are a good way to accelerate the 
-learning of team members \cite{vable2021code}.
+learning of team members (\cite{vable2021code}).
 
@@
 Code that will stand the test of time perfectly 
-is deemed 'impossible' \cite{benureau2018re}.
+is deemed 'impossible' (\cite{benureau2018re}).
 
@@
 asking for technical report 
-(against which a 'no' is just fine \cite{barnes2010publish})
+(against which a 'no' is just fine (\cite{barnes2010publish}))
 or -more relevant to the research- sending in a bug report.
 One solution for the author is to ignore such emails (as is done in a third
-of all cases \cite{teunis2015corresponding}):
+of all cases (\cite{teunis2015corresponding})):
 it can be argued that no energy should be wasted on published code
@@
 and -after much effect by others reproducing an incorrect result-
-at the cost of a scientific career \cite{baggerly2009deriving}.
+at the cost of a scientific career (\cite{baggerly2009deriving}).
 
@@
 with the creation of documentation and code going hand-in-hand,
-as well as friendly to computers \cite{huvila2022improving}.
+as well as friendly to computers (\cite{huvila2022improving}).
 
@@
 gradually as the skills of the author(s) progress.
-To quote: Publish your code, it is good enough \cite{barnes2010publish}.
+To quote \cite{barnes2010publish}: Publish your code, it is good enough.
 

### 2023-02-13 article.tex
Review

@@
 the SNP has received its name based on a computational experiment.
-The experiment that concluded the usefulness of that SNP, 
-as well as this experiment, are based on DNA sequences.
+This earlier experiment that concluded the usefulness of that SNP, 
+is based on some DNA sequences.
+This experiment is based on assumed DNA sequences.
 DNA sequences, however, are (nowadays) not simply read, 
-yet are the result of a complex computational analysis,
+yet are the result of a complex computational analysis instead,
 with its own dedicated field of research.
@@
 In practice, however, 
-only 18\% of 180 studies are easily reproducible \cite{stodden2018empirical}.
+only 18\% of 180 computational studies 
+are easily reproducible (\cite{stodden2018empirical}).
 To some, it appears that 
@@
 \fi
-
-Code holds the ground truth of an experiment; it does the actual work.
-The more complex the computation pipeline is, the easier it is
-to have a mismatch between the article (that describes what the
-code does) and the code (that actually does the work).
-It is easy (tempting?) to overlook how easy it is to have a mismatch
-between paper and code.
 
@@
 
-It is the paper that accompanies the code,
-as it is the code that generates the results.
-As humans are fallible and code gets bigger, the likelihood of
-a mismatch between the English paper and the code increases.
-We know that the most common errors are 
-simple (and may legitimately fear that those simple errors are common) 
-(\cite{baggerly2009deriving, vable2021code}),
-and this example is not too far off from the
-examples given in that paper.
+Code holds the ground truth of an experiment; it does the actual work.
+The more complex the computation pipeline is, the easier it is
+to have a mismatch between the article (that describes what the
+code does) and the code (that actually does the work).
+The moment these two disagree, it is the code that is true.
 
@@
 Code is rarely preserved (\cite{barnes2010publish}).
-If it is, keeping it running is even more rare.
 This section discusses the preservation of code for a short, medium
@@
 with GitHub being the most popular one (\cite{cosentino2017systematic}).
-For example, the \LaTeX \hspace{5mm} code of this chapter is hosted on GitHub 
-as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
 The use of code hosting websites
@@
 as shown in Figure \ref{fig:bbbqarticleissue157}.
+Also, the \LaTeX \hspace{5mm} code of this chapter is hosted on GitHub 
+as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
 \fi
@@
 Hosted code commonly keeps a history of file changes,
-a so-called commit history,
-where a commit is the description of the change, 
-as well as the changes in one or more files.
-For example, GitHub
-uses a version control system called \verb|git| (hence its name).
-In practice, this means that when a change is made to the code,
+This means that when a change is made to the code,
 a new version is created. In the case that the change was harmful,
 one can go back to an earlier version and continue again from there.
+The version control system keeps track of who-did-what transparantly.
+\iffalse
+Figure \ref{fig:daisie_contributors}
+shows the number of commits the authors of DAISIE (\cite{etienne2020daisie})
+have made to its code.
+\fi
 It is a general recommendation to put version control
@@
 as well as openly working on the code from the start (\cite{jimenez2017four}).
-Half of 56 Science magazine corresponding authors 
-have such a version control mechanism (\cite{stodden2018empirical}).
-
-% \paragraph{Metrics help assess the honest division of labour}
-
-Most code hosts, among others, GitHub and GitLab,
-automatically keep track of certain metrics.
-One such metric is the amount of code each contributor has made.
-In that way, authors involved in the writing of the software
-are honestly acknowledged. 
-\iffalse
-Figure \ref{fig:daisie_contributors}
-shows the number of commits the authors of DAISIE (\cite{etienne2020daisie})
-have made to its code.
-\fi
-The distribution of contributed code per author 
-is useful metadata, that can automatically
-be acquired when using code hosting website.
+Half of the published code has such a version control 
+system (\cite{stodden2018empirical}).
 
@@
 
-A code host, however, will not actively preserve the uploaded code:
-if the code breaks tomorrow, it will go unnoticed.
+The degradation of software is a known feature for nearly 
+four decades.
+This is called 'bit rot' by \cite{steele1983hacker},
+or 'software collapse' by \cite{hinsen2019dealing},
+in which software fails due to dependencies on other 
+software.
+Using a code hosting website, 
+that only passively stores code,
+ignores this problem.
 
@@
 To preserve code for a longer timespan, 
-continuous integration ('CI') allows one to 
-verify if code still works and, if not, to be notified on the need for an update.
-Things change: from operating systems, to libraries that code uses,
-to the code itself. 
-Embracing those changes is recommended (\cite{beck2000extreme})
-and useful code does so.
+it needs to be embraced that software degrades (\cite{beck2000extreme}).
+Continuous integration ('CI') allows one to 
+verify if code still works and, if not, to be notified.
 
@@
 
-Some code hosts, among others, GitHub, 
-allow the researcher/user to start software scripts upon uploading a change.
-This is useful to determine the reproducibility of the code: 
-one such script should run the code (on trivial data) 
-to verify if the code works at all.
-If the functionality of the code is lost, the researcher
-can immediately see which small change was the cause of this.
-Using a CI service (such as GitHub Actions for GitHub), 
-is known to significantly 
+Some code hosts,
+allow a user to trigger specialised code upon uploading a change,
+called a CI script.
+Such a CI script typically builds and tests the code.
+This practice is known to significantly 
 increase the number of bugs exposed (\cite{vasilescu2015}) and increases
 the speed at which new features are added (\cite{vasilescu2015}).
-Using CI opens up the possibility to informally annotate features of
-a repository. 
-\iffalse
-For example, Figure \ref{fig:build_badge} shows a 
-build badge that indicates that a \LaTeX document (i.e. this paper)
-could be built.
-\fi
-
-\iffalse
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\linewidth]{build_badge.png}
-  \caption{
-    A build badge for the GitHub repository of this article.
-    The description indicates which process it is that the badge
-    signals success (as in this case) or failure of. 
-  }
-  \label{fig:build_badge}
-\end{figure}
-\fi
-
-% \paragraph{Ideal paradata}
-
-For paradata to be useful, it has to be computer-friendly, yet 
-'the best paradata does not necessarily look like 'data' at all for its
-human users' \cite{huvila2022improving}.
-There are features of code that humans find useful,
-without directly being able to measure these.
-In the end, code is just 'another kind of data' and should be designed 
-as such, for example, by using tools to work on it (\cite{wilson2022twelve}).
-
-% \paragraph{Using a tool for coding style}
-
-A first example is to use a tool to enforce a coding style 
-(e.g. the Tidyverse style guide (\cite{wickham2019advanced}) for R,
-or PEP 8 (\cite{van2001pep}) for Python),
-as following a consistent coding style improves software quality (\cite{fang2001}).
-
-% \paragraph{Using a tool for complexity}
-
-A second example is to use a tool to enforce a low cyclomatic complexity.
-The cyclomatic complexity is approximately defined 
-as the number of independent paths that
-the code can be executed. 
-\iffalse
-For example, code that has only one \verb|if| statement
-has a cyclomatic complexity of 2, as the condition within the \verb|if|
-statement can be true or false,
-resulting in the body of the \verb|if| statement being either
-executed or ignored.
-\fi
-The cyclomatic complexity correlates with code complexity,
-where more complex code is likelier to contain or give rise to bugs 
-(\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}).
-
-However, for code to work longer, one can put the code and its dependencies in
-a container, that should be best in standing the test of time,
-as the degradation of software is a known feature for nearly 
-four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
-here a more modern (and in this context superior)
-definition is used: software collapse (\cite{hinsen2019dealing}),
-in which software fails due to dependencies on other 
-software.
-Where that paper recommends using older software foundations 
-with multiple implementations (\cite{hinsen2019dealing}),
-this paper recommends a modern solution that potentially has more 
-longevity:
-to publish the code installed in a virtual environment
-(including the dependencies used at that time), as described below.
-
-Such tools combine great with CI services.
-
-Using CI does not prevent code from breaking. Instead,
-it allows one to be the first to know if code breaks.
+CI can be scheduled to run on a regular basis and
+notify the user directly when the code has broken down.
 
@@
 To preserve code for the longest time, 
-both code and its dependencies are put into a so-called container.
+both code and its dependencies can be put into a so-called container.
 
 The most reproducible way of submitting the code of an experiment,
-is by providing the code with all the (software) dependencies 
-it needs installed in a virtual container.
-This is close to the golden standard suggested by 
+is to put all code with all the (software) dependencies
+in a file that acts as a virtual computational environment,
+called a 'virtual container' (from now on: 'container').
+Such a container is close to the golden standard of reproducible research
+as suggested by 
 \cite{peng2011reproducible} 
@@
 \fi
-(however, that article was written before virtualisation existed).
-Virtualisation allows one to provide a standard environment,
-that can be used on any other computer.
-The best known examples of container software are Docker and Singularity.
-Due to access rights, Singularity is the only viable option to run
-an experiment on a 
-high performance computing (HPC) environment (e.g. a computer cluster),
-as are commonly needed in the field of genetic epidemiology experiments.
+.
 
@@
 
-To repeat features of useful paradata: paradata should be comprehensive, 
-documented appropriately, documented in co-evolution with the data,
-and the data should be friendly to computers (\cite{huvila2022improving}).
+Useful paradata, in general, are (1) comprehensive,
+(2) documented appropriately, (3) documented in co-evolution with the data,
+and (4) friendly to computers (\cite{huvila2022improving}).
 In this section, these ideal properties of paradata are applied to code.
-However, code has multiple uses, which have different requirements
-for being useful paradata.
-
-Code can be used to reproduce, to replicate and to 
-extend a computational experiment (\cite{benureau2018re}) and
-each of these uses has different requirements for the code to be useful paradata.
-In this section, the recommendations how to make paradata useful
-is applied to code and compared to existing best practices in software 
-development, for the intended use of the code.
+
+However, code has multiple uses,
+as code can be used to (1) reproduce, (2) replicate, or (3)
+extend a computational experiment (\cite{benureau2018re}).
+Depending on the intended use of the code,
+there are different requirements for code being useful paradata.
 
@@
 documented (\cite{huvila2022improving}).
-What is 'useful' is determined by its intended use: 
-to reproduce, replicate or extend (see section \ref{sec:making-code-useful-paradata}).
-The amount of documentation is, however, also determined
-by the recommended practices for writing code in general.
-
-For purposes of reproducing (same code, same data) code,
-it should at least be documented how the code is run.
-However, 57\% out of 56 Science papers with obtainable code (in 
-total there were 180 papers) have documented their code well enough to run it 
-and to be understood what it ought to do (\cite{stodden2018empirical}).
-
-However, when an independent researcher can redo an experiment, 
+
+For purposes of reproducing code,
+it should at least be documented how to run the code and what it ought to do.
+Although this may be obvious, only 57\% out of 56 Science papers 
+with obtainable code (in total there were 180 papers) 
+do so (\cite{stodden2018empirical}).
+
+When it can be found out how how an experiment is run, 
 it is possible (even ideal!) that no code is read at all.
-Within that context, it could argued that no further documentation is needed.
+Within that context, it could argued that no (further) documentation is needed.
 However, all code in general should be documented 
@@
 in which documentation and code are developed hand-in-hand.
-Literate programming is the practice of writing a document mixed with code that 
-will be run.
+Literate programming is the practice of writing code and 
+documentation in the same file.
 Contemporary examples of this idea are, among others, 
@@
 code (\cite{ram2013ropensci,ram2018community}), is 
-one example of where literate programming is enforced:
-code must have extensive runnable examples
+one example of where extensive documentation is mandatory
+and all code must have examples (that are actually run)
 as part of the documentation (\cite{ropensci_2021_6619350}).
-Developing code and documentation at the same time 
-leads to both better code and documentation (\cite{reenskaug1989environment}).
 In general, when developing software, 
@@
 to write documentation while writing software,
-as well as to include many examples (\cite{lee2018ten}).
+as well as to include many examples (\cite{lee2018ten}),
+as this leads to both better code and documentation (\cite{reenskaug1989environment}).
 
@@
 
-Code should be distributed in standard ways (\cite{peng2006reproducible})
-(see section \ref{sec:code-availability}
-for the general availability of code).
-The standard way to publish code is to use a code hosting website
+% \paragraph{Code should be distributed in standard ways}
+
+Code should be distributed in standard ways (\cite{peng2006reproducible}),
+as is done by using a code hosting website
 (see subsection \ref{subsec:code-hosting}).
@@
 Error handling is one of the mechanisms to do so.
-One example is when doing a t-test:
-to compare the means of two distributions, each distribution
-must have at least two values
-and, if not, the programming language should signal the error clearly.
 In genetic epidemiology, it is common to have incomplete or missing data,
-so analyses should take into account that there are not always enough
-observations to do a t-test.
-The analytical code should clearly signal this problem.
+so analyses should take this into account with clear error messages.
 
@@
 
-Coding errors are extremely common 
+Coding errors are extremely common (\cite{baggerly2009deriving, vable2021code})
 and contribute to the reproducibility crisis in science (\cite{vable2021code}).
 Testing, in general, is an important mechanism to ensure
-the correctness of code (an interesting example is \cite{rahman2020exploratory}
-showing bugs in scientific software on the COVID-19 pandemic).
+the correctness of code.
+One clear example is \cite{rahman2020exploratory}, 
+showing bugs in scientific software on the COVID-19 pandemic.
+
+% \paragraph{TDD}
+
+Testing is so important that it is at the
+heart of a software development methodology called 'Test-Driven 
+Development' ('TDD'), in which tests are written
+before the 'real' code.
+TDD improves code quality (\cite{alkaoud2018quality,janzen2006test})
+and it is easy to integrate 
+the writing of documentation as part of the TDD cycle (\cite{shmerlin2015document}).
 
@@
 Code coverage correlates with code quality (\cite{horgan1994,del1995correlation}).
-The non-profit organisation rOpenSci (\cite{ram2018community}),
-which peer-reviews code,
-has made it a prerequisite to have a code coverage of 100\%.
-CI opens up the possibility to informally annotate the
-code coverage. 
-For example, build badges
+and, due to this, having a code coverage of (around) 100\% 
+is mandatory to pass a code peer-review
+by rOpenSci (\cite{ram2018community}).
+When CI is activated, 
+the code coverage of a project can be shown on the repository's website
 \iffalse
@@
 \fi
-are visual indications on a website that can, among others,
-show a certain R package's code has a 100\% code coverage.
-
-% \paragraph{Test-driven development}
-
-Having a 100\% code coverage is not uncommon,
-as there is a common software development practice that 
-strives for this, called test-driven development ('TDD'). 
-TDD is a software development strategy in which tests are written
-before new 'real' code is written.
-TDD improves code quality (\cite{alkaoud2018quality,janzen2006test})
-and writing comment and documentation as part of the process
-is easy (\cite{shmerlin2015document}).
+.
 
@@
 
+\iffalse
+CI opens up the possibility to informally annotate features of
+a repository. 
+For example, Figure \ref{fig:build_badge} shows a 
+build badge that indicates that a \LaTeX document (i.e. this paper)
+could be built.
+\fi
+
+\iffalse
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\linewidth]{build_badge.png}
+  \caption{
+    A build badge for the GitHub repository of this article.
+    The description indicates which process it is that the badge
+    signals success (as in this case) or failure of. 
+  }
+  \label{fig:build_badge}
+\end{figure}
+\fi
+
 % \paragraph{Adding a license}
@@
 so that it is clear that the software can be reused.
-About two-thirds of 56 Science magazine corresponding authors 
-that shared their code supply a software license (\cite{stodden2018empirical}).
+Although this may seem trivial, 
+only two-thirds of 56 computational experiments 
+supply a software license (\cite{stodden2018empirical}).
 
@@
 
-More than half of 315 scientists 
+Code reviews are recommended by software development 
+best-practices (\cite{wilson2014best}).
+However, more than half of 315 scientists 
 have their code rarely or never reviewed (\cite{vable2021code}),
@@
 (\cite{vable2021code}).
-Within software development best-practices, 
-code reviews are recommended as well (\cite{wilson2014best}) 
-The observation that code is reviewed, makes it likelier to be correct.
 
@@
 The most reproducible way of submitting the code of an experiment,
-is by providing the code with all the (software) dependencies 
-it needs installed in a virtual container.
+is by providing the code with all its (software) dependencies 
+in a container.
 
@@
 
-% \paragraph{Runnable code allows others to re-use an experiment}
-
-Containers allow a computation experiment to be highly re-usable,
-as any scientist can work with it and tweak code.
-
-For those 22 (out of 180) Science magazine corresponding authors 
-that supplied code and that code was estimated to be 
-reproducible, one in three had the results reproduced with 
-minimal or no effort (\cite{stodden2018empirical}).
+% \paragraph{Ideal paradata}
+
+For paradata to be useful, it has to be computer-friendly, yet 
+'the best paradata does not necessarily look like 'data' at all for its
+human users' \cite{huvila2022improving}.
+There are features of code that humans find useful,
+without directly being able to measure these.
+In the end, code is just 'another kind of data' and should be designed 
+as such, for example, by using tools to work on it (\cite{wilson2022twelve}).
+
+% \paragraph{Using a tool for coding style}
+
+A first example is to use a tool to enforce a coding style 
+(e.g. the Tidyverse style guide (\cite{wickham2019advanced}) for R,
+or PEP 8 (\cite{van2001pep}) for Python),
+as following a consistent coding style improves software quality (\cite{fang2001}).
+
+% \paragraph{Using a tool for complexity}
+
+A second example is to use a tool to enforce a low cyclomatic complexity.
+The cyclomatic complexity is approximately defined 
+as the number of independent paths that
+the code can be executed. 
+\iffalse
+For example, code that has only one \verb|if| statement
+has a cyclomatic complexity of 2, as the condition within the \verb|if|
+statement can be true or false,
+resulting in the body of the \verb|if| statement being either
+executed or ignored.
+\fi
+The cyclomatic complexity correlates with code complexity,
+where more complex code is likelier to contain or give rise to bugs 
+(\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}).
 
@@
 
-For research to be reproducible, one ideally has access to
-both the data used and the code.
-In some fields, such as genetic epidemiology, the data are
-sensitive, hence cannot be released,
-yet there are methods being devised to run code on sensitive
-data with assured privacy (\cite{zhang2016review,azencott2018machine}).
-
 Next to the code, it is the data used in an experiment 
@@
 to be called 'reproducible' (\cite{peng2006reproducible}).
-When the input data are sensitive, however, and thus cannot be shared with
-an independent researcher, one thus cannot reproduce the
-experiment.
-
-The solution for this problem is simple:
-supply a simulated (also called 'analytical' (\cite{peng2006reproducible})) dataset
+In some fields, such as genetic epidemiology, the data are
+sensitive, hence cannot be released, thus one cannot reproduce 
+an experiment.
+To solve this problem in the future, there are some 
+interesting methods being developed to run code on sensitive
+data with assured privacy (\cite{zhang2016review,azencott2018machine}).
+
+To alleviate the problem today,
+a developer should supply a simulated 
+(also called 'analytical' (\cite{peng2006reproducible})) dataset
 together with the code.
+This simulated data is needed to run tests, as is 
+part of the TDD methodology.
 In the case of genetic epidemiology this would mean
@@
 as can be done with the \verb|plinkr| R package (\cite{plinkr}).
-This solution, however, is known to all developers that follow
-good practices when writing code, as public (i.e. simulated) 
-datasets are needed for automatic CI testing of code.
 One extra benefit of simulated data is that these can be used
@@
 
-One needs to be trained to know the best practices,
-before these can be applied. 
-Articles that suggest these best practices, however, claim that this initial
-investment pays off.
+To know these best practices, one needs to be trained. 
+Articles that suggest these best practices (such as this one), 
+claim that this initial investment pays off.
 Code reviews are a good way to accelerate the 
@@
 
-Code needs maintenance:
-the libraries that are used by an experiment
-will change over time, inevitably causing the experiment to fail
-over time. 
-Code that will stand the test of time perfectly 
+Code needs maintenance,
+as code that will stand the test of time perfectly 
 is deemed 'impossible' (\cite{benureau2018re}).
+CI can help a maintainer to be notified when the code breaks,
+where the use of containers may slow down time,
+as an entire computational environment is preserved.
+
+% \paragraph{Uploading code may feel like a risk}
+
+Uploading code, preferably to a code hosting website, 
+may feel like a risk, as all code can be seen and scrutinised.
+However, not publishing code may put 
+oneself in the focus of attention
+and -after much effect by others reproducing an incorrect result-
+at the cost of a scientific career (\cite{baggerly2009deriving}).
 
@@
 
-There is another problem that authors of academic code face,
-which is that they may be contacted with a user 
-asking for technical report 
+When the author of code can be contacted,
+there will be users asking for technical support 
 (against which a 'no' is just fine (\cite{barnes2010publish}))
-or -more relevant to the research- sending in a bug report.
 One solution for the author is to ignore such emails (as is done in a third
@@
 it can be argued that no energy should be wasted on published code
-and work on something new instead 
-(see \cite{barnes2010publish} for a better way to deal with the problem).
+and work on something new instead.
+However, see \cite{barnes2010publish for a better way to deal with this problem.
 
@@
 
-However, upon the find of a new bug, the question is:
-does all the research that use that software still 
-result in the same conclusions?
-To prevent authors choosing to hide bug reports, 
-a code hosting website usually has a public place to report bugs,
-to increase transparency in the handling of bug reports.
-Should there be a retraction notice on code?
+When the author of code can be contacted,
+users will send in bug reports.
+If the bug is severe enough, the question arises
+if all the research that use that code still 
+results in the same conclusions.
+One such bug is described in \cite{eklund2016cluster},
+with 40,000 studies using that incorrect code.
+These reports could be ignored to work on something new.
 
@@
 
-There are some problems with providing a container with a runnable container.
-First, a container is several gigabytes and need to be stored somewhere
-online. 
-Ideally, for Singularity, this would be at the main container hosting
-site called Syslabs. Using Syslabs is free upon registration, 
-yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
-be entered, such as the URL of the article, nor the URL of hosted code.
-
-% \paragraph{Code in container}
-
-Here are multiple suggestions to make containers with runnable code FAIR.
-To make such containers findable, the most common 
-environments (Syslabs and DockerHub) to upload containers must be indexed.
-Uploading a container to those environments is, however, limited by
-a restricted amount of free storage, with the possible consequence that
-those containers are uploaded somewhere where they escape proper indexing.
-Having containers uploaded at the existing websites ensures they are
-accessible, interoperable and reusable.
-One could argue however, that the scientific community need its own,
-similar website to upload containers, as the metadata for these
-sites are limited to free-text comments and tags (DockerHub)
-or not metadata at all (Syslabs).
-
-% \paragraph{The drawbacks of publishing code}
-
-Publishing code may be disadvantageous for an author.
-For science, code should be published, 
-this allows reproducible research 
-(see \cite{haibe2020importance} for a tragic example).
-For an author, publishing code alongside an experiment opens up
-the possibility to receive questions regarding that code.
-Note, however, that not publishing code may put 
-oneself in the focus of attention
-and -after much effect by others reproducing an incorrect result-
-at the cost of a scientific career (\cite{baggerly2009deriving}).
+One problem that containers have, is their size:
+they can be several gigabytes big. 
+This makes the distribution of these containers harder.
+Ideally, containers are stored online and distributed in
+standardised ways.
+Although progress is being made, 
+there is no way to do so for all container types.
+Additionally, probably due to their novely,
+container hosting sites lack metadata.
 
@@
 
-For the preservation of code, these are the recommendations:
+For the preservation of code, these recommendations are made:
 

### 2023-02-13 article.tex
Fix curly braces

@@
 When the author of code can be contacted,
-there will be users asking for technical support 
-(against which a 'no' is just fine (\cite{barnes2010publish}))
-One solution for the author is to ignore such emails (as is done in a third
-of all cases (\cite{teunis2015corresponding})):
+there will be users asking for technical support.
+One solution for the author is to ignore such emails,
+as is done in a third of 357 cases (\cite{teunis2015corresponding}):
 it can be argued that no energy should be wasted on published code
 and work on something new instead.
-However, see \cite{barnes2010publish for a better way to deal with this problem.
+However, see \cite{barnes2010publish} for a better way to deal with this problem.
 

### 2023-02-13 article.tex
Fix spelling

@@
 one can go back to an earlier version and continue again from there.
-The version control system keeps track of who-did-what transparantly.
+The version control system keeps track of who-did-what transparently.
 \iffalse
@@
 there is no way to do so for all container types.
-Additionally, probably due to their novely,
+Additionally, probably due to their novelty,
 container hosting sites lack metadata.

### 2023-02-13 article.tex
Update fig 1, thanks reviewer 1

@@
     \textbf{Right}: the same relations specified for genetic epidemiology.
-    Input data: genetic data, phenotypic data and associations
+    Input data: genetic data, phenotypic data and associations.
     found in earlier studies.
     Results: associations between genetic data and phenotypic data.
-    Code: the computer code used in a computational experiment
-    Paper: the scholarly paper describing the (code of the) experiment
+    Code: the computer code used in a computational experiment.
+    Paper: the scholarly paper describing the (code of the) experiment.
   }

### 2023-02-13 scripts/zip_tex_files.sh
Add script to zip the TeX files

@@ -1,0 +1,9 @@
+#!/bin/bash
+#
+# Zip the TeX files.
+#
+# Usage:
+#
+#   ./scripts/zip_text_files.sh
+#
+zip tex_files.zip $(ls {article.*,header.*,figure_1.png,svmult.cls,svind.ist})

### 2023-04-11 article.tex
Improve definition of paradata

@@
 {\bf Keywords:} paradata, reproducible research, code, software,
-FAIR data, computational research, Open Science, best practices,
+computational research, Open Science, best practices,
 genetic epidemiology
@@
 
-This paper defines paradata as 'data about the data collection 
-process' (\cite{choumert2019using}).
-This definition is chosen for its simplicity, 
-yet there are multiple related and/or more nuanced 
-definitions of paradata (see for example \cite{huvila2022improving} 
-and \cite{skold2022interrogating}).
+There exists no standard definition of paradata (\cite{nicolaas2011survey,skold2022interrogating,huvila2022improving}).
+Without declaring it a formal definition, however,
+paradata is 'data about processes' (and example survey paradata 
+are call records and audit trails) \cite{nicolaas2011survey},
+or 'data about the data collection process' (\cite{choumert2019using}).
+This paper uses the latter as a working definition.
 
@@
 
-One problem that containers have, is their size:
-they can be several gigabytes big. 
-This makes the distribution of these containers harder.
+Containers do have problems.
+First, they themselves require software to run,
+with the same software decay being possible.
+Second, one needs to install that software and have the
+computer access rights (i.e. admin rights under Windows, 
+or root rights under Linux) to do so.
+Third, one needs to learn how to build and use containers.
+Lastly, containers can be several gigabytes big files,
+which makes their distribution harder.
 Ideally, containers are stored online and distributed in

### 2023-04-11 article.tex
Merge single-line paragraphs

@@
 both code and its dependencies can be put into a so-called container.
-
 The most reproducible way of submitting the code of an experiment,
@@
 In this section, these ideal properties of paradata are applied to code.
-
 However, code has multiple uses,
@@
 documented (\cite{huvila2022improving}).
-
 For purposes of reproducing code,
@@
 (see subsection \ref{subsec:code-hosting}).
-
 % \paragraph{Code must have extensive error handling}
-
-Code must be more extensive when it is (intended to be) 
+Additionally, code must be more extensive when it is (intended to be) 
 used on different data,
@@
 in a container.
-
 % \paragraph{Runnable code can precisely reproduce an experiment}
-
 Containers allow a computation experiment to be highly reproducible:
@@
 as following a consistent coding style improves software quality (\cite{fang2001}).
-
 % \paragraph{Using a tool for complexity}
-
 A second example is to use a tool to enforce a low cyclomatic complexity.
@@
 the code. 
-
 When applying these general recommendations to code, 

### 2023-04-11 article.tex
Add author biography

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Author biography}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Dr. Bilderbeek has worked in computational biology for more than a decade.
+Currently, he works for the National Bioinformatics Infrastructure Sweden (NBIS)
+in Uppsala, Sweden, 
+as an application expert and trainer, with a focus on reproducible research.
+All his work can be found on GitHub at \url{https://github.com/richelbilderbeek}
+and his YouTube channel.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography

### 2023-04-12 article.tex
Add suggestion, thanks Isto

@@
 Without declaring it a formal definition, however,
-paradata is 'data about processes' (and example survey paradata 
-are call records and audit trails) \cite{nicolaas2011survey},
-or 'data about the data collection process' (\cite{choumert2019using}).
+paradata can be understood in general sense as 'data about processes’ (with e.g. survey paradata
+termed alternatively as records and audit trails) \cite{nicolaas2011survey},
+or in more specific sense in relation to data collection, as 'data about the data collection process' (\cite{choumert2019using}).
 This paper uses the latter as a working definition.

### 2023-10-08 feedback/20230927/article.tex
Organize feedback in folders

--- 2023-04-12 article.tex
@@
 
+% IH: replaced references to "this paper" with "this chapter". Check that ok.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 or in more specific sense in relation to data collection, as 'data about the data collection process' (\cite{choumert2019using}).
-This paper uses the latter as a working definition.
+This chapter uses the latter as a working definition.
 
@@
 it can help understand how scholarly knowledge is produced (\cite{huvila2022improving}).
-This paper discusses these general reasons applied to code.
+% IH: As the series editors asked that all chapters should in the end somehow link to knowledge management, may be here could be a place to add a few words about it e.g. see below
+% IH: and to integrate this understanding as a part of the management of scholarly knowledge.
+This chapter discusses these general reasons applied to code.
 
@@
 
-This paper uses genetic epidemiology as a specific example,
+This chapter uses genetic epidemiology as a specific example,
 to illustrate in which way code is paradata,
@@
 (see Table \ref{tab:definitions} in the Supplementary Materials
-for the definitions of terms used in this paper)
+for the definitions of terms used in this chapter)
 \fi
@@
 
-This paper started with some suggestions to 
+This chapter started with some suggestions to 
 make paradata useful for data re(use):
@@
 
-Dr. Bilderbeek has worked in computational biology for more than a decade.
-Currently, he works for the National Bioinformatics Infrastructure Sweden (NBIS)
-in Uppsala, Sweden, 
-as an application expert and trainer, with a focus on reproducible research.
-All his work can be found on GitHub at \url{https://github.com/richelbilderbeek}
-and his YouTube channel.
+% IH: Moved to a separate document
+%Dr. Bilderbeek has worked in computational biology for more than a decade.
+%Currently, he works for the National Bioinformatics Infrastructure Sweden (NBIS)
+%in Uppsala, Sweden, 
+%as an application expert and trainer, with a focus on reproducible research.
+%All his work can be found on GitHub at \url{https://github.com/richelbilderbeek}
+%and his YouTube channel.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{This paper}
+\subsection{This chapter}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 at \url{https://github.com/richelbilderbeek/chapter_paradata}.
-Due this, one can, among others, see the complete version history of this paper.
+Due this, one can, among others, see the complete version history of this chapter.
 This alone could allow research to be done on, for example,
@@
   \end{tabular}
-  \caption{Terms used in this paper, and their definitions}
+  \caption{Terms used in this chapter, and their definitions}
   \label{tab:definitions}
@@
   \end{tabular}
-  \caption{Recommendations described in this paper}
+  \caption{Recommendations described in this chapter}
   \label{tab:recommendations}
